### 2017-08-29 Assignment-1/Solns-Assgn-1.tex
Create Solns-Assgn-1.tex

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}

\title{Solutions to the Assignment - 1 : CS6230 - \\
Optimization Methods in Machine Learning}
\author{Vishwak Srinivasan\\
\texttt{CS15BTECH11043}}
\date{}

\begin{document}
\maketitle

\section*{Question 1}
\begin{flushleft}
Without loss of generality, assume that \(C\) is compact and \(D\) is just closed, since we know that \(C - D\) is closed. Now we need to show that \(\exists c^{*} \in C, d^{*} \in D\) such that \(\displaystyle c^{*}, d^{*} = \text{arg} \inf_{x \in C, y \in D} d(x, y)\). This can be seen from the definition of the compactness of \(C\) and the closedness of \(D\). 
\(\newline\)

Now construct \(z = c^{*} - d^{*}\). Let us show the following: \(\displaystyle \inf_{x \in C} x^{T}z \geq c^{*^{T}}z\). We prove this by contradiction:
\(\newline\)

Assume that \(\displaystyle \inf_{x \in C} x^{T}z < c^{*^{T}}z\). Since \(x\) and \(c^{*} \in C\), the convex combination of \(x\) and \(c^{*}\) must \(\in C\). Now let us find out the distance between \(x_{\epsilon} = \epsilon x + (1 - \epsilon) c^{*}\) and \(d^{*}\). This is equal to:
\begin{gather}
||x_{\epsilon} - d^{*}||^{2} = ||\epsilon x + (1 - \epsilon)c^{*} - d^{*}||^{2} = ||c^{*} - d^{*} + \epsilon(x - c^{*})||^{2} \\
\implies ||x_{\epsilon} - d^{*}||^{2} = \langle c^{*} - d^{*} + \epsilon(x - c^{*}) , c^{*} - d^{*} + \epsilon(x - c^{*}) \rangle \\
\implies ||x_{\epsilon} - d^{*}||^{2} = ||c^{*} - d^{*}||^{2} + \epsilon^{2}||x - c^{*}||^{2} + 2\epsilon\langle x - c^{*}, c^{*} - d^{*}\rangle
\end{gather}

From the assumption taken, we know that \(\langle x - c^{*} , z\rangle < 0\). The last term in equation 3 will use this fact and hence for a choice of \(\epsilon\), \(||x_{\epsilon} - d^{*}||^{2} < ||c^{*} - d^{*}||^{2}\), thereby violating the fact that \(c^{*}, d^{*}\) are the closest points from \(C, D\) respectively.
\(\newline\)

We also know that: \(\langle c^{*}, z \rangle > \langle d^{*}, z \rangle\). This can be seen by sending the \(\langle d^{*}, z \rangle\) term to the LHS, and then that becomes \(||z||^{2}\). 
\(\newline\)

Just as how we proved the claim that \(\displaystyle \inf_{x \in C} x^{T}z \geq c^{*^{T}}z\), we will prove \(\displaystyle d^{*^{T}}z \geq \sup_{y \in D} y^{T}z\). Assume that \(d^{*^{T}}z < \sup_{y\in D} y^{T}z\). Now, consider the distance between \(y_{\epsilon} = \epsilon y + (1-\epsilon)d^{*}\) and \(c^{*}\). We know that \(y_{\epsilon} \in D\) by convexity of \(D\).

\begin{gather}
||c^{*} - y_{\epsilon}||^{2} = ||c^{*} - \epsilon y - (1 - \epsilon)d^{*}||^{2} = ||c^{*} - d^{*} + \epsilon(d^{*} - y)||^{2} \\
\implies ||c^{*} - y_{\epsilon}||^{2} = \langle c^{*} - d^{*} + \epsilon(d^{*} - y) , c^{*} - d^{*} + \epsilon(d^{*} - y) \rangle \\
\implies ||c^{*} - y_{\epsilon}||^{2} = ||c^{*} - d^{*}||^{2} + \epsilon^{2}||d^{*} - y||^{2} + 2\epsilon\langle c^{*} - d^{*}, d^{*} - y \rangle 
\end{gather}

From the assumption taken, we know that \(\langle d^{*} - y, z\rangle < 0\). Again, the last term in equation 6 will use this fact and hence for choice of \(\epsilon\), \(||c^{*} - y_{\epsilon}||^{2} < ||c^{*} - d^{*}||^{2}\), thereby violating the fact that \(c^{*}, d^{*}\) are the closest points from \(C, D\) respectively. 
\(\newline\)

Using the above facts: \(\displaystyle \inf_{x \in C} x^{T}z \geq c^{*^{T}}z \geq d^{*^{T}}z \geq \sup_{y \in D} y^{T}z \implies \inf_{x \in C} x^{T}z > \sup_{y \in D} y^{T}z\). Hence proved.
\end{flushleft}

\section*{Question 2}
\begin{flushleft}
Consider the functions \(f_{1}(x) = \sqrt{|x|}\) and \(f_{2}(x) = \sqrt{|x - 1|}\). Now we know that the \(\alpha\)-sublevel sets of both \(f_{1}\) and \(f_{2}\) are convex for all \(\alpha\), but the functions \(f_{1}\) and \(f_{2}\) are themselves not convex. This makes them quasiconvex functions. 
\(\newline\)

Now take the sum of these functions. Let \(f(x) = f_{1}(x) + f_{2}(x) = \sqrt{|x|} + \sqrt{|x-1|}\). Note that this function has two local minima: one at \(x = 0\) and the other at \(x = 1\). Consider \(\alpha \in [1, \sqrt{2}]\). Note that the \(\alpha\)-sublevel sets for this choice of \(\alpha\) is not convex, hence telling us that the sum of two quasi-convex functions is not necessarily convex.
\end{flushleft}

\section*{Question 4}
\begin{flushleft}
Proving the implication: if \(f\) is convex, then inequality in the question holds:
\(\newline\)

If \(f\) is convex: then:
\begin{equation}
\label{4_1}
f(y) \geq f(x) + \nabla f(x)^{T} (y - x)
\end{equation}
Inverting \(x\) and \(y\) we get:
\begin{equation}
\label{4_2}
f(x) \geq f(y) + \nabla f(y)^{T} (x - y)
\end{equation}

Adding equations \ref{4_1} and \ref{4_2}, we get:
\begin{equation}
\left(\nabla f(x)^{T} -  \nabla f(y)^{T}\right)(y-x) \leq 0 \implies \left(\nabla f(x)^{T} - \nabla f(y)^{T}\right) ( x - y) \geq 0
\end{equation}

Proving the implication: if the condition in the question holds, then the function is convex
\(\newline\)

Now from \((\nabla f(\mathbf{x}) - \nabla f(\mathbf{y}))^{T}(x - y ) \geq 0\). This suggests that \(\nabla f\) is increasing. Consider a function: \(g(\theta) = f(\theta\mathbf{x} + (1 - \theta)\mathbf{y}) \implies g'(0) = \nabla f(\mathbf{y})^{T}(\mathbf{y} - \mathbf{x})\).

Now we know that: \(g(1) - g(0) = \displaystyle \int_{0}^{1} g'(\theta) d\theta \geq g'(0) d\theta\). This is because \(\nabla f\) is non-decreasing as shown above. \(g(1) - g(0) \geq g'(0) = \nabla f(\mathbf{x})^{T} (\mathbf{y} - \mathbf{x}) \implies g(0) \geq g(1) + \nabla f(\mathbf{y})^{T}(\mathbf{x} -\mathbf{y})\). From this: we get: \(f(\mathbf{x}) \geq f(\mathbf{y}) + \nabla f(\mathbf{y})^{T}(\mathbf{x} - \mathbf{y})\), which is the first order condition.
\end{flushleft}

\section*{Question 8}
\subsection*{Question 8.a.}
\begin{flushleft}
\begin{equation}
\label{8_1_1}
f(\mathbf{x}) = \displaystyle - \sum_{i=1}^{n} \log(x_{i}) = - \log(x_{1}) - \log(x_{2}) - \ldots - \log(x_{n})
\end{equation}

Calculating the gradient of equation \ref{8_1_1} we get:
\begin{equation}
\nabla f(\mathbf{x}) = \langle -\frac{1}{x_{1}}, -\frac{1}{x_{2}}, \ldots, -\frac{1}{x_{n}} \rangle
\end{equation}

Calculating the Hessian of equation \ref{8_1_1} we get:
\begin{equation}
\displaystyle \nabla^{2} f(\mathbf{x}) [i,j] = \begin{cases}
\frac{1}{x^{2}_{i}} & \text{if } i = j \\
0 & \text{if } i \neq j
\end{cases}
\end{equation}

Now note that the eigenvalues of \(\nabla^{2} f(\mathbf{x})\) are \(\Lambda = \lbrace\frac{1}{x^{2}_{1}}, \frac{1}{x^{2}_{2}}, \ldots, \frac{1}{x^{2}_{n}}\rbrace\). Since the domain of \(f\) is \(\mathbb{R}^{n}_{++}\), we can most certainly assure that \(\nexists \lambda \in \Lambda\), such that \(\lambda \leq 0\). Hence it can be said that \(f\) is \(k\)-strongly convex with \(k = \min \Lambda\).
\end{flushleft}

\subsection*{Question 8.b.}
\begin{flushleft}
\begin{equation}
\label{8_2_1}
f(\mathbf{x}) = \displaystyle - \sum_{i=1}^{n} x_{i} \log(x_{i}) = -x_{1}\log(x_{1}) -x_{2}\log(x_{2}) -\ldots -x_{n}\log(x_{n})
\end{equation}

Calculating the gradient of equation \ref{8_2_1} we get:
\begin{equation}
\nabla f(\mathbf{x}) = \langle -(1 + \log(x_{1})), -(1 + \log(x_{2})), \ldots, -(1 + \log(x_{n})) \rangle
\end{equation}

Calculating the Hessian of equation \ref{8_2_1} we get:
\begin{equation}
\displaystyle \nabla^{2} f(\mathbf{x}) [i,j] = \begin{cases}
-\frac{1}{x_{i}} & \text{if } i = j \\
0 & \text{if } i \neq j
\end{cases}
\end{equation}

Note that this matrix is diagonal and the eigenvalues of this matrix are \(\Lambda = \lbrace\frac{-1}{x_1}, \frac{-1}{x_2}, \ldots, \frac{-1}{x_n}\rbrace\). Since the domain of \(f\) is \(\mathbb{R}^{n}_{+}\), we can definitely tell that the eigenvalues are all less than \(0\), hence telling us that the function is non-convex.
\end{flushleft}
\section*{Question 9}
\subsection*{Question 9.a.}
\begin{flushleft}
\textit{Claim: } \(\displaystyle f(\mathbf{x}) = \frac{1}{\mathbf{x}^T P \mathbf{x} + 1}\) is quasi-concave for \(P \in \mathbb{S}^{n}_{++}\). 
\(\newline\)

\textit{Proof: } Consider the \(\alpha\)-superlevel sets of \(f(\mathbf{x})\). Denote it by \(\mathcal{U}_{\alpha}\).
\begin{equation}
\mathcal{U}_{\alpha} = \lbrace \mathbf{x} : \frac{1}{\mathbf{x}^{T} P \mathbf{x} + 1} \geq \alpha \rbrace
\end{equation}

Assuming \(\alpha > 0\), we get: \(\displaystyle \frac{1}{\alpha} - 1 \geq \mathbf{x}^{T} P \mathbf{x} \geq 0 \implies \alpha \in (0, 1)\). Consider \(\mathbf{y}\) also such that \(\displaystyle \frac{1}{\alpha} - 1 \geq \mathbf{y}^{t} P \mathbf{y}\). Now for all \(\theta \in [0,1]\), \(\mathbf{z} = \theta\mathbf{x} + (1 - \theta)\mathbf{y}\). We will show that: \(\mathbf{z}^{T} P \mathbf{z} \leq \frac{1}{\alpha} - 1\). We know that \(\displaystyle \mathbf{z}^{T} P \mathbf{z} \leq \theta^{2} \mathbf{x}^{T} P \mathbf{x} + (1 - \theta)^2 \mathbf{y}^{T} P \mathbf{y} \leq \theta^{2} \left(\frac{1}{\alpha} - 1\right) + (1 - \theta)^2 \left(\frac{1}{\alpha} - 1\right) = \frac{1}{\alpha} - 1\). Hence we now know that the \(\alpha\)-superlevel set of \(f(\mathbf{x})\) for \(\alpha \in (0, 1)\) is convex.
\(\newline\)

Now if \(\alpha = 1\) then, we get \(\mathbf{x}^{T} P \mathbf{x} \leq 0\), and since \(P \in \mathbb{S}^{n}_{++}\), \(\mathbf{x}^{T} P \mathbf{x} = 0\) is only possible, and \(\mathbf{x} = \mathbf{0} \implies \mathcal{U}_{1} = \{\mathbf{0}\}\). A singleton set is convex too.
\(\newline\)

Now if \(\alpha > 1\) then, we get \(\mathbf{x}^{T} P \mathbf{x} \leq \frac{1}{\alpha} - 1 < 0 \implies \mathbf{x}^{T} P \mathbf{x} < 0\). We know that \(\nexists \mathbf{x}\) for which this holds \(\implies U_{\alpha > 1} = \emptyset\) , and we know that an empty set is convex too.
\(\newline\)

Consider \(\alpha = 0\). Now we get \(1 \geq 0\), which is true for all \(\mathbf{x} \in \mathbb{R}^{n}\), hence \(\mathcal{U}_{0} = \mathbb{R}^{n}\), which is convex.
\(\newline\)

Last case, consider \(\alpha < 0\). Now we get \(\mathbf{x}^{T} P \mathbf{x} \geq \displaystyle \frac{1}{\alpha} - 1\). Since \(\alpha < 0 \implies \frac{1}{\alpha} < 0 \implies \frac{1}{\alpha} - 1 < 0\). But we know that \(\mathbf{x}^{T} P \mathbf{x} \geq 0\), and hence \(\mathcal{U}_{\alpha < 0} = \mathbb{R}^{n}\) which is convex.
\(\newline\)

Summarizing the cases:
\begin{center}
\begin{tabular}{| c | c |}
\hline
Choice of \(\alpha\) & \(\mathcal{U}_{\alpha}\) \\
\hline
\hline
\(\alpha > 1\) & \(\emptyset\) \\
\hline
\(\alpha = 1\) & \(\{\mathbf{0}\}\) \\
\hline
\(\alpha \in (0, 1)\) & A convex set \\
\hline
\(\alpha = 0\) & \(\mathbb{R}^{n}\) \\
\hline
\(\alpha < 0\) & \(\mathbb{R}^{n}\) \\
\hline
\end{tabular}
\end{center}

From the aforementioned cases, it should be obvious that no matter what \(\alpha\) we choose, we get a convex superlevel set, hence proving the claim that \(f(\mathbf{x})\) is quasi-concave.

\end{flushleft}
\subsection*{Question 9.b.}
\begin{flushleft}
Using the composition property stated in Pg. 102 in Boyd and Vanderberghe, where \(A = \mathbb{I}, b = \mathbf{0}\) and \(\mathbf{c} = \mathbf{a}, d = 1\) (given in this question) and the facts that:
\begin{itemize}
\item \(f(\mathbf{x}) = \mathbf{x}^{T} P \mathbf{x}\) is convex, and hence quasiconvex
\item \(\displaystyle f\left(\mathbf{\frac{x}{\mathbf{a}^{T}\mathbf{x} + 1}}\right) = \left(\frac{\mathbf{x}^{T} P \mathbf{x}}{(\mathbf{a}^{T}\mathbf{x} + 1)^2}\right)\) is hence quasiconvex from the fact mentioned above in Boyd and Vanderberghe.
\end{itemize}
\end{flushleft}

\end{document}

### 2017-08-29 Programming-Assignment-1/fixed_step_size.py
PA-1 Update 1

@@ -1,0 +1,14 @@
+import numpy as np
+from matplotlib import pyplot as plt
+
+def quad(x, y):
+	return 1.125*x**2 + 0.5*x*y + 0.75*y**2 + 2*x + 2*y
+	
+def log_reg(x, y):
+	return 0.5*(x**2 + y**2) + 50*np.log(1 + np.exp(-0.5*y)) + 50*(1 + np.exp(0.2*x))
+	
+def himmelblaus(x, y):
+	return 0.1*(x**2 + y - 11)**2 + 0.1*(x + y**2 - 7)**2
+	
+def rosenbrock(x, y):
+	return 0.002*(1 - x)**2 + 0.2*(y*x**2)**2

### 2017-08-29 Programming-Assignment-1/fixed_step_size.py
PA-1 Update 2

@@ -1,14 +1,12 @@
 import numpy as np
+from functions import *
 from matplotlib import pyplot as plt
 
-def quad(x, y):
-	return 1.125*x**2 + 0.5*x*y + 0.75*y**2 + 2*x + 2*y
-	
-def log_reg(x, y):
-	return 0.5*(x**2 + y**2) + 50*np.log(1 + np.exp(-0.5*y)) + 50*(1 + np.exp(0.2*x))
-	
-def himmelblaus(x, y):
-	return 0.1*(x**2 + y - 11)**2 + 0.1*(x + y**2 - 7)**2
-	
-def rosenbrock(x, y):
-	return 0.002*(1 - x)**2 + 0.2*(y*x**2)**2
+def update(cur_point, lr, func):
+	gradients = {'quad'		: grad_quad,
+		     'log_reg'		: grad_log_reg,
+		     'himmelblaus'	: grad_himmelblaus,
+		     'rosenbrock'	: grad_rosenbrock
+		    }
+	updt	= cur_point - lr*gradients[func](*cur_point)
+	return updt

### 2017-08-29 Programming-Assignment-1/functions.py

@@ -1,0 +1,51 @@
+import numpy as np
+
+def quad(x, y):
+	return 1.125*x**2 + 0.5*x*y + 0.75*y**2 + 2*x + 2*y
+	
+def log_reg(x, y):
+	return 0.5*(x**2 + y**2) + 50*np.log(1 + np.exp(-0.5*y)) + 50*np.log(1 + np.exp(0.2*x))
+	
+def himmelblaus(x, y):
+	return 0.1*(x**2 + y - 11)**2 + 0.1*(x + y**2 - 7)**2
+	
+def rosenbrock(x, y):
+	return 0.002*(1 - x)**2 + 0.2*(y - x**2)**2
+	
+def grad_quad(x, y):
+	dx	= 2.25*x + 0.5*y + 2
+	dy	= 1.5*y + 0.5*x + 2
+	return np.array([dx, dy])
+	
+def grad_log_reg(x, y):
+	dx1	= x
+	dx2	= 0
+	dx3	= 10*np.exp(0.2*x)/(1 + np.exp(0.2*x))
+	dy1	= y
+	dy2	= -25*np.exp(-0.5*y)/(1 + np.exp(-0.5*y))
+	dy3	= 0
+	dx	= dx1 + dx2 + dx3
+	dy	= dy1 + dy2 + dy3
+	return np.array([dx, dy])
+	
+def grad_himmelblaus(x, y):
+	dt1	= 0.2*(x**2 + y - 11)
+	dt2	= 0.2*(x + y**2 - 7)
+	dt1x	= 2*x
+	dt2x	= 1
+	dt1y	= 1
+	dt2y	= 2*y
+	dx	= dt1*dt1x + dt2*dt2x
+	dy	= dt1*dt1y + dt2*dt2y
+	return np.array([dx, dy])
+	
+def grad_rosenbrock(x, y):
+	dt1	= 0.004**(1 - x)
+	dt2	= 0.4*(y - x**2)
+	dt1x	= -1
+	dt2x	= -2*x
+	dt1y	= 0
+	dt2y	= 1
+	dx	= dt1*dt1x + dt2*dt2x
+	dy	= dt1*dt1y + dt2*dt2y
+	return np.array([dx, dy])

### 2017-08-29 Programming-Assignment-1/fixed_step_size.py
PA-1 Update 3

@@ -1,7 +1,9 @@
 import numpy as np
 from functions import *
+from argparse import ArgumentParser
 from matplotlib import pyplot as plt
 
+# Fixed Step Size Updates
 def update(cur_point, lr, func):
 	gradients = {'quad'		: grad_quad,
 		     'log_reg'		: grad_log_reg,
@@ -10,3 +12,58 @@ def update(cur_point, lr, func):
 		    }
 	updt	= cur_point - lr*gradients[func](*cur_point)
 	return updt
+			
+# Argument parser for arguments
+parser	= ArgumentParser()
+parser.add_argument('--func_name', required=True, help='quad | log_reg | himmelblaus | rosenbrock')
+parser.add_argument('--init_type', required=True, help='rand | static')
+parser.add_argument('--lr', required=True, type=float, help='learning rate')
+opt	= parser.parse_args()
+
+# Set initial point
+if opt.init_type == 'rand':
+	xs	= np.random.uniform(low=-6, high=6, size=2)
+
+elif opt.init_type == 'static':
+	xs	= np.array([2, 3])
+	
+plt.subplot(111)
+plt.xlabel('$x$', size=20)
+plt.ylabel('$y$', size=20)
+
+if opt.func_name == 'quad':
+	levels	= np.linspace(0, 110, 88)
+	xmin, xmax, xstep	= -6, 6, 0.05
+	ymin, ymax, ystep	= -6, 6, 0.05
+	X, Y	= np.meshgrid(np.arange(xmin, xmax + xstep, xstep), np.arange(ymin, ymax + ystep, ystep))
+	Z	= quad(X, Y)
+
+elif opt.func_name == 'log_reg':
+	levels	= np.linspace(0, 265, 212)
+	xmin, xmax, xstep	= -6, 6, 0.05
+	ymin, ymax, ystep	= -6, 6, 0.05
+	X, Y	= np.meshgrid(np.arange(xmin, xmax + xstep, xstep), np.arange(ymin, ymax + ystep, ystep))
+	Z	= log_reg(X, Y)
+
+elif opt.func_name == 'himmelblaus':
+	levels	= np.linspace(0, 220, 176)
+	xmin, xmax, xstep	= -6, 6, 0.05
+	ymin, ymax, ystep	= -6, 6, 0.05
[…]

### 2017-08-29 Programming-Assignment-1/functions.py

@@ -40,7 +40,7 @@ def grad_himmelblaus(x, y):
 	return np.array([dx, dy])
 	
 def grad_rosenbrock(x, y):
-	dt1	= 0.004**(1 - x)
+	dt1	= 0.004*(1 - x)
 	dt2	= 0.4*(y - x**2)
 	dt1x	= -1
 	dt2x	= -2*x

### 2017-08-29 Programming-Assignment-1/backtracking_ls.py
PA-1 Update 4

@@ -1,23 +1,39 @@
 import numpy as np
+from numpy.linalg import norm
 from functions import *
 from argparse import ArgumentParser
 from matplotlib import pyplot as plt
 
 # Fixed Step Size Updates
-def update(cur_point, lr, func):
-	gradients = {'quad'		: grad_quad,
-		     'log_reg'		: grad_log_reg,
-		     'himmelblaus'	: grad_himmelblaus,
-		     'rosenbrock'	: grad_rosenbrock
-		    }
-	updt	= cur_point - lr*gradients[func](*cur_point)
-	return updt
+def update(cur_point, func):
+	gradients	= {'quad'		: grad_quad,
+		 	   'log_reg'		: grad_log_reg,
+			   'himmelblaus'	: grad_himmelblaus,
+		    	   'rosenbrock'		: grad_rosenbrock
+		    	  }
+	function	= {'quad'		: quad,
+			   'log_reg'		: log_reg,
+			   'himmelblaus'	: himmelblaus,
+			   'rosenbrock'		: rosenbrock
+			  }
+	lr	= 1
+	alpha	= 0.5
+	beta	= 0.5
+	fn	= function[func]
+	grd	= gradients[func]
+	i	= 0
+	
+	while fn(*(cur_point - lr*grd(*cur_point))) > fn(*cur_point) - alpha*lr*norm(grd(*cur_point))**2 and i < 10:
+		lr	= beta*lr
+		i	= i + 1
+	
+	updt	= cur_point - lr*grd(*cur_point)
+	return updt, lr
 			
 # Argument parser for arguments
 parser	= ArgumentParser()
 parser.add_argument('--func_name', required=True, help='quad | log_reg | himmelblaus | rosenbrock')
 parser.add_argument('--init_type', required=True, help='rand | static')
-parser.add_argument('--lr', required=True, type=float, help='learning rate')
 opt	= parser.parse_args()
 
 # Set initial point
@@ -61,9 +77,10 @@ elif opt.func_name == 'rosenbrock':
 
 plt.contour(X, Y, Z, levels=levels, cmap=plt.cm.jet)
 
+lr = 'init'
 for i in range(0, 1000):
 	plt.plot(*xs, 'k.', markersize=5)
-	xs	= update(xs, opt.lr, opt.func_name)
[…]

### 2017-08-29 Programming-Assignment-1/inv_step_size.py

@@ -17,7 +17,6 @@ def update(cur_point, lr, func):
 parser	= ArgumentParser()
 parser.add_argument('--func_name', required=True, help='quad | log_reg | himmelblaus | rosenbrock')
 parser.add_argument('--init_type', required=True, help='rand | static')
-parser.add_argument('--lr', required=True, type=float, help='learning rate')
 opt	= parser.parse_args()
 
 # Set initial point
@@ -63,7 +62,7 @@ plt.contour(X, Y, Z, levels=levels, cmap=plt.cm.jet)
 
 for i in range(0, 1000):
 	plt.plot(*xs, 'k.', markersize=5)
-	xs	= update(xs, opt.lr, opt.func_name)
-	plt.title('Iteration ${0}$ Current Point: $({1}, {2})$'.format(i+1, round(xs[0], 5), round(xs[1], 5)), size=20)
+	xs	= update(xs, 1/(i+1), opt.func_name)
+	plt.title('Iteration ${0}$ Learning Rate: ${1}$ Current Point: $({2}, {3})$'.format(i+1, round(1/(i+1), 5), round(xs[0], 5), round(xs[1], 5)), size=20)
 	plt.pause(0.0001)
 plt.show()

### 2017-08-30 Programming-Assignment-1/backtracking_ls.py
PA-1 Update 5

@@ -77,10 +77,25 @@ elif opt.func_name == 'rosenbrock':
 
 plt.contour(X, Y, Z, levels=levels, cmap=plt.cm.jet)
 
+all_updates	= []
+all_bcktrck	= []
 lr = 'init'
 for i in range(0, 1000):
-	plt.plot(*xs, 'k.', markersize=5)
-	plt.title('Iteration ${0}$ Learning Rate = ${1}$ Current Point: $({2}, {3})$'.format(i+1, lr, round(xs[0], 5), round(xs[1], 5)), size=20)
+	all_updates.append(xs)
 	xs, lr	= update(xs, opt.func_name)
-	plt.pause(0.0001)
+	all_bcktrck.append(np.array([i+1, -np.log2(lr)]))
+all_updates.append(xs)
+
+all_updates	= np.array(all_updates)
+plt.plot(all_updates[:,0], all_updates[:,1], 'k.-')
+plt.title('Iteration ${0}$ Current Point: $({1}, {2})$'.format(i+1, round(xs[0], 5), round(xs[1], 5)), size=20)
 plt.show()
+
+plt.clf()
+all_bcktrck	= np.array(all_bcktrck)
+plt.xlim(0, 1001)
+plt.ylim(-0.5, 10.5)
+plt.yticks(np.arange(0, 11, 1.0))
+plt.plot(all_bcktrck[:,0], all_bcktrck[:,1], 'k.')
+plt.title('Variation of Learning Rate with iteration')
+plt.show() 

### 2017-08-30 Programming-Assignment-1/fixed_step_size.py

@@ -61,9 +61,13 @@ elif opt.func_name == 'rosenbrock':
 
 plt.contour(X, Y, Z, levels=levels, cmap=plt.cm.jet)
 
+all_updates	= []
 for i in range(0, 1000):
-	plt.plot(*xs, 'k.', markersize=5)
+	all_updates.append(xs)
 	xs	= update(xs, opt.lr, opt.func_name)
-	plt.title('Iteration ${0}$ Current Point: $({1}, {2})$'.format(i+1, round(xs[0], 5), round(xs[1], 5)), size=20)
-	plt.pause(0.0001)
+all_updates.append(xs)
+
+all_updates	= np.array(all_updates)
+plt.plot(all_updates[:,0], all_updates[:,1], 'k.-')
+plt.title('Iteration ${0}$ Current Point: $({1}, {2})$'.format(i+1, round(xs[0], 5), round(xs[1], 5)), size=20)
 plt.show()

### 2017-08-30 Programming-Assignment-1/inv_step_size.py

@@ -60,9 +60,13 @@ elif opt.func_name == 'rosenbrock':
 
 plt.contour(X, Y, Z, levels=levels, cmap=plt.cm.jet)
 
+all_updates	= []
 for i in range(0, 1000):
-	plt.plot(*xs, 'k.', markersize=5)
+	all_updates.append(xs)
 	xs	= update(xs, 1/(i+1), opt.func_name)
-	plt.title('Iteration ${0}$ Learning Rate: ${1}$ Current Point: $({2}, {3})$'.format(i+1, round(1/(i+1), 5), round(xs[0], 5), round(xs[1], 5)), size=20)
-	plt.pause(0.0001)
+all_updates.append(xs)
+
+all_updates	= np.array(all_updates)
+plt.plot(all_updates[:,0], all_updates[:,1], 'k.-')
+plt.title('Iteration ${0}$ Current Point: $({2}, {3})$'.format(i+1, round(1/(i+1), 5), round(xs[0], 5), round(xs[1], 5)), size=20)
 plt.show()

### 2017-08-30 Programming-Assignment-1/backtracking_ls.py
PA-1 Update 6

@@ -38,7 +38,12 @@ opt	= parser.parse_args()
 
 # Set initial point
 if opt.init_type == 'rand':
-	xs	= np.random.uniform(low=-6, high=6, size=2)
+	if opt.func_name == 'rosenbrock':
+		x	= np.random.uniform(low=-3, high=3)
+		y	= np.random.uniform(low=-6, high=6)
+		xs	= np.array([x, y])
+	else:
+		xs	= np.random.uniform(low=-6, high=6, size=2)
 
 elif opt.init_type == 'static':
 	xs	= np.array([2, 3])
@@ -79,7 +84,6 @@ plt.contour(X, Y, Z, levels=levels, cmap=plt.cm.jet)
 
 all_updates	= []
 all_bcktrck	= []
-lr = 'init'
 for i in range(0, 1000):
 	all_updates.append(xs)
 	xs, lr	= update(xs, opt.func_name)
@@ -87,7 +91,7 @@ for i in range(0, 1000):
 all_updates.append(xs)
 
 all_updates	= np.array(all_updates)
-plt.plot(all_updates[:,0], all_updates[:,1], 'k.-')
+plt.plot(all_updates[:,0], all_updates[:,1], 'k.-', markersize=5)
 plt.title('Iteration ${0}$ Current Point: $({1}, {2})$'.format(i+1, round(xs[0], 5), round(xs[1], 5)), size=20)
 plt.show()
 

### 2017-08-30 Programming-Assignment-1/fixed_step_size.py

@@ -22,7 +22,12 @@ opt	= parser.parse_args()
 
 # Set initial point
 if opt.init_type == 'rand':
-	xs	= np.random.uniform(low=-6, high=6, size=2)
+	if opt.func_name == 'rosenbrock':
+		x	= np.random.uniform(low=-3, high=3)
+		y	= np.random.uniform(low=-6, high=6)
+		xs	= np.array([x, y])
+	else:
+		xs	= np.random.uniform(low=-6, high=6, size=2)
 
 elif opt.init_type == 'static':
 	xs	= np.array([2, 3])
@@ -68,6 +73,6 @@ for i in range(0, 1000):
 all_updates.append(xs)
 
 all_updates	= np.array(all_updates)
-plt.plot(all_updates[:,0], all_updates[:,1], 'k.-')
+plt.plot(all_updates[:,0], all_updates[:,1], 'k.-', markersize=5)
 plt.title('Iteration ${0}$ Current Point: $({1}, {2})$'.format(i+1, round(xs[0], 5), round(xs[1], 5)), size=20)
 plt.show()

### 2017-08-30 Programming-Assignment-1/inv_step_size.py

@@ -21,7 +21,12 @@ opt	= parser.parse_args()
 
 # Set initial point
 if opt.init_type == 'rand':
-	xs	= np.random.uniform(low=-6, high=6, size=2)
+	if opt.func_name == 'rosenbrock':
+		x	= np.random.uniform(low=-3, high=3)
+		y	= np.random.uniform(low=-6, high=6)
+		xs	= np.array([x, y])
+	else:
+		xs	= np.random.uniform(low=-6, high=6, size=2)
 
 elif opt.init_type == 'static':
 	xs	= np.array([2, 3])
@@ -67,6 +72,6 @@ for i in range(0, 1000):
 all_updates.append(xs)
 
 all_updates	= np.array(all_updates)
-plt.plot(all_updates[:,0], all_updates[:,1], 'k.-')
+plt.plot(all_updates[:,0], all_updates[:,1], 'k.-', markersize=5)
 plt.title('Iteration ${0}$ Current Point: $({2}, {3})$'.format(i+1, round(1/(i+1), 5), round(xs[0], 5), round(xs[1], 5)), size=20)
 plt.show()

### 2017-08-30 Programming-Assignment-1/PA-1_report.tex
PA-1 Update 7

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{float}

\title{Report for Programming Assignment 1}
\author{Vishwak Srinivasan\\
\texttt{CS15BTECH11043}}
\date{}

\begin{document}
\maketitle
\begin{flushleft}
The ``special" functions in consideration for this assignment were:
\begin{gather*}
f_{Q}(x, y) = 1.125x^2 + 0.5xy + 0.75y^2 + 2x + 2y \\
f_{LL}(x, y) = 0.5(x^2 + y^2) + 50\log\left(1 + e^{-0.5y}\right) + 50\log\left(1 + e^{0.2x}\right) \\
f_{H}(x, y) = 0.1(x^2 + y - 11)^2 + 0.1(x + y^2 - 7)^2 \\
f_{R}(x, y) = 0.002(1 - x)^2 + 0.2(y - x^2)^2
\end{gather*}

These functions were plotted using \texttt{GNUplot}, a terminal based plotting tool. The code for plotting can be found in the file: \texttt{plotter.gnuplot}. A few important observations could be made:
\begin{itemize}
\item In the domain of consideration i.e., \([-6, 6]\times[-6, 6]\) for \(f_{Q}, f_{LL} \text{ \& } f_{H}\) and \([-3, 3] \times [-6, 6]\) for \(f_{R}\) --- \(f_{Q}\) and \(f_{LL}\) were found to convex, and \(f_{H}\) and \(f_{R}\) were found to be non-convex. All these inferences were based of the surfaces plotted.
\item \(f_{Q}\) and \(f_{LL}\) have 1 minimum each, whereas \(f_{H}\) has 4 minima and 1 local maximum. The problem with \(f_{R}\) is the existence of a valley like structure.
\end{itemize}

\begin{figure}[H]
\begin{minipage}{0.45\linewidth}
\centering
\includegraphics[width=0.75\textwidth]{quadratic_function}
\caption{Note the basin (dark region in the middle)}
\end{minipage}
\hfill
\begin{minipage}{0.45\linewidth}
\centering
\includegraphics[width=0.75\textwidth]{ridge_regularized_logistic_regression}
\caption{Note the basin (slightly darker region near \((-4, 6)\))}
\end{minipage}
\end{figure}

\begin{figure}[H]
\begin{minipage}{0.45\linewidth}
\centering
\includegraphics[width=0.75\textwidth]{himmelblaus_function}
\caption{Note the 4 basins and a plateau approximately in the middle of them}
\end{minipage}
\hfill
\begin{minipage}{0.45\linewidth}
\centering
\includegraphics[width=0.75\textwidth]{rosenbrock_banana_function}
\caption{Note the valley (dark parabolic structure)}
\end{minipage}
\end{figure}

Gradient Descent was the algorithm used to find attain the closest minimum from the point of initialization. The reason \textit{closest minimum} is explicitly mentioned is because of the existence of multiple minima in the Himmelblaus function, and the dynamics of Gradient Descent dictate that the result of the algorithm will lie at the closest basin of attraction.
\(\newline\)

The first Gradient Descent method used was Vanilla Gradient Descent with a fixed step size to be chosen from \(\lbrace 0.3, 0.1, 0.01 \rbrace\). The point of initializations were one fixed point -- \((2,3)\) for all functions and two other points chosen randomly from the domain in consideration of the functions. My implementation chooses these points from a uniform distribution over the domain of the functions.

\end{flushleft}
\end{document}

### 2017-08-30 Programming-Assignment-1/PA-1_report.tex
PA-1 Update 8

@@
 All these inferences were based of the surfaces plotted.
-\item \(f_{Q}\) and \(f_{LL}\) have 1 minimum each, whereas \(f_{H}\) has 4 minima and 1 local maximum. The problem with \(f_{R}\) is the existence of a valley like structure.
+\item \(f_{Q}\) and \(f_{LL}\) have 1 minimum each, whereas \(f_{H}\) has 4 local minima and 1 local maximum. The problem with \(f_{R}\) is the existence of a valley like structure.
 \end{itemize}
@@
 \includegraphics[width=0.75\textwidth]{ridge_regularized_logistic_regression}
-\caption{Note the basin (slightly darker region near \((-4, 6)\))}
+\caption{Note the basin (slightly darker region near \((-4, 4)\))}
 \end{minipage}
@@
 My implementation chooses these points from a uniform distribution over the domain of the functions.
+\(\newline\)
+
+We are going to evaluate the performance of different learning rate settings for the static initialization at \((2,3)\) by considering the different in the function values at the optimal and last update i.e., \(1000^{th}\) iteration.
+The optimal values calculated analytically (or using the Newton-Raphson method for \(f_{LL}(x, y)\)) are as follows:
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+Function & Optimal Value in domain & Attained at \\
+\hline
+Quadratic Function & \(0.0\) & \((-0.64, -1.12)\)\\
+\hline
+Ridge Regularized Logistic Regression & \(\approx 40.4012\) & \((-3.3742, 3.5787)\)\\
+\hline 
+Himmelblaus Function & \(0.0\) & \((3, 2)\)\\
+\hline
+Rosenbrock's Banana Function & \(0.0\) & \((1, 1)\)\\
+\hline
+\end{tabular}
+\end{center}
+
+\begin{center}
+\begin{tabular}{|c|c|c|c|}
+\hline
+& 0.3 & 0.1 & 0.01 \\
+\hline
+\(f_{Q}(x, y)\) & & & \\
+\hline
+\(f_{LL}(x, y)\) & & & \\
+\hline
+\(f_{H}(x, y)\) & & & \\
+\hline
+\(f_{R}(x, y)\) & & & \\
+\hline
+\end{tabular}
+\end{center}
 

### 2017-08-30 Programming-Assignment-1/backtracking_ls.py

@@ -60,7 +60,7 @@ if opt.func_name == 'quad':
 	Z	= quad(X, Y)
 
 elif opt.func_name == 'log_reg':
-	levels	= np.linspace(0, 265, 212)
+	levels	= np.linspace(40, 265, 180)
 	xmin, xmax, xstep	= -6, 6, 0.05
 	ymin, ymax, ystep	= -6, 6, 0.05
 	X, Y	= np.meshgrid(np.arange(xmin, xmax + xstep, xstep), np.arange(ymin, ymax + ystep, ystep))

### 2017-08-30 Programming-Assignment-1/fixed_step_size.py

@@ -44,7 +44,7 @@ if opt.func_name == 'quad':
 	Z	= quad(X, Y)
 
 elif opt.func_name == 'log_reg':
-	levels	= np.linspace(0, 265, 212)
+	levels	= np.linspace(40, 265, 180)
 	xmin, xmax, xstep	= -6, 6, 0.05
 	ymin, ymax, ystep	= -6, 6, 0.05
 	X, Y	= np.meshgrid(np.arange(xmin, xmax + xstep, xstep), np.arange(ymin, ymax + ystep, ystep))

### 2017-08-30 Programming-Assignment-1/inv_step_size.py

@@ -43,7 +43,7 @@ if opt.func_name == 'quad':
 	Z	= quad(X, Y)
 
 elif opt.func_name == 'log_reg':
-	levels	= np.linspace(0, 265, 212)
+	levels	= np.linspace(40, 265, 180)
 	xmin, xmax, xstep	= -6, 6, 0.05
 	ymin, ymax, ystep	= -6, 6, 0.05
 	X, Y	= np.meshgrid(np.arange(xmin, xmax + xstep, xstep), np.arange(ymin, ymax + ystep, ystep))

### 2017-08-30 Programming-Assignment-1/newton_raphson.py

@@ -1,0 +1,23 @@
+import numpy as np
+
+def f(z, a):
+	return z + 50*a*np.exp(a*z)/(1 + np.exp(a*z))
+	
+def df(z, a):
+	return 1 + 50*a*(a*np.exp(a*z)/(1 + np.exp(a*z))**2)
+	
+def fxy(x, y):
+	return 0.5*(x**2 + y**2) + 50*(np.log(1 + np.exp(-0.5*y))) + 50*(np.log(1 + np.exp(0.2*x)))
+
+x	= -1
+for i in range(0, 50):
+	x	= x - f(x, 0.2)/df(x, 0.2)
+	
+y	= 1
+for i in range(0, 50):
+	y	= y - f(y, -0.5)/df(y, -0.5)
+	
+print('{0} at {1}'.format(f(x, 0.2), x))
+print('{0} at {1}'.format(f(y, -0.5), y))
+
+print(fxy(x, y))

### 2017-08-31 Programming-Assignment-1/PA-1_report.tex
PA-1 Update 9

@@
 \centering
-\includegraphics[width=0.75\textwidth]{quadratic_function}
+\includegraphics[width=0.75\textwidth]{./images/quadratic_function}
 \caption{Note the basin (dark region in the middle)}
@@
 \centering
-\includegraphics[width=0.75\textwidth]{ridge_regularized_logistic_regression}
+\includegraphics[width=0.75\textwidth]{./images/ridge_regularized_logistic_regression}
 \caption{Note the basin (slightly darker region near \((-4, 4)\))}
@@
 \centering
-\includegraphics[width=0.75\textwidth]{himmelblaus_function}
+\includegraphics[width=0.75\textwidth]{./images/himmelblaus_function}
 \caption{Note the 4 basins and a plateau approximately in the middle of them}
@@
 \centering
-\includegraphics[width=0.75\textwidth]{rosenbrock_banana_function}
+\includegraphics[width=0.75\textwidth]{./images/rosenbrock_banana_function}
 \caption{Note the valley (dark parabolic structure)}
@@
 The reason \textit{closest minimum} is explicitly mentioned is because of the existence of multiple minima in the Himmelblaus function, and the dynamics of Gradient Descent dictate that the result of the algorithm will lie at the closest basin of attraction.
-\(\newline\)
-
-The first Gradient Descent method used was Vanilla Gradient Descent with a fixed step size to be chosen from \(\lbrace 0.3, 0.1, 0.01 \rbrace\).
-The point of initializations were one fixed point -- \((2,3)\) for all functions and two other points chosen randomly from the domain in consideration of the functions.
-My implementation chooses these points from a uniform distribution over the domain of the functions.
 \(\newline\)
@@
 
+The first Gradient Descent method used was Vanilla Gradient Descent with a fixed step size to be chosen from \(\lbrace 0.3, 0.1, 0.01 \rbrace\).
+The point of initializations were one fixed point -- \((2,3)\) for all functions and two other points chosen randomly from the domain in consideration of the functions.
+My implementation chooses these points from a uniform distribution over the domain of the functions.
+The below table shows the difference \(f(\mathbf{x}^{(1000)}) - f(\mathbf{x}^{*})\) for initialization at \((2,3)\).
+\(\newline\)
+
 \begin{center}
@@
 \hline
-\(f_{Q}(x, y)\) & & & \\
+\(f_{Q}(x, y)\) & 0 & 0 & 0\\
 \hline
-\(f_{LL}(x, y)\) & & & \\
+\(f_{LL}(x, y)\) & 0 & 0 & 0\\
 \hline
-\(f_{H}(x, y)\) & & & \\
+\(f_{H}(x, y)\) & 1.769 & 0 & 0 \\
 \hline
-\(f_{R}(x, y)\) & & & \\
+\(f_{R}(x, y)\) & 0.00092 & 0.00106 & 0.00112 \\
 \hline
@@
 
+From the above table it can be seen that for any setting of learning rate from the given values, the algorithm converges to the minimum for the convex functions: \(f_{Q}\) and \(f_{LL}\).
+This is natural from what we have learnt that gradient descent will eventually converge to the minimum for convex functions.
+In \(f_{H}\), convergence was seen for the step size being set to \(0.1\) and \(0.01\), but there was an oscillation observed in \(0.3\), hence the point got pushed away from the minimum.
+In the case of \(f_{R}\), the valley like structure makes the gradient updates slower and slower and hence none of the settings make it to the minimum.
+\(\newline\)
+
+The next Gradient Descent method used was Backtracking Line Search with Gradient Descent. Here we consider the approximation:
+\begin{equation*}
+f(\mathbf{x} + t\nabla f(\mathbf{x})) \approx f(\mathbf{x}) - t\alpha ||\nabla f(\mathbf{x})||^{2}_{2}
+\end{equation*}
+
+The update step for \(t\) is \(t := \beta t\).
+\(\alpha = 0.5\) and \(\beta = 0.5\) is our settings.
+Below are the value of \(f(\mathbf{x}^{(1000)}) - f(\mathbf{x}^{*})\) for initialization at \((2,3)\).
+\(\newline\)
+
+\begin{center}
+\begin{tabular}{|c|c|c|c|}
+\hline
+\(f_{Q}(x, y)\) & \(f_{LL}(x, y)\) & \(f_{H}(x, y)\) & \(f_{R}(x, y)\) \\
+\hline
+0 & 0 & 0 & 0.00078 \\
+\hline
+\end{tabular}
+\end{center}
+
+This shows the effectiveness of backtracking line search based Gradient Descent.
+It still could not navigate to the minimum in \(f_{R}(x, y)\), that is the nature of the valley.
+As learnt in class, we can see the effectiveness of backtracking line search here too.
+\(\newline\)
+
+The last Gradient Descent method used was an annealing schedule for the learning rate.
+We set the learning rate to \(\frac{1}{k}\) where \(k\) is the iteration number.
+Below are the values of \(f(\mathbf{x}^{(1000)}) - f(\mathbf{x}^{*})\) for initialization at \((2, 3)\).
+
+\begin{center}
+\begin{tabular}{|c|c|c|c|}
+\hline
+\(f_{Q}(x, y)\) & \(f_{LL}(x, y)\) & \(f_{H}(x, y)\) & \(f_{R}(x, y)\) \\
+\hline
+0 & 0 & 0 & 0.00062 \\
+\hline
+\end{tabular}
+\end{center}
+
+The paths are all affected by the initial high learning rate.
+In the case of \(f_{Q}\), we jump to the other side of the basin in the first step and then jump back again, and then slowly begin to converge.
+In the case of \(f_{LL}\), we jump to the other side of the minimum and then to and fro about the minimum and then converge.
+The most interesting case was in \(f_{H}\).
+Due to the high learning rate, we end near another local minimum close to \((-3.779, -3.283)\) and infact do not converge at the closest local minimum to the point of initialization, which is \((3, 2)\).
+In \(f_{R}\), we make a relatively large jump, but still get stuck in the valley.
+From what we have learnt in class, setting large learning rate initially might not help if our initialization is close to any minimum/basin of attraction.
+Below are the graphs:
+
+\begin{figure}[H]
+\begin{minipage}{0.45\linewidth}
+\includegraphics[width=0.9\textwidth]{./images/quad_inv}
+\caption{\(f_{Q}\)}
+\end{minipage}
+\hfill
+\begin{minipage}{0.45\linewidth}
+\includegraphics[width=0.9\textwidth]{./images/log_reg_inv}
+\caption{\(f_{LL}\)}
+\end{minipage}
+\end{figure}
+
+\begin{figure}[H]
+\begin{minipage}{0.45\linewidth}
+\includegraphics[width=0.9\textwidth]{./images/himmelblau_inv}
+\caption{\(f_{H}\)}
+\end{minipage}
+\hfill
+\begin{minipage}{0.45\linewidth}
+\includegraphics[width=0.9\textwidth]{./images/rosenbrock_inv}
+\caption{\(f_{R}\)}
+\end{minipage}
+\end{figure}
+
 \end{flushleft}

### 2017-08-31 Programming-Assignment-1/backtracking_ls.py

@@ -90,6 +90,15 @@ for i in range(0, 1000):
 	all_bcktrck.append(np.array([i+1, -np.log2(lr)]))
 all_updates.append(xs)
 
+if opt.func_name == 'quad':
+	print(quad(*xs))
+elif opt.func_name == 'log_reg':
+	print(log_reg(*xs))
+elif opt.func_name == 'himmelblaus':
+	print(himmelblaus(*xs))
+elif opt.func_name == 'rosenbrock':
+	print(rosenbrock(*xs))
+
 all_updates	= np.array(all_updates)
 plt.plot(all_updates[:,0], all_updates[:,1], 'k.-', markersize=5)
 plt.title('Iteration ${0}$ Current Point: $({1}, {2})$'.format(i+1, round(xs[0], 5), round(xs[1], 5)), size=20)

### 2017-08-31 Programming-Assignment-1/fixed_step_size.py

@@ -72,6 +72,15 @@ for i in range(0, 1000):
 	xs	= update(xs, opt.lr, opt.func_name)
 all_updates.append(xs)
 
+if opt.func_name == 'quad':
+	print(quad(*xs))
+elif opt.func_name == 'log_reg':
+	print(log_reg(*xs))
+elif opt.func_name == 'himmelblaus':
+	print(himmelblaus(*xs))
+elif opt.func_name == 'rosenbrock':
+	print(rosenbrock(*xs))
+
 all_updates	= np.array(all_updates)
 plt.plot(all_updates[:,0], all_updates[:,1], 'k.-', markersize=5)
 plt.title('Iteration ${0}$ Current Point: $({1}, {2})$'.format(i+1, round(xs[0], 5), round(xs[1], 5)), size=20)

### 2017-08-31 Programming-Assignment-1/inv_step_size.py

@@ -71,6 +71,15 @@ for i in range(0, 1000):
 	xs	= update(xs, 1/(i+1), opt.func_name)
 all_updates.append(xs)
 
+if opt.func_name == 'quad':
+	print(quad(*xs))
+elif opt.func_name == 'log_reg':
+	print(log_reg(*xs))
+elif opt.func_name == 'himmelblaus':
+	print(himmelblaus(*xs))
+elif opt.func_name == 'rosenbrock':
+	print(rosenbrock(*xs))
+
 all_updates	= np.array(all_updates)
 plt.plot(all_updates[:,0], all_updates[:,1], 'k.-', markersize=5)
 plt.title('Iteration ${0}$ Current Point: $({2}, {3})$'.format(i+1, round(1/(i+1), 5), round(xs[0], 5), round(xs[1], 5)), size=20)

### 2017-08-31 Programming-Assignment-1/newton_raphson.py

@@ -1,23 +1,25 @@
+# Newton-Raphson to find the roots of equation
+# g_a(x) = x + 50*a*exp(ax)/(1 + exp(ax))
+# This happens to be the form of the partial derivatives of the Ridge Regularized Logistic Regression function
 import numpy as np
 
-def f(z, a):
+def g(z, a):
 	return z + 50*a*np.exp(a*z)/(1 + np.exp(a*z))
-	
-def df(z, a):
+
+def dg(z, a):
 	return 1 + 50*a*(a*np.exp(a*z)/(1 + np.exp(a*z))**2)
-	
-def fxy(x, y):
+
+def f_ll(x, y):
 	return 0.5*(x**2 + y**2) + 50*(np.log(1 + np.exp(-0.5*y))) + 50*(np.log(1 + np.exp(0.2*x)))
 
 x	= -1
 for i in range(0, 50):
-	x	= x - f(x, 0.2)/df(x, 0.2)
-	
+	x	= x - g(x, 0.2)/dg(x, 0.2)	
+
 y	= 1
 for i in range(0, 50):
-	y	= y - f(y, -0.5)/df(y, -0.5)
+	y	= y - g(y, -0.5)/dg(y, -0.5)
 	
-print('{0} at {1}'.format(f(x, 0.2), x))
-print('{0} at {1}'.format(f(y, -0.5), y))
-
-print(fxy(x, y))
+print('Partial Derivative w.r.t x => {0} at {1} \
+	\nPartial Derivative w.r.t y => {2} at {3} \
+	\nFunction Value is: {4}'.format(round(g(x, 0.2), 10), round(x, 4), round(g(y, -0.5), 10), round(y, 4), round(f_ll(x, y), 4)))

### 2017-08-31 Programming-Assignment-1/PA-1_report.tex
PA-1 Update 10

@@
 In \(f_{H}\), convergence was seen for the step size being set to \(0.1\) and \(0.01\), but there was an oscillation observed in \(0.3\), hence the point got pushed away from the minimum.
+
+\begin{figure}[H]
+\begin{minipage}{0.32\linewidth}
+\includegraphics[width=0.9\textwidth]{./images/himmelblau_0_01.eps}
+\caption{\(f_{H}(x, y) \rightarrow \eta\) = 0.01}
+\end{minipage}
+\hfill
+\begin{minipage}{0.32\linewidth}
+\includegraphics[width=0.9\textwidth]{./images/himmelblau_0_1.eps}
+\caption{\(f_{H}(x, y) \rightarrow \eta\) = 0.1}
+\end{minipage}
+\hfill
+\begin{minipage}{0.32\linewidth}
+\includegraphics[width=0.9\textwidth]{./images/himmelblau_0_3.eps}
+\caption{\(f_{H}(x, y) \rightarrow \eta\) = 0.3}
+\end{minipage}
+\end{figure}
 In the case of \(f_{R}\), the valley like structure makes the gradient updates slower and slower and hence none of the settings make it to the minimum.
@@
 \end{equation*}
+\(\newline\)
 

### 2017-08-31 Programming-Assignment-1/PA-1_report.tex
PA-1 Update 11

@@
 \documentclass{article}
-\usepackage[margin=1in]{geometry}
+\usepackage[margin=0.73in]{geometry}
 \usepackage{amsmath}
@@
 \usepackage{float}
+\usepackage{hyperref}
 
@@
 \maketitle
+
+\section*{Instructions for examining the programs}
+\begin{itemize}
+\item There are three important programs: \texttt{fixed\_step\_size.py} for fixed step size (Question 2), \texttt{backtracking\_ls.py} for backtracing line search (Question 3) and \texttt{inv\_step\_size.py} for learning rate = \(\frac{1}{k}\) (Question 4).
+\item All of the above mentioned files depend on \texttt{NumPy} and a module \texttt{functions.py}, included with these programs.
+\item To test a program, run \texttt{python3 <name\_of\_the\_program> -h}. The options are there for your convenience.
+\item Since analytically arriving at a solution for \(f_{LL}\) was not possible, I wrote a smaller program to find the roots of partial derivatives using the Newton-Raphson Method \texttt{newton\_raphson.py}.
+\item The plots were generated using \texttt{GNUplot}. The plots can be obtained by running \texttt{gnuplot plotter.gnuplot}.
+\end{itemize}
+
+\section*{Main Report}
 \begin{flushleft}
@@
 
-These functions were plotted using \texttt{GNUplot}, a terminal based plotting tool.
-The code for plotting can be found in the file: \texttt{plotter.gnuplot}.
-A few important observations could be made:
 \begin{itemize}
@@
 \centering
-\includegraphics[width=0.75\textwidth]{./images/quadratic_function}
+\includegraphics[width=0.9\textwidth]{./images/quadratic_function}
 \caption{Note the basin (dark region in the middle)}
@@
 \centering
-\includegraphics[width=0.75\textwidth]{./images/ridge_regularized_logistic_regression}
+\includegraphics[width=0.9\textwidth]{./images/ridge_regularized_logistic_regression}
 \caption{Note the basin (slightly darker region near \((-4, 4)\))}
@@
 \centering
-\includegraphics[width=0.75\textwidth]{./images/himmelblaus_function}
+\includegraphics[width=0.9\textwidth]{./images/himmelblaus_function}
 \caption{Note the 4 basins and a plateau approximately in the middle of them}
@@
 \centering
-\includegraphics[width=0.75\textwidth]{./images/rosenbrock_banana_function}
+\includegraphics[width=0.9\textwidth]{./images/rosenbrock_banana_function}
 \caption{Note the valley (dark parabolic structure)}
@@
 
-\begin{center}
+\begin{table}[H]
+\centering
 \begin{tabular}{|c|c|c|c|}
 \hline
-& 0.3 & 0.1 & 0.01 \\
+\(\eta \rightarrow\) & 0.3 & 0.1 & 0.01 \\
 \hline
@@
 \end{tabular}
-\end{center}
+\caption{\(f(\mathbf{x}^{(1000)}) - f(\mathbf{x}^{*})\) for Fixed Step Size}
+\end{table}
 
@@
 
-\begin{center}
+\begin{table}[H]
+\centering
 \begin{tabular}{|c|c|c|c|}
@@
 \end{tabular}
-\end{center}
+\caption{\(f(\mathbf{x}^{(1000)}) - f(\mathbf{x}^{*})\) for Backtracking Line Search}
+\end{table}
 
@@
 \end{tabular}
-\end{center}
+\caption{\(f(\mathbf{x}^{(1000)}) - f(\mathbf{x}^{*})\) for Annealing Schedule}
+\end{table}
 
@@
 
+From Yann LeCun's paper \href{http://yann.lecun.com/exdb/publis/pdf/lecun-simard-pearlmutter-93.pdf}{Automatic Learning Rate Maximization by On-Line Estimation of the Hessian's Eigenvectors}, the optimal learning rate seems to be \(\eta_{\text{opt}} = \frac{1}{\lambda_{max}}\), where \(\lambda_{max}\) is the maximum eigenvalue of the Hessian matrix.
+If the learning rate is too high, then it might be the case that the learning rate set is too greater than the optimal learning rate itself.
+For example: in the case of Himmelblau's function \(f_{H}(x, y)\), the eigenvalues of the Hessian at \((2, 3)\) are : \(1.338, 10.462\), which means that the optimal learning rate is: \(\approx 0.095\) to start with.
+
 \end{flushleft}

### 2017-10-03 Assignment-2/Solns-Assgn-2.tex
Assgn-2 Update 1

\documentclass[10.5pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bookmark}

\title{Solutions to the Assignment - 2 : CS6230 - \\
Optimization Methods in Machine Learning}
\author{Vishwak Srinivasan\\
\texttt{CS15BTECH11043}}
\date{}

\begin{document}
\maketitle

\section*{Question 2}
\begin{flushleft}
To prove convexity, consider 2 points: \(x\) and \(y\).
\subsubsection*{Case 1}
Let \(x = 0\) and \(y = 0\). Because of this, any convex combination of \(x\) and \(y\) has to be \(0\). Formally, \(\forall \hspace{1.5mm} \theta \in [0, 1]\),
\begin{gather*}
f(\theta x + (1 - \theta) y) = f(0) = 1 \\
\theta f(x) + (1 - \theta) f(y) = \theta + 1 - \theta = 1
\end{gather*}
Since \(f(\theta x + (1 - \theta) y) = \theta f(x) + (1 - \theta) f(y)\), convexity holds.

\subsubsection*{Case 2}
Without loss of generality, let \(x > 0\) and \(y = 0\). \(\forall \hspace{1.5mm} \theta \in [0, 1]\),
\begin{gather*}
f(\theta x + (1 - \theta)y) = f(\theta x) = \begin{cases} 1 & \text{if} \theta = 0 \\ 0 & \text{otherwise} \end{cases} \\
\theta f(x) + (1 - \theta) f(y) = (1 - \theta)
\end{gather*}
If \(\theta = 0\), then \[f(\theta x + (1 - \theta)y) = \theta f(x) + (1 - \theta) f(y)\] Else, \[f(\theta x + (1 - \theta)y) = 0 \leq \theta f(x) + (1 - \theta) f(y) = (1  - \theta)\] Convexity holds.

\subsubsection*{Case 3}
Let \(x > 0\) and \(y > 0\). Because of this, any convex combination of \(x\) and \(y\) has to be \(> 0\). Formally, \(\forall \hspace{1.5mm} \theta \in [0, 1]\),
\begin{gather*}
f(\theta x + (1 - \theta)y) = 0 \\
\theta f(x) + (1 - \theta) f(y) = 0
\end{gather*}
Since \(f(\theta x + (1 - \theta) y) = \theta f(x) + (1 - \theta) f(y)\), convexity holds.

Since in all cases, convexity held, the function is verified to be convex.

Consider the task of finding the subgradients at \(x = 0\). By the definition of subgradients:
\begin{gather*}
f(y) \geq f(x) + g^{T}(y - x) \\
f(y) \geq 1 + gy
\end{gather*}

Now if \(y = 0\), then \(g\) could be any number. Otherwise if \(y > 0\), then we get this condition: \(g \leq \frac{-1}{y}\). Note that \(g\) should hold for all \(y\). Hence for the smallest \(y\) close to \(0\), then set of subgradients may be an empty set, hence leading the intersection to be an empty set. Since the set of subgradients is empty, the subdifferentials at \(x = 0\) are empty, leading to conclude that \(f\) is not subdifferentiable at \(x = 0\).
\end{flushleft}

\section*{Question 4}
\subsection*{Part a}
\begin{flushleft}
We know that if \(\nabla g\) is L-Lipschitz, then:
\begin{equation}
\label{grad-lips}
\nabla^2 g(x) \preceq LI
\end{equation}
where \(I\) is the identity matrix of corresponding dimensions.
\(\newline\)

Hence by using a \(2^{nd}\) order Taylor series expansion of \(g\) around a given \(x\), we get:
\begin{equation}
\label{second-taylor}
\displaystyle g(y) \approx g(x) + \nabla g(x)^{T}(y - x) + \frac{(y - x)^{T}\nabla^2 g(x) (y - x)}{2}
\end{equation}
Note that we have assumed that \(g\) is twice-differentiable. The proof breaks down if this is not taken into consideration.
\(\newline\)

Using \ref{grad-lips} in \ref{second-taylor}, we get:
\begin{equation}
\label{grad-lips-taylor}
\displaystyle g(y) \leq g(x) + \nabla g(x)^{T}(y - x) + \frac{L}{2}||y - x||_{2}^{2}
\end{equation}

Since \(f(y) = g(y) + h(y)\), substituting \ref{grad-lips-taylor}:
\begin{equation}
\label{part-a}
f(y) \leq g(x) + \nabla g(x)^{T}(y - x) + \frac{L}{2}||y - x||_{2}^{2} + h(y)
\end{equation}
\end{flushleft}

\subsection*{Part b}
\begin{flushleft}
Let \(y = x^{+} = x - tG_{t}(x)\). Substituting this in \ref{part-a}, we get:
\begin{equation}
\label{update-subs}
f(x^{+}) \leq g(x) + \nabla g(x)^{T}(-tG_{t}(x)) + \frac{Lt^{2}}{2}||G_{t}(x)||_{2}^{2} + h\left(x - tG_{t}(x)\right)
\end{equation}

Since \(t \leq \frac{1}{L}\), \(L \leq \frac{1}{t}\) and consequently \(\frac{Lt^2}{2} \leq \frac{t}{2}\). Substitute this is \ref{update-subs} to get:
\begin{equation}
\label{part-b}
f(x^{+}) \leq g(x) - t\nabla g(x)^TG_{t}(x) + \frac{t}{2}||G_{t}(x)||_{2}^{2} + h\left(x - tG_{t}(x)\right)
\end{equation}
\end{flushleft}

\subsection*{Part c}
\begin{flushleft}
We know that if \(z = \textbf{prox}_{h, t}(x)\), then:
\begin{equation}
\label{subgrad-optim}
x - z \in \partial h(z) \implies x - \textbf{prox}_{h, t}(x) \in \partial h(\textbf{prox}_{h, t}(x))
\end{equation}

Substitute \(x \leftarrow x - t\nabla g(x)\) in \ref{subgrad-optim} to get:
\begin{equation}
\label{update-subgrad-optim}
x - t\nabla g(x) - \textbf{prox}_{h, t}(x - t\nabla g(x)) \in \partial h(\textbf{prox}_{h, t}(x - t\nabla g(x)))
\end{equation}

Using the definition of \(\textbf{prox}_{h, t}(x - t\nabla g(x)) = x - tG_{t}(x)\) in \ref{update-subgrad-optim}, we get:
\begin{equation}
\label{part-c}
x - t\nabla g(x) - x + tG_{t}(x) \in \partial h(x - tG_{t}(x)) \implies tG_{t}(x) - tg(x) \in \partial h(x - tG_{t}(x))
% @TODO: Need to check this equation
\end{equation}
\end{flushleft}

\subsection*{Part d}
\begin{flushleft}
Note that the required inequation has a \(f(z)\) in the RHS, indicating the presence of \(g(z)\) and \(h(z)\) in our steps.
Using the definition of convexity of \(g\) and \(h\) we get:
\begin{equation}
\label{cvx-g}
g(z) \geq g(x) + \nabla g(x)^{T}(z - x) \implies g(x) \leq g(z) - \nabla g(x)^{T}(z - x)
\end{equation}
\begin{equation}
\label{cvx-h}
h(z) \geq h(x) + s(x)^{T}(z - x) \implies h(x) \leq h(z) - s(x)^{T}(z - x)
\end{equation}

Substituting \(x \leftarrow x - tG_{t}(x)\) in \ref{cvx-h} and using the result in \ref{part-c}, we get:
\begin{equation}
\label{update-cvx-h}
h(x - tG_{t}(x)) \leq h(z) - s(x - tG_{t}(x))^{T}(z - x) = h(z) - (G_{t}(x) - \nabla g(x))^{T}(z - x + tG_{t}(x))
\end{equation}

Using \ref{part-b} and the substituting the results from \ref{cvx-g} and \ref{update-cvx-h}:
\begin{equation}
\label{step-d1}
f(x^{+}) \leq g(z) - \nabla g(x)^{T}(z - x) - t\nabla g(x)^{T}G_{t}(x) + \frac{t}{2}||G_{t}(x)||_{2}^{2} + h(z) - (G_{t}(x) - \nabla g(x))^{T}(z - x + tG_{t}(x))
\end{equation}

Performing the rearrangement: \[(G_{t}(x) - \nabla g(x))^{T}(z - x + tG_{t}(x)) = (G_{t}(x) - \nabla g(x))^{T}(z - x) + t||G_{t}(x)||_{2}^{2} - t\nabla g(x)^{T}G_{t}(x)\] and substituting in \ref{step-d1} we proceed to get:
\begin{equation}
\label{step-d2}
f(x^{+}) \leq g(z) - (\nabla g(x) + G_{t}(x) - \nabla g(x))^{T}(z - x) - t\nabla g(x)^{T}G_{t}(x) + t\nabla g(x)^{T}G_{t}(x) - \frac{t}{2}||G_{t}(x)||_{2}^{2}
\end{equation}

Cancelling out the required terms in \ref{step-d2} and using the fact that \(f(z) = g(z) + h(z)\)
\begin{equation}
\label{part-d}
f(x^{+}) \leq f(z) - G_{t}(x)^{T}(z - x) - \frac{t}{2}||G_{t}(x)||_{2}^{2} = f(z) + G_{t}(x)^{T}(x - z) - \frac{t}{2}||G_{t}(x)||_{2}^{2}
\end{equation}

\subsection*{Part e}
We know that \(x^{+}\) is the updated after \(G_{t}(x)\) is applied to \(x\). Substituting \(z = x\) in \ref{part-d}, we can see that:
\begin{equation}
\label{guarantee}
f(x^{+}) \leq f(x) - \frac{t}{2}||G_{t}(x)||_{2}^{2}
\end{equation}

Since \(||.||\) is positive, \ref{guarantee} guarantees a decrease in the function value after a step is made. Use the update rule \(x^{+} = x - tG_{t}(x)\) to get \(G_{t}(x) = \frac{x - x^{+}}{t}\). Substituting: \(z = x^{*}\) and the equation for \(G_{t}(x)\) in \ref{part-d}, we get:
\begin{equation}
\label{step-e1}
f(x^{+}) \leq f(x^{*}) + \frac{1}{2t}\left((x - x^{+})^{T}(2x - 2x^{*}) - (x - x^{+})^{T}(x - x^{+})\right)
\end{equation}

Playing with the second term:
\begin{gather*}
2x - 2x^{*} - x + x^{+} = (x - x^{*}) - (x^{*} - x^{+}) \\
\implies (x - x^{+})^{T}(2x - 2x^{*} - (x - x^{+})) = (x - x^{+})^{T}(x - x^{*}) - (x - x^{+})^{T}(x^{*} - x^{+})
\end{gather*}

Splitting each \(x - x^{+}\) as \((x - x^{*}) + (x^{*} - x^{+})\):
\begin{gather*}
(x - x^{+})^{T}(x - x^{*}) = ||x - x^{*}||_{2}^{2} + (x^{*} - x^{+})^{T}(x - x^{*}) \\
(x - x^{+})^{T}(x^{*} - x^{+}) = (x - x^{*})^{T}(x^{*} - x^{+}) + ||x^{*} - x^{+}||_{2}^{2} \\
\implies (x - x^{+})^{T}(x - x^{*}) - (x - x^{+})^{T}(x^{*} - x^{+}) = ||x - x^{*}||_{2}^{2} - ||x^{+} - x^{*}||_{2}^{2} \\
\implies (x - x^{+})^{T}(2x - 2x^{*} - (x - x^{+})) = ||x - x^{*}||_{2}^{2} - ||x^{+} - x^{*}||_{2}^{2}
\end{gather*}

Substituting the final results back in \ref{step-e1}, we get:
\begin{equation}
\label{part-e}
f(x^{+}) \leq f(x^{*}) + \frac{1}{2t}\left(||x - x^{*}||_{2}^{2} - ||x^{+} - x^{*}||_{2}^{2}\right)
\end{equation}
\end{flushleft}

\subsection*{Part f}
\begin{flushleft}
We know from \ref{guarantee} that \(f(x^{(k)}) \leq f(x^{(k-1)})\). Substitute \(x^{+} \leftarrow x^{(k)}\) and \(x \leftarrow x^{(k-1)}\) in \ref{part-e}. We now get:
\begin{equation}
\label{step-f1}
f(x^{(k)}) - f(x^{*}) \leq \frac{1}{2t}\left(||x^{(k-1)} - x^{*}||_{2}^{2} - ||x^{(k)} - x^{*}||_2^{2}\right)
\end{equation}

Perform a summation over \(k\) running from 1 to \(K\) on both sides of \ref{step-f1}, which yields us:
\begin{equation}
\label{step-f2}
\displaystyle \sum_{k=1}^{K} f(x^{(k)}) - f(x^{*}) \leq \frac{1}{2t} \left(\sum_{k=1}^{K} ||x^{(k-1)} - x^{*}||_{2}^{2} - ||x^{(k)} - x^{*}||_2^{2}\right)
\end{equation}

The sum on the RHS of \ref{step-f2} is conveniently a telescoping sum.
\begin{equation}
\label{step-f3}
\displaystyle \sum_{k=1}^{K} f(x^{(k)}) - f(x^{*}) \leq \frac{1}{2t} \left(||x^{0} - x^{*}||_{2}^{2} - ||x^{K} - x^{*}||_{2}^{2}\right)
\end{equation}

Since \(f(x^{(K)}) \leq f(x^{(K-1)}) \leq \ldots \leq f(x^{(1)})\), we can lower-bound the value \(f(x^{(k)}) - f(x^{*})\) by \(f(x^{(K)}) - f(x^{*})\). Also \(||x^{(0)} - x^{*}||_{2}^{2} - ||x^{(K)} - x^{*}||_{2}^{2} \leq ||x^{(0)} - x^{*}||_{2}^{2}\). Using these substitutions in \ref{step-f3}, we get:
\begin{equation}
\label{part-f}
f(x^{(K)}) - f(x^{*}) \leq \frac{||x^{(0)} - x^{*}||_{2}^{2}}{2tK}
\end{equation}
\end{flushleft}
\end{document}

### 2017-10-06 Assignment-2/Solns-Assgn-2.tex
Assgn-2 Update 2

@@
-\documentclass[10.5pt]{article}
+\documentclass[11pt]{article}
 \usepackage[margin=1in]{geometry}
@@
 \maketitle
+
+\section*{Question 1}
+\subsection*{Part a}
+\begin{flushleft}
+\begin{equation}
+\label{ques-1a}
+\displaystyle f(x) = \max_{i=1\ldots m} (a_{i}^{T}x + b_{i})
+\end{equation}
+
+Note that \ref{ques-1a} is of the form: \(\displaystyle f(x) = \max_{i=1\ldots m} f_{i}(x)\). The steps involved in finding the subgradient of \(f\) at a given point are as follows:
+
+\begin{enumerate}
+\item See if \(f_{i}\) is subdifferentiable \(\forall i \in \{1 \ldots m\}\).
+\item If there is no intersection, then the subdifferential of \(f\) is the subdifferential of the unique maximum.
+\item If there is intersection, then the subdifferential of \(f\) at the point of intersection of is the set of all convex combinations of the subdifferentials of the function intersecting at that point.
+\end{enumerate}
+
+The first bullet point is simple to check: \(\nabla f_{i}(x) = a_{i}\).
+The second and third bullet point can be combined to form the following: the active functions at a given point are the set of functions such that \(f(x) = f_{\text{active}}(x)\).
+Note that in the second point, this set is singleton, whereas in the third point, this set maybe as big as \(m\).
+Hence the subdifferential of \(f\) is given by:
+\begin{equation}
+\label{ans-1a}
+\partial f = \text{Conv}\left(\{\partial f_{i}(x) | f_{i}(x) = f(x) \text{ and } i \in \{1 \ldots m\}\}\right)
+\end{equation}
+
+where the \(\text{Conv}(S)\) in \ref{ans-1a} denotes the set of all convex combinations of the elements of \(S\)/convex hull of \(S\).
+\end{flushleft}
+
+\subsection*{Part b}
+\begin{flushleft}
+\begin{equation}
+\label{ques-1b}
+\displaystyle f(x) = \max_{i=1\ldots m} |a_{i}^{T}x + b_{i}|
+\end{equation}
+
+Note that \ref{ques-1b} takes the same form as \ref{ques-1a}. In fact, for \ref{ques-1b}, one can apply the results of \ref{ques-1a} twice. 
+The steps are as follows:
+\begin{enumerate}
+\item See if \(f_{i}\) is subdifferentiable \(\forall i \in \{1 \ldots m\}\).
+\item Get the subdifferential using the convex combination of the subdifferentials of every active function.
+\end{enumerate}
+
+For the first step, we know that \(\displaystyle f_{i} = \max \{-a_{i}^{T}x - b_{i}, a_{i}^Tx + b_{i}\}\).
+From the results of Question 1a and also the fact that there are more than one active functions at \(x = \mathbf{0}\), we can tell that:
+\begin{equation}
+\label{sub-ans-1b}
+\partial f_{i}(0) = \text{Conv}(-a_{i}, a_{i}) = \{a_{i} - 2\theta a_{i} | \theta \in [0, 1]\}
+\end{equation}
+
+Using \ref{sub-ans-1b}, we replicate the same statement from Question 1a, which is:
+\begin{equation}
+\label{ans-1b}
+\partial f = \text{Conv}\left(\{\partial f_{i}(x) | f_{i}(x) = f(x) \text{ and } i \in \{1 \ldots m\}\}\right)
+\end{equation}
+
+where the \(\text{Conv}(S)\) in \ref{sub-ans-1b} and \ref{ans-1b} denotes the set of all convex combinations of the elements of \(S\)/convex hull of \(S\).
+
+\end{flushleft}
 

### 2017-10-10 Assignment-2/Solns-Assgn-2.tex
Assgn-2 Update 3

@@
 \usepackage{bookmark}
+\usepackage{hyperref}
 
@@
 
-Using \ref{sub-ans-1b}, we replicate the same statement from Question 1a, which is:
+Using \ref{sub-ans-1b}, we replicate the same statement from \ref{ans-1a}, which is:
 \begin{equation}
@@
 where the \(\text{Conv}(S)\) in \ref{sub-ans-1b} and \ref{ans-1b} denotes the set of all convex combinations of the elements of \(S\)/convex hull of \(S\).
+\end{flushleft}
 
+\subsection*{Part c}
+\begin{flushleft}
+\begin{equation}
+\label{ques-1c}
+\displaystyle f(x) = \sup_{0 \leq t \leq 1} x_{1} + x_{2}t + \ldots + x_{n}t^{n-1}
+\end{equation}
+
+Since the set \([0, 1]\) over which the supremum is being taken over in \ref{ques-1c} is closed and bounded, meaning it is compact, following from the lecture notes\footnote{\href{https://see.stanford.edu/materials/lsocoee364b/01-subgradients_notes.pdf}{Boyd and Vanderberghe Lecture Notes on Subgradients}}, we can apply the same intuition as done for \ref{ques-1a} and \ref{ques-1b}.
+In this case, it is required that \(p(t)\) is sub-differentiable w.r.t.
+\(x\).
+This is assured since the gradient is:
+\begin{equation}
+\label{sub-ans-1c}
+\displaystyle \nabla_{x} p(t) =  \langle 1, t, \ldots, t^{n-1} \rangle
+\end{equation}
+
+Using \ref{sub-ans-1c}, we use the same statement as in \ref{ans-1a} and \ref{ans-1b},
+\begin{equation}
+\label{ans-1c}
+\displaystyle \partial f = \text{Conv}\left(\{\nabla_{x} f(x) | f(x) = \sup p(t), t \in [0, 1]\}\right)
+\end{equation}
+
+where the \(\text{Conv}(S)\) if \ref{ans-1c} denotes the set of all convex combinations of the elements of \(S\)/convex hull of \(S\).
 \end{flushleft}
+
+\subsection*{Part d}
+\begin{flushleft}
+\begin{equation}
+\label{ques-1d}
+\displaystyle f(x) = x_{[1]} + x_{[2]} + \ldots + x_{[k]}
+\end{equation}
+
+Clearly, those indices not involved in the first \(k\) highest values will have their partial derivatives to be \(0\) and those indices involved in the first \(k\) highest values will have their partial derivatives to be \(1\).
+\(\newline\)
+
+To get these indices, we need to maintain a copy of the vector and sort the elements of that copy.
+Note the indices involved in the top \(k\) values.
+Call the set to be \(S\).
+Hence the subgradient for \ref{ques-1d} would be:
+\begin{equation}
+\label{ans-1d}
+\displaystyle g \text{ such that } g_{i} = \begin{cases} 1 & \text{ if } i \in S \\ 0 & \text{ otherwise} \end{cases}
+\end{equation}
+
+Note that \ref{ans-1d} will vary based on the value of \(k\) too. 
+\end{flushleft}
+\newpage
 
@@
 \begin{equation}
+\label{sub-part-c}
+x - t\nabla g(x) - x + tG_{t}(x) \in \partial h(x - tG_{t}(x)) \implies tG_{t}(x) - t\nabla g(x) \in \partial h(x - tG_{t}(x)) 
+\end{equation}
+
+Using \ref{sub-part-c}, we have:
+\begin{equation}
 \label{part-c}
-x - t\nabla g(x) - x + tG_{t}(x) \in \partial h(x - tG_{t}(x)) \implies tG_{t}(x) - tg(x) \in \partial h(x - tG_{t}(x))
-% @TODO: Need to check this equation
+\implies G_{t}(x) - \nabla g(x) \in \partial h(x - tG_{t}(x))
+% TODO: Need to check this equation
 \end{equation}

### 2017-10-10 Assignment-2/Solns-Assgn-2.tex
Assgn-2 Update 4

@@
 \documentclass[11pt]{article}
-\usepackage[margin=1in]{geometry}
+\usepackage[margin=0.8in]{geometry}
 \usepackage{amsmath}
@@
 
+\section*{Question 3}
+\subsection*{Part a}
+\begin{flushleft}
+Consider the SVD of \(W\) to be \(W = U_{W} \Sigma_{W} V^{T}_{W}\). From the conditions imposed on \(W\), we get:
+
+Hence the SVD of \(UV^{T} + W\) can be given by:
+\begin{equation}
+\label{svd-uv-w}
+UV^{T} + W = U I V^{T} + U_{W} \Sigma_{W} V^{T}_{W} = \begin{bmatrix} U & U_{W} \end{bmatrix} \begin{bmatrix} I & 0 \\ 0 & \Sigma_{W} \end{bmatrix} \begin{bmatrix} V^{T} \\ V_{W}^{T} \end{bmatrix}
+\end{equation}
+
+This is a valid SVD because:
+\begin{equation*}
+U^{T}W = 0 \implies U^{T}U_{W}\Sigma_{W}V^{T}_{W} = 0 \implies U^{T}U_{W}\Sigma_{W}V^{T}_{W}V_{W} = 0 \implies U^{T}U_{W}\Sigma_{W} = 0 \implies U^{T}U_{W} = 0
+\end{equation*}
+
+Similarly it can be shown that \(VV_{W}^{T} = 0\) (by using \(WU = 0\) and pre-multiplying \(U^{T}_{W}\)). From these results, it can be shown that:
+\begin{gather*}
+\begin{bmatrix} U & U_{W} \end{bmatrix}^{T} \begin{bmatrix} U & U_{W} \end{bmatrix} = I_{2r} \\
+\begin{bmatrix} V & V_{W} \end{bmatrix}^{T} \begin{bmatrix} V & V_{W} \end{bmatrix} = I_{2r}
+\end{gather*}
+
+Since \(||W||_{\text{op}} \leq 1\), we can tell that the diagonal entries of \(\Sigma_{W}\) are \(\leq 1\).
+Hence the maximum diagonal element in the singular matrix for \(UV^{T} + W\) from \ref{svd-uv-w} is \(1\), which suggests that \(||UV^{T} + W||_{\text{op}} \leq 1\)
+\end{flushleft}
+
+\subsection*{Part b}
+\begin{flushleft}
+Using the SVD of \(A\), we have:
+\begin{equation}
+\label{uv-w-a-prod}
+(UV^{T} + W)^{T}A = (VU^{T} + W^{T})(U\Sigma V^{T}) = VU^{T}U\Sigma V^{T} + W^{T}U\Sigma V^{T} = V\Sigma V^{T}
+\end{equation}
+
+Now, using the properties of circularity of the trace and \ref{uv-w-a-prod}, we get:
+\begin{equation}
+\label{part-3b}
+\text{tr}((UV^{T} + W)^{T}A) = \text{tr}(V\Sigma V^{T}) = \text{tr}(V^{T}V\Sigma) = \text{tr}(\Sigma) = ||A||_{*}
+\end{equation}
+\end{flushleft}
+
+\subsection*{Part c}
+\begin{flushleft}
+We know that the dual norm of the operator norm is the trace norm. Hence by the definition of trace norm:
+\begin{equation}
+\label{trace-norm-dual}
+\displaystyle ||A||_{*} = \max_{||B||_{\text{op}} \leq 1} \text{tr}(B^{T} A)
+\end{equation}
+From the result in \ref{ans-1a}, \ref{trace-norm-dual} and \(\nabla_{A} \text{tr}(B^{T}A) = B\), we know how to obtain the subdifferential of \(||A||_{*}\). This happens to be:
+\begin{equation}
+\label{subdiff-trace}
+\displaystyle \partial ||A||_{*} = \text{Conv}\left(\{B \ni \text{tr}(B^{T}A) = ||A||_{*}, ||B||_{\text{op}} \leq 1\}\right)
+\end{equation}
+From the results in \ref{part-3b} and \ref{svd-uv-w}, we can see that \(UV^{T} + W\) for any \(W\) satisfying the same condition as the elements of \(\partial ||A||_{*}\). Hence we can conclude that:
+\begin{multline}
+\label{subset-rel}
+\text{Conv}\left(\{B \ni \text{tr}(B^{T}A) = ||A||_{*}, ||B||_{\text{op}} \leq 1\}\right) \\ \supseteq \{UV^{T} + W \ni ||W||_{\text{op}} \leq 1, U^{T}W = 0, WV = 0, \text{tr}((UV^{T} + W)^{T}A) = ||A||_{*}\}
+\end{multline}
+
+Using \ref{subset-rel} and \ref{subdiff-trace}, we get the final result:
+\begin{equation}
+\label{ans-3c}
+\displaystyle \partial ||A||_{*} \supseteq \{UV^{T} + W \ni ||W||_{\text{op}} \leq 1, U^{T}W = 0, WV = 0, \text{tr}((UV^{T} + W)^{T}A) = ||A||_{*} \}
+\end{equation}
+
+\end{flushleft}
+
 \section*{Question 4}
@@
 \begin{flushleft}
-We know that if \(z = \textbf{prox}_{h, t}(x)\), then:
+By subgradient optimality condition:
 \begin{equation}
 \label{subgrad-optim}
-x - z \in \partial h(z) \implies x - \textbf{prox}_{h, t}(x) \in \partial h(\textbf{prox}_{h, t}(x))
+0 \in \partial\left(\frac{||z-x||^2}{2t} + h(z)\right)  \implies \frac{x - z}{t} \in \partial h(z) \implies \frac{x - \textbf{prox}_{h, t}(x)}{t} \in \partial h(\textbf{prox}_{h, t}(x))
 \end{equation}
@@
 \label{update-subgrad-optim}
-x - t\nabla g(x) - \textbf{prox}_{h, t}(x - t\nabla g(x)) \in \partial h(\textbf{prox}_{h, t}(x - t\nabla g(x)))
+\frac{x - t\nabla g(x) - \textbf{prox}_{h, t}(x - t\nabla g(x))}{t} \in \partial h(\textbf{prox}_{h, t}(x - t\nabla g(x)))
 \end{equation}
@@
 \label{sub-part-c}
-x - t\nabla g(x) - x + tG_{t}(x) \in \partial h(x - tG_{t}(x)) \implies tG_{t}(x) - t\nabla g(x) \in \partial h(x - tG_{t}(x)) 
+\frac{x - t\nabla g(x) - x + tG_{t}(x)}{t} \in \partial h(x - tG_{t}(x)) \implies \frac{tG_{t}(x) - t\nabla g(x)}{t} \in \partial h(x - tG_{t}(x)) 
 \end{equation}

### 2017-11-05 Paper-Presentation/main.tex
Start slides prep - PPT

\documentclass{beamer}
\usepackage{amsmath}
\usepackage{amssymb}
%\documentclass{article}
%\usepackage{beamerarticle}
\usetheme{CambridgeUS}

\title[]{Newton-Type Methods for Non-Convex Optimization Under Inexact Hessian Information}
\subtitle{Presented as a part of\\
CS6230 - Optimization Methods in Machine Learning}

\author{Vishwak Srinivasan \and Ayushi Patel}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSubsection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

% Let's get started
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{Introduction}
\subsection{The problem statement}

\begin{frame}{The problem statement}
  \begin{itemize}
  \item {
    Consider a unconstrained optimization problem
    \begin{equation}
        \min_{\mathbf{x} \in \mathbb{R}^{d}} F(\mathbf{x})
    \end{equation}
    where \(F : \mathbb{R}^{d} \rightarrow \mathbb{R}\), is \textit{smooth} and \textit{non-convex}.
  }

  \item<2-> {
    Many known methods to help solve the \emph{convex} version.
    \begin{itemize}
        \item First order methods
        \item Second order methods
    \end{itemize}
  }

  \item<3-> {
    First order methods are ``fine" - advances in tensor computing, automatic differentiation and so on.
    \begin{equation}
        \mathbf{x} := \mathbf{x} - \eta \nabla F(\mathbf{x})
    \end{equation}
    \pause
  }
  \item<4-> {
    Are second order methods just as fine?
    \begin{equation}
        \mathbf{x} := \mathbf{x} - \eta (\nabla^{2} F(\mathbf{x}))^{-1} \nabla F(\mathbf{x})
    \end{equation}
    \uncover<5->{How do we compute \(\nabla^{2} F(\mathbf{x})\)?}
  }
  \end{itemize}
\end{frame}

\subsection{Issues with the problem statement and motivation}

\begin{frame}{Issues with the problem statement}
  \begin{itemize}
  \item {
    Existing bounds for first order methods are for \emph{convex} \(F\).
    \begin{itemize}
      \uncover<2->{\item Are these methods even applicable to a \emph{non-convex} setting, as defined above?}
      \uncover<3->{\item Are there any mathematical guarantees?}
    \end{itemize}
  }

  \item<4-> {
    Second order methods: computation of \(\nabla^{2} F(\mathbf{x})\)
    \begin{itemize}
      \uncover<5->{\item How hard is this? Time Complexity Analysis}
      \uncover<6->{\item How hard is this? Space Complexity Analysis}
    \end{itemize}
  }
  \end{itemize}

  \uncover<7->{
    In Summary:\\
    Gradient based methods might not be able to optimize non-convex problems efficiently, exact hessian based methods could take a long time and a lot of space.
  }
\end{frame}

\section{Second Main Section}

\subsection{Another Subsection}

\begin{frame}{Blocks}
\begin{block}{Block Title}
You can also highlight sections of your presentation in a block, with it's own title
\end{block}
\begin{theorem}
There are separate environments for theorems, examples, definitions and proofs.
\end{theorem}
\begin{example}
Here is an example of an example block.
\end{example}
\end{frame}

% Placing a * after \section means it will not show in the
% outline or table of contents.
\section*{Summary}

\begin{frame}{Summary}
  \begin{itemize}
  \item
    The \alert{first main message} of your talk in one or two lines.
  \item
    The \alert{second main message} of your talk in one or two lines.
  \item
    Perhaps a \alert{third message}, but not more than that.
  \end{itemize}

  \begin{itemize}
  \item
    Outlook
    \begin{itemize}
    \item
      Something you haven't solved.
    \item
      Something else you haven't solved.
    \end{itemize}
  \end{itemize}
\end{frame}

% All of the following is optional and typically not needed. 
\appendix
\section<presentation>*{\appendixname}
\subsection<presentation>*{For Further Reading}

\begin{frame}[allowframebreaks]
  \frametitle<presentation>{For Further Reading}

  \begin{thebibliography}{10}

  \beamertemplatebookbibitems
  % Start with overview books.

  \bibitem{Author1990}
    A.~Author.
    \newblock {\em Handbook of Everything}.
    \newblock Some Press, 1990.

  \beamertemplatearticlebibitems
  % Followed by interesting articles. Keep the list short.

  \bibitem{Someone2000}
    S.~Someone.
    \newblock On this and that.
    \newblock {\em Journal of This and That}, 2(1):50--100,
    2000.
  \end{thebibliography}
\end{frame}

\end{document}

### 2017-11-05 Paper-Presentation/main.tex
Introduction and Definitions added

@@
 \usepackage{amssymb}
+\usepackage{bookmark}
+\usepackage{hyperref}
+\newcommand{\h}{\nabla^{2}}
+\newcommand{\g}{\nabla}
+\newcommand{\xbold}{\mathbf{x}}
 %\documentclass{article}
@@
 
-\title[]{Newton-Type Methods for Non-Convex Optimization Under Inexact Hessian Information}
+\title[Xu, Roosta-Khorasani and Mahoney]{Newton-Type Methods for Non-Convex Optimization Under Inexact Hessian Information}
 \subtitle{Presented as a part of\\
@@
 \author{Vishwak Srinivasan \and Ayushi Patel}
-
-% Delete this, if you do not want the table of contents to pop up at
-% the beginning of each subsection:
-\AtBeginSubsection[]
-{
-  \begin{frame}<beamer>{Outline}
-    \tableofcontents[currentsection]
-  \end{frame}
-}
 
@@
     \begin{equation}
-        \min_{\mathbf{x} \in \mathbb{R}^{d}} F(\mathbf{x})
+        \label{main-obj}
+        \min_{\xbold \in \mathbb{R}^{d}} F(\xbold)
     \end{equation}
@@
   }
-
   \item<2-> {
@@
   }
-
   \item<3-> {
@@
     \begin{equation}
-        \mathbf{x} := \mathbf{x} - \eta \nabla F(\mathbf{x})
+        \xbold := \xbold - \eta \g F(\xbold)
     \end{equation}
@@
     \begin{equation}
-        \mathbf{x} := \mathbf{x} - \eta (\nabla^{2} F(\mathbf{x}))^{-1} \nabla F(\mathbf{x})
+        \xbold := \xbold - \eta (\h F(\xbold))^{-1} \g F(\xbold)
     \end{equation}
-    \uncover<5->{How do we compute \(\nabla^{2} F(\mathbf{x})\)?}
+    \uncover<5->{How do we compute \(\h F(\xbold)\)?}
   }
@@
 
-\subsection{Issues with the problem statement and motivation}
-
-\begin{frame}{Issues with the problem statement}
+\subsection{Motivation for the work}
+\begin{frame}{Issues with the first order methods}
   \begin{itemize}
-  \item {
-    Existing bounds for first order methods are for \emph{convex} \(F\).
+  \item<1->{
+    Performance of first order methods can be seriously hindered by ill-conditioning.
     \begin{itemize}    
-      \uncover<2->{\item Are these methods even applicable to a \emph{non-convex} setting, as defined above?}
-      \uncover<3->{\item Are there any mathematical guarantees?}
+        \item<2->{Condition number: \(\kappa = \frac{L}{\gamma} = \frac{\lambda_{\max}(\h F)}{\lambda_{\min}(\h F)}\), where \(L\) is smoothness parameter, \(\gamma\) is strong convexity parameter}
+        \item<3->{Convergence criteria for gradient descent relies on \(\kappa\)} %add citations
     \end{itemize}
   }
-
-  \item<4-> {
-    Second order methods: computation of \(\nabla^{2} F(\mathbf{x})\)
+  \item<4->{
+    First order methods involve fine-tuning hyperparameters.
     \begin{itemize}    
-      \uncover<5->{\item How hard is this? Time Complexity Analysis}
-      \uncover<6->{\item How hard is this? Space Complexity Analysis}
+        \item<5->{For gradient descent, we have 1 (learning rate) and for Adam we have 4 (learning rate, tolerance, \(\beta_{1}\) and \(\beta_{2}\))}
+    \end{itemize}
+  }
+  \item<6->{
+    Theoretical guarantees exist for \emph{convex} settings only.
+    \begin{itemize}    
+        \item<7->{Work by Nesterov showed that we take \(O(1/\epsilon)\) iterations to converge to a first order stationary point using gradient descent} % add citations
+        \item<8->{\emph{Non-convex} problems have saddle-points to escape from too}
     \end{itemize}
@@
   \end{itemize}
-
-  \uncover<7->{
-    In Summary:\\
-    Gradient based methods might not be able to optimize non-convex problems efficiently, exact hessian based methods could take a long time and a lot of space.
-  }
 \end{frame}
 
-\section{Second Main Section}
-
-\subsection{Another Subsection}
-
-\begin{frame}{Blocks}
-\begin{block}{Block Title}
-You can also highlight sections of your presentation in a block, with it's own title
-\end{block}
-\begin{theorem}
-There are separate environments for theorems, examples, definitions and proofs.
-\end{theorem}
-\begin{example}
-Here is an example of an example block.
-\end{example}
-\end{frame}
-
-% Placing a * after \section means it will not show in the
-% outline or table of contents.
-\section*{Summary}
-
-\begin{frame}{Summary}
+\begin{frame}{Issues with the second order methods}
   \begin{itemize}
-  \item
-    The \alert{first main message} of your talk in one or two lines.
-  \item
-    The \alert{second main message} of your talk in one or two lines.
-  \item
-    Perhaps a \alert{third message}, but not more than that.
+  \item<1-> {
+    Second order methods: computation of \(\h F(\xbold)\)
+    \begin{itemize}
+      \uncover<2->{\item How hard is this? Time Complexity Analysis}
+      \uncover<3->{\item How hard is this? Space Complexity Analysis}
     \end{itemize}
-
-  \begin{itemize}
-  \item
-    Outlook
-    \begin{itemize}
-    \item
-      Something you haven't solved.
-    \item
-      Something else you haven't solved.
-    \end{itemize}
+  }
+  \item<4->{
+    Theoretical guarantees exist for \emph{convex} settings only.
+  } %add citations here as well
   \end{itemize}
@@
 
-% All of the following is optional and typically not needed. 
-\appendix
-\section<presentation>*{\appendixname}
-\subsection<presentation>*{For Further Reading}
-
-\begin{frame}[allowframebreaks]
-  \frametitle<presentation>{For Further Reading}
-
-  \begin{thebibliography}{10}
-
-  \beamertemplatebookbibitems
-  % Start with overview books.
-
-  \bibitem{Author1990}
-    A.~Author.
-    \newblock {\em Handbook of Everything}.
-    \newblock Some Press, 1990.
-
-  \beamertemplatearticlebibitems
-  % Followed by interesting articles. Keep the list short.
-
-  \bibitem{Someone2000}
-    S.~Someone.
-    \newblock On this and that.
-    \newblock {\em Journal of This and That}, 2(1):50--100,
-    2000.
-  \end{thebibliography}
+\section{Definitions}
+\begin{frame}{Special Notations}
+\begin{itemize}
+\item<1->{\(\g F(\xbold)\) is the gradient of \(F\) at \(\xbold\)}
+\item<2->{\(\h F(\xbold)\) is the (Exact) Hessian of \(F\) at \(\xbold \implies\) \(F\) is twice-differentiable}
+\item<3->{\(H_{t} \triangleq H(\xbold_{t})\) denotes the In-exact Hessian of \(F\) at the \(t^{th}\) iteration or at \(\xbold_{t}\)}
+\item<4->{\(A \succ B\) for two symmetric matrices \(A, B\) implies \(A - B \succ 0\). Similar notation holds for positive-semidefiniteness}
+\end{itemize}
 \end{frame}
 
+\begin{frame}{Definitions}
+\begin{alertblock}{\((\epsilon_{g}, \epsilon_{h})\)-Optimality}
+Given \(\epsilon_{g}, \epsilon_{h} \in (0, 1)\), \(\mathbf{\hat{x}}\) is an \((\epsilon_{g}, \epsilon_{h})\)-optimal solution to the problem \ref{main-obj}, if:
+\begin{itemize}
+\item \(||\g F(\mathbf{\hat{x}}) || \leq \epsilon_{g}\)
+\item \(\lambda_{\min} (\h F(\mathbf{\hat{x}})) \leq \epsilon_{h}\)
+\end{itemize}
+\end{alertblock}
+\pause
+Note: If the function \(F\) satisfies strict saddle property, then the above defined optimality condition ensures ``closeness'' to a local-minimum. % add citations
+\pause
+\begin{alertblock}{Lipschitz and Bounded Hessian (Hessian Regularity)}
+Given \(\xbold_{t}\) and \(\mathbf{s}_{t}\) the \(t^{th}\) iterate and \(t^{th}\) update step:
+\begin{itemize}
+\item \(||\h F(\xbold) - \h F(\xbold_{t})|| \leq L||\xbold - \xbold_{t}|| \hspace{4mm} \forall \xbold \in [\xbold_{t}, \xbold_{t} + \mathbf{s}_{t}]\)
+\item \(||\h F(\xbold_{t}) || \leq B\)
+\end{itemize}
+\end{alertblock}
+\end{frame}
 \end{document}

### 2017-11-05 Paper-Presentation/main.tex
Added TR and CR

@@
     \begin{itemize}    
-        \item<7->{Work by Nesterov showed that we take \(O(1/\epsilon)\) iterations to converge to a first order stationary point using gradient descent} % add citations
-        \item<8->{\emph{Non-convex} problems have saddle-points to escape from too}
+        \item<7->{\emph{Non-convex} problems have saddle-points to escape from}
     \end{itemize}
@@
 
-\section{Definitions}
+\section{Notations, Definitions and Preliminaries}
 \begin{frame}{Special Notations}
@@
 \begin{alertblock}{\((\epsilon_{g}, \epsilon_{h})\)-Optimality}
+\label{sec-opt}
 Given \(\epsilon_{g}, \epsilon_{h} \in (0, 1)\), \(\mathbf{\hat{x}}\) is an \((\epsilon_{g}, \epsilon_{h})\)-optimal solution to the problem \ref{main-obj}, if:
@@
 \item \(||\g F(\mathbf{\hat{x}}) || \leq \epsilon_{g}\)
-\item \(\lambda_{\min} (\h F(\mathbf{\hat{x}})) \leq \epsilon_{h}\)
+\item \(\lambda_{\min} (\h F(\mathbf{\hat{x}})) \geq -\epsilon_{h}\)
 \end{itemize}
@@
 \pause
-Note: If the function \(F\) satisfies strict saddle property, then the above defined optimality condition ensures ``closeness'' to a local-minimum. % add citations
+Note: If the function \(F\) satisfies \underline{strict saddle property}, then the above defined optimality condition ensures ``closeness'' to a local-minimum. % add citations
 \pause
@@
 \end{frame}
+
+\subsection{Preliminaries}
+\begin{frame}{Trust Region Methods}
+\begin{itemize}
+\item<1->{Method used to optimize over \emph{non-convex} functions, proven to be extremely useful at or near saddle points, by leveraging the negative curvature around that point}
+\item<2->{Helps achieve two things:
+  \begin{itemize}
+    \item<3->Finds optimal direction to move in
+    \item<4->Finds optimal distance to move in the determined direction
+  \end{itemize}
+  }
+\item<5->{Mathematically: if \(\mathbf{s}_{t}\) is the update to be made i.e., \(\xbold_{t+1} - \xbold_{t} = \mathbf{s}_{t}\), then
+\begin{equation}
+\label{TR}
+\mathbf{s}_{t} = \text{arg}\min_{\mathbf{s}} \frac{1}{2}\mathbf{s}^{T} \h F(\xbold_{t}) \mathbf{s} + \mathbf{s}^{T} \g F(\xbold_{t}) \hspace{5mm}\ni ||\mathbf{s}||_{2} \leq \Delta_{t}
+\end{equation} % add citation
+}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Trust Region Methods: Theoretical Guarantees}
+\begin{itemize}
+\item<1->{Solve a minimization problem which is a constrained quadratic problem}
+\item<2->{Similarities to other eigenvalue problems, leading to special methods drawing inferences from existing methods e.g., truncated conjugate gradient methods, Lanczos trust-region based methods}
+\item<3->{To reach an \((\epsilon_{g}, \epsilon_{h})\)-optimal point as defined earlier, we need \(O\left(\max\{\epsilon_{h}^{-1}\epsilon_{g}^{-2}, \epsilon_{h}^{-3}\}\right)\)} %add citation
+\item<4->{To reach an \((\epsilon_{g}, \epsilon_{h})\)-optimal point, we need \(O\left(\max\{\epsilon_{g}^{-3}, \epsilon_{h}^{-3}\}\right)\)} %add citation
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Cubic Regularization Methods}
+\begin{itemize}
+\item<1->{Method used to optimize over \emph{non-convex} functions, just as done by trust region methods.}
+\item<2->{Provides the same benefits provided by trust region methods
+  \begin{itemize}
+    \item<3->Finds optimal direction to move in
+    \item<4->Finds optimal distance to move in the determined direction
+  \end{itemize}
+  }
+\item<5->{Mathematically: if \(\mathbf{s}_{t}\) is the update to be made i.e., \(\xbold_{t+1} - \xbold_{t} = \mathbf{s}_{t}\), then
+\begin{equation}
+\label{CR}
+\mathbf{s}_{t} = \text{arg}\min_{\mathbf{s}} \frac{1}{2}\mathbf{s}^{T} \h F(\xbold_{t}) \mathbf{s} + \mathbf{s}^{T} \g F(\xbold_{t}) + \frac{\sigma_{t}}{3}||\mathbf{s}||^3
+\end{equation} % add citation
+}
+\item<6->{Notice familiarity?}
+\item<7->{What is different?}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Cubic Regularization Methods: Theoretical Guarantess}
+\begin{itemize}
+\item<1->{Solve a minimization problem which is an unconstrained quadratic problem, as opposed to constrained objective in \ref{TR}}
+\item<2->{Methods to solve this include Lanczos type iterations, fast approximate matrix inversion}
+\item<3->{To reach an \((\epsilon_{g}, \epsilon_{h})\)-optimal point as defined earlier, we need \(O\left(\max\{\epsilon_{g}^{-2}, \epsilon_{h}^{-3}\}\right)\)} %add citation
+\item<4->{Assuming more tight conditions helps improve the dependence on \(\epsilon_{g}\) - we need \(O\left(\max\{\epsilon_{g}^{-3/2}, \epsilon_{h}^{-3}\}\right)\)} %add citation
+\end{itemize}
+\end{frame}
 \end{document}

### 2017-11-05 Paper-Presentation/main.tex
Added TR and CR with Inexact Hessian

@@
 \newcommand{\xbold}{\mathbf{x}}
+\newcommand{\sbold}{\mathbf{s}}
+\newcommand{\mineig}{\lambda_{\min}}
+\newcommand{\maxeig}{\lambda_{\max}}
 %\documentclass{article}
@@
     \begin{itemize}    
-        \item<2->{Condition number: \(\kappa = \frac{L}{\gamma} = \frac{\lambda_{\max}(\h F)}{\lambda_{\min}(\h F)}\), where \(L\) is smoothness parameter, \(\gamma\) is strong convexity parameter}
+        \item<2->{Condition number: \(\kappa = \frac{L}{\gamma} = \frac{\maxeig(\h F)}{\mineig(\h F)}\), where \(L\) is smoothness parameter, \(\gamma\) is strong convexity parameter}
         \item<3->{Convergence criteria for gradient descent relies on \(\kappa\)} %add citations
@@
 \item \(||\g F(\mathbf{\hat{x}}) || \leq \epsilon_{g}\)
-\item \(\lambda_{\min} (\h F(\mathbf{\hat{x}})) \geq -\epsilon_{h}\)
+\item \(\mineig (\h F(\mathbf{\hat{x}})) \geq -\epsilon_{h}\)
 \end{itemize}
@@
 \begin{alertblock}{Lipschitz and Bounded Hessian (Hessian Regularity)}
-Given \(\xbold_{t}\) and \(\mathbf{s}_{t}\) the \(t^{th}\) iterate and \(t^{th}\) update step:
+Given \(\xbold_{t}\) and \(\sbold_{t}\) the \(t^{th}\) iterate and \(t^{th}\) update step respectively:
 \begin{itemize}
-\item \(||\h F(\xbold) - \h F(\xbold_{t})|| \leq L||\xbold - \xbold_{t}|| \hspace{4mm} \forall \xbold \in [\xbold_{t}, \xbold_{t} + \mathbf{s}_{t}]\)
-\item \(||\h F(\xbold_{t}) || \leq B\)
+\item \(||\h F(\xbold) - \h F(\xbold_{t})|| \leq L||\xbold - \xbold_{t}|| \hspace{4mm} \forall \xbold \in [\xbold_{t}, \xbold_{t} + \sbold_{t}]\)
+\item \(||\h F(\xbold_{t}) || \leq B, \hspace{4mm} B < \infty\)
 \end{itemize}
@@
   }
-\item<5->{Mathematically: if \(\mathbf{s}_{t}\) is the update to be made i.e., \(\xbold_{t+1} - \xbold_{t} = \mathbf{s}_{t}\), then
+\item<5->{Mathematically: if \(\sbold_{t}\) is the update to be made i.e., \(\xbold_{t+1} - \xbold_{t} = \sbold_{t}\), then
 \begin{equation}
 \label{TR}
-\mathbf{s}_{t} = \text{arg}\min_{\mathbf{s}} \frac{1}{2}\mathbf{s}^{T} \h F(\xbold_{t}) \mathbf{s} + \mathbf{s}^{T} \g F(\xbold_{t}) \hspace{5mm}\ni ||\mathbf{s}||_{2} \leq \Delta_{t}
+\sbold_{t} = \text{arg}\min_{\sbold} \frac{1}{2}\sbold^{T} \h F(\xbold_{t}) \sbold + \sbold^{T} \g F(\xbold_{t}) \hspace{5mm}\ni ||\sbold||_{2} \leq \Delta_{t}
 \end{equation} % add citation
@@
   }
-\item<5->{Mathematically: if \(\mathbf{s}_{t}\) is the update to be made i.e., \(\xbold_{t+1} - \xbold_{t} = \mathbf{s}_{t}\), then
+\item<5->{Mathematically: if \(\sbold_{t}\) is the update to be made i.e., \(\xbold_{t+1} - \xbold_{t} = \sbold_{t}\), then
 \begin{equation}
 \label{CR}
-\mathbf{s}_{t} = \text{arg}\min_{\mathbf{s}} \frac{1}{2}\mathbf{s}^{T} \h F(\xbold_{t}) \mathbf{s} + \mathbf{s}^{T} \g F(\xbold_{t}) + \frac{\sigma_{t}}{3}||\mathbf{s}||^3
+\sbold_{t} = \text{arg}\min_{\sbold} \frac{1}{2}\sbold^{T} \h F(\xbold_{t}) \sbold + \sbold^{T} \g F(\xbold_{t}) + \frac{\sigma_{t}}{3}||\sbold||^3
 \end{equation} % add citation
@@
 \end{frame}
+
+\section{Algorithms and associated convergence analysis}
+\begin{frame}{Conditions on the Inexact Hessian}
+\begin{alertblock}{Inexact Hessian Regularity}
+Given \(\xbold_{t}\) and \(\sbold_{t}\) the \(t^{th}\) iterate and \(t^{th}\) update step respectively:
+\begin{itemize}
+\item \(||\left(H(\xbold_{t}) - \h F(\xbold_{t})\right)\sbold_{t}|| \leq \epsilon||\sbold_{t}||, \hspace{4mm} \epsilon \in (0,1)\)
+\item \(||H(\xbold_{t}) || \leq B_{H}\)
+\end{itemize}
+\end{alertblock}
+\pause
+\begin{alertblock}{Sufficient Descent Condition-1}
+Assume that we solve \ref{TR} approximately to find \(\sbold_{t}\) such that:
+\begin{itemize}
+\item \(-m_{t}(\sbold_{t}) \geq -m_{t}(\sbold_{t}^{C}) \geq \min\lbrace\Phi_{1}(\g F(\xbold_{t}), H_{t}), \Psi_{1}(\g F(\xbold_{t}), \Delta_{t})\rbrace \)
+\item \(-m_{t}(\sbold_{t}) \geq -m_{t}(\sbold_{t}^{E}) \geq \frac{1}{2}\nu|\mineig(H_{t})|\Delta_{t}^{2}, \text{ if } \mineig(H_{t}) < 0 \)
+\end{itemize}
+where \(\sbold_{t}^{C}\) is a Cauchy point, along the negative gradient direction and \(\sbold_{t}^{E}\) is along negative curvature direction such that \((\sbold_{t}^{E})^{T} H_{t} \sbold_{t}^{E} \leq \nu\mineig(H_{t})||\sbold_{t}^{E}||^{2} < 0\) for some \(\nu \in (0, 1]\).
+\end{alertblock}
+\end{frame}
+
+\begin{frame}{Algorithm for Trust Region method}
+\begin{columns}
+\column{0.45\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{./images/algo-1.png}
+\column{0.55\textwidth}
+\begin{theorem}
+Consider any \(0 < \epsilon_{g}, \epsilon_{h} < 1\). Suppose
+\begin{itemize}
+\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon < (1 - \eta)\nu\epsilon_{h}\)}
+\item<2->{\(\h F(\xbold)\) satisfies Hessian regularity}
+\item<3->{if the approximate solution to \ref{TR} satisfies Sufficient Descent Condition-1}, 
+\uncover<4->{then the algorithm terminates in at most \(T = O\left(\max\{\epsilon_{h}^{-1}\epsilon_{g}^{-2}, \epsilon_{h}^{-3}\}\right)\) iterations}
+\end{itemize}
+\end{theorem}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Conditions on the Inexact Hessian}
+\begin{alertblock}{Sufficient Descent Condition-2}
+Assume that we solve \ref{CR} approximately to find \(\sbold_{t}\) such that:
+\begin{itemize}
+\item \(-m_{t}(\sbold_{t}) \geq -m_{t}(\sbold_{t}^{C}) \geq \max\lbrace\Phi_{2}(\sigma_{t}, \sbold_{t}, B_{H}, \g F(\xbold_{t})), \Psi_{2}(\sigma_{t}, B_{H}, \g F(\xbold_{t}))\rbrace\)
+\item \(-m_{t}(\sbold_{t}) \geq -m_{t}(\sbold_{t}^{C}) \geq \frac{\nu|\mineig(H_{t})|}{6} \max\lbrace ||\sbold_{t}^{E}||^{2}, \frac{\nu^{2}|\mineig(H_{t})|^{2}}{\sigma_{t}^{2}}\rbrace \text{ if } \mineig(H_{t}) < 0\)
+\end{itemize}
+\(\sbold_{t}^{C}\) is a Cauchy point, along the negative gradient direction and \(\sbold_{t}^{E}\) is along negative curvature direction such that \((\sbold_{t}^{E})^{T} H_{t} \sbold_{t}^{E} \leq \nu\mineig(H_{t})||\sbold_{t}^{E}||^{2} < 0\) for some \(\nu \in (0, 1]\).
+\end{alertblock}
+\pause
+\begin{alertblock}{Sufficient Descent for Optimal Complexity}
+Assume that we solve \ref{CR} approximately to find \(\sbold_{t}\) such that:
+\[||\nabla m_{t}(\sbold_{t})|| \leq \theta_{t}||\g F(\xbold_{t})||, \hspace{4mm} \text{ where }\theta_{t} = \zeta\min\{1, ||\sbold_{t}||\} \text{ and } \zeta \in (0, 1) \]
+\end{alertblock}
+\end{frame}
+
+\begin{frame}{Algorithm for Cubic Regularization method}
+\begin{columns}
+\column{0.45\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{./images/algo-2.png}
+\column{0.55\textwidth}
+\begin{theorem}
+Consider any \(0 < \epsilon_{g}, \epsilon_{h} < 1\). Suppose
+\begin{itemize}
+\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon = \min\{\epsilon_{0}, \zeta \epsilon_{g}\}, \zeta \in (0, 0.5)\)}
+\item<2->{\(\h F(\xbold)\) satisfies Hessian regularity}
+\item<3->{if the approximate solution to \ref{CR} satisfies Sufficient Descent Condition-2 and Sufficient Descent for Optimal Complexity}, 
+\uncover<4->{then the algorithm terminates in at most \(T = O\left(\max\{\epsilon_{g}^{-3/2}, \epsilon_{h}^{-3}\}\right)\) iterations}
+\end{itemize}
+\end{theorem}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Characteristics of Inexact Hessian in these methods}
+\begin{itemize}
+\item<1->{Same convergence rates with actual Hessians}
+\item<2->{Both the methods - trust region and cubic regularization, at termination satisfy
+  \begin{equation*}
+  ||\g F(\xbold_{T})|| \leq \epsilon_{g} \hspace{4mm} \text{ and } \hspace{4mm} \mineig(\h F(\xbold_{T})) \geq -(\epsilon_{h} + \epsilon)
+  \end{equation*}
+  \uncover<3->{in terms of the optimal point definition, these are \((\epsilon_{g}, \epsilon_{h} + \epsilon)\)-optimal points}
+} 
+\end{itemize}
+\end{frame}
 \end{document}

### 2017-11-05 Paper-Presentation/main.tex
Started Finite-sum minimization

@@
 \end{frame}
+
+\section{Finite-Sum Minimization}
+\begin{frame}{Randomized Sub-sampling}
+\begin{itemize}
+\item<1->{Consider a modified unconstrained optimization problem
+  \begin{equation}
+    \min_{\xbold \in \mathbb{R}^{d}} F(\xbold) \triangleq \frac{1}{n}\displaystyle \sum_{i=1}^{n} f_{i}(\xbold)
+  \end{equation}
+where each \(f_{i}\) is smooth, but possibly non-convex.}
+\item<2->{Seen in Machine Learning problems, where each \(f_{i}\) corresponds to the loss or mis-fit due to the \(i^{th}\) sample in the mini-batch/batch}
+\item<3->{Large-scale problems have \(n, d >> 1\).}
+\item<4->{Consider a probability distribution over \({1, 2, \ldots, n}\), denoted by \(\mathbf{p} = \{p_{i}\}_{i=1}^{n}\) and \(Pr(X = i) = p_{i} > 0, \hspace{2mm} \sum p_{i} = 1\).
+Based on \(\mathbf{p}\), we pick a sample \(\mathcal{S}\).}
+\item<5->{Define the sub-sampled Hessian to be:
+  \begin{equation}
+    H(\xbold) \triangleq \frac{1}{n|\mathcal{S}|}\sum_{j \in \mathcal{S}} \frac{1}{p_{j}} \h f_{j}(\xbold)
+  \end{equation}
+}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Randomized Sub-Sampling}
+\begin{theorem}
+Consider \(\displaystyle \sup_{\xbold\in\mathbb{R}^{d}} ||\h f_{i}(\xbold)|| \leq K_{i}, \hspace{2mm} \forall i \in \{1, 2, \ldots, n\}\).
+Let \(\displaystyle K_{\max} = \max_{i=1,\ldots, n} K_{i}\) and \(\hat{K} = \frac{1}{n} \displaystyle \sum_{i=1}^{n} K_{i}\).
+If \(p_{i} = \frac{1}{n} \hspace{1mm} \forall i \in \{1, 2, \ldots, n\}\), and \(|\mathcal{S}| \geq \frac{16K^{2}_{max}}{\epsilon^{2}}\log\frac{2d}{\delta} \), then with probability at least \(1 - \delta\), we have:
+\begin{equation}
+Pr\left(||H(\xbold) - \h F(\xbold)|| \leq \epsilon\right) \geq 1 - \delta
+\end{equation}
+\end{theorem}
+\pause
+\uncover<2->{Two major implications:}
+\begin{itemize}
+\item<3->{Higher the value of \(d\), more larger sample set is required. But this dependence is logarithmic.}
+\item<4->{Higher the number of samples, lower the value of \(\epsilon\), in turn better ``inexact'' Hessian.}
+\end{itemize}
+\end{frame}
 \end{document}

### 2017-11-06 Paper-Presentation/main.tex
Complete PPT

@@
-\documentclass{beamer}
+\documentclass[10pt]{beamer}
 \usepackage{amsmath}
@@
 Given \(\xbold_{t}\) and \(\sbold_{t}\) the \(t^{th}\) iterate and \(t^{th}\) update step respectively:
+\vspace{-2mm}
 \begin{itemize}
@@
 \section{Finite-Sum Minimization}
-\begin{frame}{Randomized Sub-sampling}
+\subsection{Random Sub-sampling}
+\begin{frame}{General problem}
 \begin{itemize}
@@
 
-\begin{frame}{Randomized Sub-Sampling}
+\begin{frame}{General problem - Uniform sampling of the Hessian}
 \begin{theorem}
 Consider \(\displaystyle \sup_{\xbold\in\mathbb{R}^{d}} ||\h f_{i}(\xbold)|| \leq K_{i}, \hspace{2mm} \forall i \in \{1, 2, \ldots, n\}\).
-Let \(\displaystyle K_{\max} = \max_{i=1,\ldots, n} K_{i}\) and \(\hat{K} = \frac{1}{n} \displaystyle \sum_{i=1}^{n} K_{i}\).
+Let \(\displaystyle K_{\max} = \max_{i=1,\ldots, n} K_{i}\).
 If \(p_{i} = \frac{1}{n} \hspace{1mm} \forall i \in \{1, 2, \ldots, n\}\), and \(|\mathcal{S}| \geq \frac{16K^{2}_{max}}{\epsilon^{2}}\log\frac{2d}{\delta} \), then with probability at least \(1 - \delta\), we have:
@@
 \pause
-\uncover<2->{Two major implications:}
+\uncover<2->{Some major implications:}
 \begin{itemize}
@@
 \end{frame}
+
+\begin{frame}{Special sub-problem}
+\begin{itemize}
+\item<1->{Applying a minor modification \(\xbold \rightarrow \mathbf{a}_{i}^{T}\xbold\), redefine the unconstrained optimization problem
+  \begin{equation}
+    \min_{\xbold \in \mathbb{R}^{d}} F(\xbold) \triangleq \frac{1}{n}\displaystyle \sum_{i=1}^{n} f_{i}(\mathbf{a}_{i}^{T}\xbold)
+  \end{equation}
+where each \(f_{i}\) is smooth, but possibly non-convex.}
+\item<2->{Provides a more clear analogy to Machine Learning problems, where \(\mathbf{a}_{i}\) is a ``data-vector" e.g., Linear Kernel SVMs, Logistic Regression}
+\item<3->{\(\h F(\xbold) = \frac{1}{n} \displaystyle \sum_{i=1}^{n} f_{i}''(\mathbf{a}_{i}^{T}\xbold) \mathbf{a}_{i}\mathbf{a}_{i}^{T}\)}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Special sub-problem - Non-uniform sampling of the Hessian}
+\begin{theorem}
+Consider \(\displaystyle \sup_{\xbold\in\mathbb{R}^{d}} ||\h f_{i}(\xbold)|| \leq K_{i}, \hspace{2mm} \forall i \in \{1, 2, \ldots, n\}\).
+Let \(\hat{K} = \frac{1}{n}\sum_{i=1}^{n} K_{i}\).
+If \(p_{i} = \displaystyle \frac{|f_{i}''(\mathbf{a}_{i}^{T}\xbold)| ||\mathbf{a}_{i}||^{2}_{2}}{N}\), where \(N\) is a normalizing factor and \(|\mathcal{S}| \geq \frac{4\hat{K}^{2}}{\epsilon^{2}}\log\frac{2d}{\delta}\), then with probability at least \(1 - \delta\), we have:
+\begin{equation}
+Pr\left(||H(\xbold) - \h F(\xbold)|| \leq \epsilon\right) \geq 1 - \delta
+\end{equation}
+\end{theorem}
+\pause
+\uncover<2->{Some major implications:}
+\begin{itemize}
+\item<3->{Higher the value of \(d\), more larger sample set is required. But this dependence is logarithmic.}
+\item<4->{Higher the number of samples, lower the value of \(\epsilon\), in turn better ``inexact'' Hessian.}
+\item<5->{Probabilities are also determined by the ``length" of the data.}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Usefulness and general ideas presented by these bounds}
+\begin{itemize}
+\item<1->{Both bounds depend on \(d\) in a logarithmic sense.}
+\item<2->{Larger the size of \(\mathcal{S}\), better the ``inexact" Hessian.
+  \begin{itemize}
+    \item<3->{\(\hat{K} \leq K_{\max}\). So?}
+    \item<4->{Smaller sample size \(|\mathcal{S}|\) required to achieve same level of accuracy in non-uniform sampling compared to uniform sampling}
+    \item<5->{Proves helpful if distribution of \(K_{i}'s\) is skewed i.e., \(\hat{K} << K_{\max}\)}
+  \end{itemize}
+}
+\item<3->{Choice of samples depend on the amount of curvature \(f_{i}''(\mathbf{a}_{i}^{T}\xbold)\), higher curvature \(\implies\) higher probability of being sampled.}
+\end{itemize}
+\end{frame}
+
+\subsection{Convergence Analysis of Finite-Sum Minimization}
+\begin{frame}{Probabilistic Convergence for Trust Region on Finite-Sum Minimization}
+\begin{theorem}
+Consider \(0 < \epsilon_{g}, \epsilon_{h} < 1\).
+\begin{itemize}
+\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon < (1- \eta)\nu\epsilon_{h}\)}
+\item<2->{\(\h F(\xbold)\) satisfies Hessian regularity}
+\item<3->{\(\delta_{0} = \delta\min\{\epsilon_{g}^{2}\epsilon_{h}, \epsilon_{h}^{3}\}\) for some \(\delta \in (0, 1)\)}
+\item<4->{\(\mathcal{S}\) sampled with parameters \(\epsilon, \delta_{0}\) using either uniform or non-uniform sampling.}
+\item<5->{if the approximate solution to \ref{TR} satisfies Sufficient Descent Condition-1, \uncover<6->{then the algorithm terminates in at most \(T = O\left(\max\{\epsilon_{h}^{-1}\epsilon_{g}^{-2}, \epsilon_{h}^{-3}\}\right)\) iterations, and with probability \(1 - \delta\), \(||\g F(\xbold_{T})|| \leq \epsilon_{g} \hspace{4mm} \text{ and } \hspace{4mm} \mineig(\h F(\xbold_{T})) \geq -(\epsilon + \epsilon_{h})\).}}
+\end{itemize}
+\end{theorem}
+\end{frame}
+
+\begin{frame}{Probabilistic Convergence for Cubic Regularization on Finite-Sum Minimization}
+\begin{theorem}
+Consider \(0 < \epsilon_{g}, \epsilon_{h} < 1\).
+\begin{itemize}
+\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon = \min\{\epsilon_{0}, \zeta\epsilon_{g}\}, \zeta \in (0, 0.5)\)}
+\item<2->{\(\h F(\xbold)\) satisfies Hessian regularity}
+\item<3->{\(\delta_{0} = \delta\min\{\epsilon_{g}^{3/2}, \epsilon_{h}^{3}\}\) for some \(\delta \in (0, 1)\)}
+\item<4->{\(\mathcal{S}\) sampled with parameters \(\epsilon, \delta_{0}\) using either uniform or non-uniform sampling.}
+\item<5->{if the approximate solution to \ref{CR} satisfies Sufficient Descent Condition-2 and Sufficient Descent for Optimal Complexity, \uncover<6->{then the algorithm terminates in at most \(T = O\left(\max\{\epsilon_{g}^{-3/2}, \epsilon_{h}^{-3}\}\right)\) iterations, and with probability \(1 - \delta\), \(||\g F(\xbold_{T})|| \leq \epsilon_{g} \hspace{4mm} \text{ and } \hspace{4mm} \mineig(\h F(\xbold_{T})) \geq -(\epsilon + \epsilon_{h})\).}}
+\end{itemize}
+\end{theorem}
+\end{frame}
+
+\section{Conclusion}
+\begin{frame}{Summary}
+\begin{itemize}
+\item<1->{Approximation of curvature information}
+\item<2->{Introduction to Trust Region and Cubic Regularization}
+\item<3->{Achieving approximate optimality based on conditions or properties of the inexact Hessian}
+\item<4->{Introduction to subsampling strategies}
+\item<5->{Probabilistic convergence for Trust Region and Cubic Regularization on the finite-sum minimization problem}
+\item<6->{Very theoretical - please view \href{arxiv.org/abs/1708.07827}{\beamergotobutton{Second-Order Optimization for Non-Convex Machine Learning: An Empirical Study}}}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Questions??}
+\end{frame}
 \end{document}

### 2017-11-06 Paper-Presentation/main.tex
Added References

@@
 \newcommand{\maxeig}{\lambda_{\max}}
+\newcommand{\eg}{\epsilon_{g}}
+\newcommand{\eh}{\epsilon_{h}}
 %\documentclass{article}
@@
 
-% Let's get started
 \begin{document}
@@
         \item<2->{Condition number: \(\kappa = \frac{L}{\gamma} = \frac{\maxeig(\h F)}{\mineig(\h F)}\), where \(L\) is smoothness parameter, \(\gamma\) is strong convexity parameter}
-        \item<3->{Convergence criteria for gradient descent relies on \(\kappa\)} %add citations
+        \item<3->{Convergence criteria for gradient descent relies on \(\kappa\) \cite{benrecht-simons}}
     \end{itemize}
@@
     \begin{itemize}    
-        \item<5->{For gradient descent, we have 1 (learning rate) and for Adam we have 4 (learning rate, tolerance, \(\beta_{1}\) and \(\beta_{2}\))}
+        \item<5->{For gradient descent, we have 1 (learning rate) and for Adam \cite{adam-iclr} we have 4 (learning rate, tolerance, \(\beta_{1}\) and \(\beta_{2}\))}
     \end{itemize}
@@
     \begin{itemize}    
-        \item<7->{\emph{Non-convex} problems have saddle-points to escape from}
+        \item<7->{\emph{Non-convex} problems have saddle-points to escape from \cite{du-nips}, \cite{jin-icml}}
     \end{itemize}
@@
     Theoretical guarantees exist for \emph{convex} settings only.
-  } %add citations here as well
+  }
   \end{itemize}
@@
 \begin{frame}{Definitions}
-\begin{alertblock}{\((\epsilon_{g}, \epsilon_{h})\)-Optimality}
+\begin{alertblock}{\((\eg, \eh)\)-Optimality}
 \label{sec-opt}
-Given \(\epsilon_{g}, \epsilon_{h} \in (0, 1)\), \(\mathbf{\hat{x}}\) is an \((\epsilon_{g}, \epsilon_{h})\)-optimal solution to the problem \ref{main-obj}, if:
+Given \(\eg, \eh \in (0, 1)\), \(\mathbf{\hat{x}}\) is an \((\eg, \eh)\)-optimal solution to the problem \ref{main-obj}, if:
 \begin{itemize}
-\item \(||\g F(\mathbf{\hat{x}}) || \leq \epsilon_{g}\)
-\item \(\mineig (\h F(\mathbf{\hat{x}})) \geq -\epsilon_{h}\)
+\item \(||\g F(\mathbf{\hat{x}}) || \leq \eg\)
+\item \(\mineig (\h F(\mathbf{\hat{x}})) \geq -\eh\)
 \end{itemize}
@@
 \pause
-Note: If the function \(F\) satisfies \underline{strict saddle property}, then the above defined optimality condition ensures ``closeness'' to a local-minimum. % add citations
+Note: If the function \(F\) satisfies \underline{strict saddle property} \cite{ge-colt}, then the above defined optimality condition ensures ``closeness'' to a local-minimum.
 \pause
@@
 Given \(\xbold_{t}\) and \(\sbold_{t}\) the \(t^{th}\) iterate and \(t^{th}\) update step respectively:
-\vspace{-2mm}
 \begin{itemize}
@@
 \begin{itemize}
-\item<1->{Method used to optimize over \emph{non-convex} functions, proven to be extremely useful at or near saddle points, by leveraging the negative curvature around that point}
+\item<1->{Method used to optimize over \emph{non-convex} functions, proven to be extremely useful at or near saddle points, by leveraging the negative curvature around that point \cite{conn-siam}, \cite{sorensen-siam}}
 \item<2->{Helps achieve two things:
@@
 \item<2->{Similarities to other eigenvalue problems, leading to special methods drawing inferences from existing methods e.g., truncated conjugate gradient methods, Lanczos trust-region based methods}
-\item<3->{To reach an \((\epsilon_{g}, \epsilon_{h})\)-optimal point as defined earlier, we need \(O\left(\max\{\epsilon_{h}^{-1}\epsilon_{g}^{-2}, \epsilon_{h}^{-3}\}\right)\)} %add citation
-\item<4->{To reach an \((\epsilon_{g}, \epsilon_{h})\)-optimal point, we need \(O\left(\max\{\epsilon_{g}^{-3}, \epsilon_{h}^{-3}\}\right)\)} %add citation
+\item<3->{To reach an \((\eg, \eh)\)-optimal point as defined earlier, we need \(O\left(\max\{\eh^{-1}\eg^{-2}, \eh^{-3}\}\right)\) \cite{cartis-joc}} 
+\item<4->{To reach an \((\eg, \eh)\)-optimal point, we need \(O\left(\max\{\eg^{-3}, \eh^{-3}\}\right)\) \cite{geovani-oms}}
 \end{itemize}
@@
 \sbold_{t} = \text{arg}\min_{\sbold} \frac{1}{2}\sbold^{T} \h F(\xbold_{t}) \sbold + \sbold^{T} \g F(\xbold_{t}) + \frac{\sigma_{t}}{3}||\sbold||^3
-\end{equation} % add citation
+\end{equation}
 }
@@
 \item<2->{Methods to solve this include Lanczos type iterations, fast approximate matrix inversion}
-\item<3->{To reach an \((\epsilon_{g}, \epsilon_{h})\)-optimal point as defined earlier, we need \(O\left(\max\{\epsilon_{g}^{-2}, \epsilon_{h}^{-3}\}\right)\)} %add citation
-\item<4->{Assuming more tight conditions helps improve the dependence on \(\epsilon_{g}\) - we need \(O\left(\max\{\epsilon_{g}^{-3/2}, \epsilon_{h}^{-3}\}\right)\)} %add citation
+\item<3->{To reach an \((\eg, \eh)\)-optimal point as defined earlier, we need \(O\left(\max\{\eg^{-2}, \eh^{-3}\}\right)\) \cite{cartis-joc}}
+\item<4->{Assuming more tight conditions helps improve the dependence on \(\eg\) - we need \(O\left(\max\{\eg^{-3/2}, \eh^{-3}\}\right)\) \cite{cartis-siam}} 
 \end{itemize}
@@
 \begin{theorem}
-Consider any \(0 < \epsilon_{g}, \epsilon_{h} < 1\). Suppose
+Consider any \(0 < \eg, \eh < 1\). Suppose
 \begin{itemize}
-\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon < (1 - \eta)\nu\epsilon_{h}\)}
+\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon < (1 - \eta)\nu\eh\)}
 \item<2->{\(\h F(\xbold)\) satisfies Hessian regularity}
 \item<3->{if the approximate solution to \ref{TR} satisfies Sufficient Descent Condition-1}, 
-\uncover<4->{then the algorithm terminates in at most \(T = O\left(\max\{\epsilon_{h}^{-1}\epsilon_{g}^{-2}, \epsilon_{h}^{-3}\}\right)\) iterations}
+\uncover<4->{then the algorithm terminates in at most \(T = O\left(\max\{\eh^{-1}\eg^{-2}, \eh^{-3}\}\right)\) iterations}
 \end{itemize}
@@
 \begin{theorem}
-Consider any \(0 < \epsilon_{g}, \epsilon_{h} < 1\). Suppose
+Consider any \(0 < \eg, \eh < 1\). Suppose
 \begin{itemize}
-\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon = \min\{\epsilon_{0}, \zeta \epsilon_{g}\}, \zeta \in (0, 0.5)\)}
+\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon = \min\{\epsilon_{0}, \zeta \eg\}, \zeta \in (0, 0.5)\)}
 \item<2->{\(\h F(\xbold)\) satisfies Hessian regularity}
 \item<3->{if the approximate solution to \ref{CR} satisfies Sufficient Descent Condition-2 and Sufficient Descent for Optimal Complexity}, 
-\uncover<4->{then the algorithm terminates in at most \(T = O\left(\max\{\epsilon_{g}^{-3/2}, \epsilon_{h}^{-3}\}\right)\) iterations}
+\uncover<4->{then the algorithm terminates in at most \(T = O\left(\max\{\eg^{-3/2}, \eh^{-3}\}\right)\) iterations}
 \end{itemize}
@@
   \begin{equation*}
-  ||\g F(\xbold_{T})|| \leq \epsilon_{g} \hspace{4mm} \text{ and } \hspace{4mm} \mineig(\h F(\xbold_{T})) \geq -(\epsilon_{h} + \epsilon)
+  ||\g F(\xbold_{T})|| \leq \eg \hspace{4mm} \text{ and } \hspace{4mm} \mineig(\h F(\xbold_{T})) \geq -(\eh + \epsilon)
   \end{equation*}
-  \uncover<3->{in terms of the optimal point definition, these are \((\epsilon_{g}, \epsilon_{h} + \epsilon)\)-optimal points}
+  \uncover<3->{in terms of the optimal point definition, these are \((\eg, \eh + \epsilon)\)-optimal points}
 } 
@@
 \begin{theorem}
-Consider \(0 < \epsilon_{g}, \epsilon_{h} < 1\).
+Consider \(0 < \eg, \eh < 1\).
 \begin{itemize}
-\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon < (1- \eta)\nu\epsilon_{h}\)}
+\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon < (1- \eta)\nu\eh\)}
 \item<2->{\(\h F(\xbold)\) satisfies Hessian regularity}
-\item<3->{\(\delta_{0} = \delta\min\{\epsilon_{g}^{2}\epsilon_{h}, \epsilon_{h}^{3}\}\) for some \(\delta \in (0, 1)\)}
+\item<3->{\(\delta_{0} = \delta\min\{\eg^{2}\eh, \eh^{3}\}\) for some \(\delta \in (0, 1)\)}
 \item<4->{\(\mathcal{S}\) sampled with parameters \(\epsilon, \delta_{0}\) using either uniform or non-uniform sampling.}
-\item<5->{if the approximate solution to \ref{TR} satisfies Sufficient Descent Condition-1, \uncover<6->{then the algorithm terminates in at most \(T = O\left(\max\{\epsilon_{h}^{-1}\epsilon_{g}^{-2}, \epsilon_{h}^{-3}\}\right)\) iterations, and with probability \(1 - \delta\), \(||\g F(\xbold_{T})|| \leq \epsilon_{g} \hspace{4mm} \text{ and } \hspace{4mm} \mineig(\h F(\xbold_{T})) \geq -(\epsilon + \epsilon_{h})\).}}
+\item<5->{if the approximate solution to \ref{TR} satisfies Sufficient Descent Condition-1, \uncover<6->{then the algorithm terminates in at most \(T = O\left(\max\{\eh^{-1}\eg^{-2}, \eh^{-3}\}\right)\) iterations, and with probability \(1 - \delta\), \(||\g F(\xbold_{T})|| \leq \eg \hspace{4mm} \text{ and } \hspace{4mm} \mineig(\h F(\xbold_{T})) \geq -(\epsilon + \eh)\).}}
 \end{itemize}
@@
 \begin{theorem}
-Consider \(0 < \epsilon_{g}, \epsilon_{h} < 1\).
+Consider \(0 < \eg, \eh < 1\).
 \begin{itemize}
-\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon = \min\{\epsilon_{0}, \zeta\epsilon_{g}\}, \zeta \in (0, 0.5)\)}
+\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon = \min\{\epsilon_{0}, \zeta\eg\}, \zeta \in (0, 0.5)\)}
 \item<2->{\(\h F(\xbold)\) satisfies Hessian regularity}
-\item<3->{\(\delta_{0} = \delta\min\{\epsilon_{g}^{3/2}, \epsilon_{h}^{3}\}\) for some \(\delta \in (0, 1)\)}
+\item<3->{\(\delta_{0} = \delta\min\{\eg^{3/2}, \eh^{3}\}\) for some \(\delta \in (0, 1)\)}
 \item<4->{\(\mathcal{S}\) sampled with parameters \(\epsilon, \delta_{0}\) using either uniform or non-uniform sampling.}
-\item<5->{if the approximate solution to \ref{CR} satisfies Sufficient Descent Condition-2 and Sufficient Descent for Optimal Complexity, \uncover<6->{then the algorithm terminates in at most \(T = O\left(\max\{\epsilon_{g}^{-3/2}, \epsilon_{h}^{-3}\}\right)\) iterations, and with probability \(1 - \delta\), \(||\g F(\xbold_{T})|| \leq \epsilon_{g} \hspace{4mm} \text{ and } \hspace{4mm} \mineig(\h F(\xbold_{T})) \geq -(\epsilon + \epsilon_{h})\).}}
+\item<5->{if the approximate solution to \ref{CR} satisfies Sufficient Descent Condition-2 and Sufficient Descent for Optimal Complexity, \uncover<6->{then the algorithm terminates in at most \(T = O\left(\max\{\eg^{-3/2}, \eh^{-3}\}\right)\) iterations, and with probability \(1 - \delta\), \(||\g F(\xbold_{T})|| \leq \eg \hspace{4mm} \text{ and } \hspace{4mm} \mineig(\h F(\xbold_{T})) \geq -(\epsilon + \eh)\).}}
 \end{itemize}
@@
 \end{frame}
+
+\begin{frame}[t,allowframebreaks]
+\frametitle{References}
+\bibliographystyle{plain}
+\bibliography{main.bib}
+\end{frame}
 \end{document}

### 2017-11-06 Paper-Presentation/main.tex
Typo fixes

@@
         \item<2->{Condition number: \(\kappa = \frac{L}{\gamma} = \frac{\maxeig(\h F)}{\mineig(\h F)}\), where \(L\) is smoothness parameter, \(\gamma\) is strong convexity parameter}
-        \item<3->{Convergence criteria for gradient descent relies on \(\kappa\) \cite{benrecht-simons}}
+        \item<3->{Convergence criteria for gradient descent relies on \(\kappa\) (\cite{benrecht-simons})}
     \end{itemize}
@@
     \begin{itemize}    
-        \item<5->{For gradient descent, we have 1 (learning rate) and for Adam \cite{adam-iclr} we have 4 (learning rate, tolerance, \(\beta_{1}\) and \(\beta_{2}\))}
+        \item<5->{For gradient descent, we have 1 (learning rate) and for Adam (\cite{adam-iclr}) we have 4 (learning rate, tolerance, \(\beta_{1}\) and \(\beta_{2}\))}
     \end{itemize}
@@
     \begin{itemize}    
-        \item<7->{\emph{Non-convex} problems have saddle-points to escape from \cite{du-nips}, \cite{jin-icml}}
+        \item<7->{\emph{Non-convex} problems have saddle-points to escape from (\cite{du-nips}, \cite{jin-icml})}
     \end{itemize}
@@
 \pause
-Note: If the function \(F\) satisfies \underline{strict saddle property} \cite{ge-colt}, then the above defined optimality condition ensures ``closeness'' to a local-minimum.
+Note: If the function \(F\) satisfies \underline{strict saddle property} (\cite{ge-colt}), then the above defined optimality condition ensures ``closeness'' to a local-minimum.
 \pause
@@
 \begin{itemize}
-\item<1->{Method used to optimize over \emph{non-convex} functions, proven to be extremely useful at or near saddle points, by leveraging the negative curvature around that point \cite{conn-siam}, \cite{sorensen-siam}}
+\item<1->{Method used to optimize over \emph{non-convex} functions, proven to be extremely useful at or near saddle points, by leveraging the negative curvature around that point (\cite{conn-siam}, \cite{sorensen-siam})}
 \item<2->{Helps achieve two things:
@@
 \item<2->{Similarities to other eigenvalue problems, leading to special methods drawing inferences from existing methods e.g., truncated conjugate gradient methods, Lanczos trust-region based methods}
-\item<3->{To reach an \((\eg, \eh)\)-optimal point as defined earlier, we need \(O\left(\max\{\eh^{-1}\eg^{-2}, \eh^{-3}\}\right)\) \cite{cartis-joc}} 
-\item<4->{To reach an \((\eg, \eh)\)-optimal point, we need \(O\left(\max\{\eg^{-3}, \eh^{-3}\}\right)\) \cite{geovani-oms}}
+\item<3->{To reach an \((\eg, \eh)\)-optimal point as defined earlier, we need \(O\left(\max\{\eh^{-1}\eg^{-2}, \eh^{-3}\}\right)\) (\cite{cartis-joc})} 
+\item<4->{To reach an \((\eg, \eh)\)-optimal point, we need \(O\left(\max\{\eg^{-3}, \eh^{-3}\}\right)\) (\cite{geovani-oms})}
 \end{itemize}
@@
 \item<2->{Methods to solve this include Lanczos type iterations, fast approximate matrix inversion}
-\item<3->{To reach an \((\eg, \eh)\)-optimal point as defined earlier, we need \(O\left(\max\{\eg^{-2}, \eh^{-3}\}\right)\) \cite{cartis-joc}}
-\item<4->{Assuming more tight conditions helps improve the dependence on \(\eg\) - we need \(O\left(\max\{\eg^{-3/2}, \eh^{-3}\}\right)\) \cite{cartis-siam}} 
+\item<3->{To reach an \((\eg, \eh)\)-optimal point as defined earlier, we need \(O\left(\max\{\eg^{-2}, \eh^{-3}\}\right)\) (\cite{cartis-joc})}
+\item<4->{Assuming more tight conditions helps improve the dependence on \(\eg\) - we need \(O\left(\max\{\eg^{-3/2}, \eh^{-3}\}\right)\) (\cite{cartis-siam})}
 \end{itemize}
@@
 \item \(||\left(H(\xbold_{t}) - \h F(\xbold_{t})\right)\sbold_{t}|| \leq \epsilon||\sbold_{t}||, \hspace{4mm} \epsilon \in (0,1)\)
-\item \(||H(\xbold_{t}) || \leq B_{H}\)
+\item \(||H(\xbold_{t}) || \leq B_{H}, \hspace{4mm} B_{H} < \infty\)
 \end{itemize}
@@
 \item<3->{Large-scale problems have \(n, d >> 1\).}
-\item<4->{Consider a probability distribution over \({1, 2, \ldots, n}\), denoted by \(\mathbf{p} = \{p_{i}\}_{i=1}^{n}\) and \(Pr(X = i) = p_{i} > 0, \hspace{2mm} \sum p_{i} = 1\).
+\item<4->{Consider a probability distribution over \(\{1, 2, \ldots, n\}\), denoted by \(\mathbf{p} = \{p_{i}\}_{i=1}^{n}\) and \(Pr(X = i) = p_{i} > 0, \hspace{2mm} \sum p_{i} = 1\).
 Based on \(\mathbf{p}\), we pick a sample \(\mathcal{S}\).}
@@
 }
-\item<3->{Choice of samples depend on the amount of curvature \(f_{i}''(\mathbf{a}_{i}^{T}\xbold)\), higher curvature \(\implies\) higher probability of being sampled.}
+\item<6->{Choice of samples depend on the amount of curvature \(f_{i}''(\mathbf{a}_{i}^{T}\xbold)\), higher curvature \(\implies\) higher probability of being sampled.}
 \end{itemize}

### 2017-11-09 Programming-Assignment-2/primal.py
PA-2 Update 1

@@ -1,0 +1,68 @@
+import h5py as h
+import cvxpy as C
+import numpy as np
+from matplotlib import pyplot as plt
+from argparse import ArgumentParser as AP
+
+p = AP()
+p.add_argument('--C1', required=True, type=float, help='C1 parameter')
+p.add_argument('--C2', required=True, type=float, help='C2 parameter')
+opt = p.parse_args()
+
+# Load the dataset
+f = h.File('toy.hdf5', 'r')
+X = np.array(list(f['X'])).T
+y = np.array(list(f['y'])).T
+print('Dataset loaded: X.shape = {0}'.format(X.shape))
+
+# Declare the variables to optimize over
+m, d = X.shape[0], X.shape[1]
+beta = C.Variable(d)
+beta_0 = C.Variable(1)
+slackvar = C.Variable(m)
+
+# Declare the objective function
+objective = (C.sum_squares(beta))*0.5
+for i in range(0, m):
+    if y[i] == 1:
+        objective += slackvar[i]*opt.C1
+    else:
+        objective += slackvar[i]*opt.C2
+
+# Declare the constraints to enforce
+constraints = [slackvar >= 0]
+for i in range(0, m):
+    constraints += [y[i]*(beta.T*X[i] + beta_0) >= 1 - slackvar[i]]
+
+# Define the problem
+problem = C.Problem(C.Minimize(objective), constraints)
+
+# Solve the problem
+problem.solve(solver=C.CVXOPT)
+
+print("Problem exited with status: {0} and value attained: {1}".format(problem.status, problem.value))
+
+# Plotting section
+label_flag = [False, False]
+for i in range(0, m):
+    if y[i] == 1:
+        if label_flag[0] == False:        
+            plt.plot(X[i,0], X[i,1], 'go', label='+1')
+            label_flag[0] = True
+        else:
+            plt.plot(X[i,0], X[i,1], 'go')
+    else:
+        if label_flag[1] == False:        
+            plt.plot(X[i,0], X[i,1], 'bo', label='-1')
+            label_flag[1] = True
+        else:
+            plt.plot(X[i,0], X[i,1], 'bo')
+
+x = np.arange(np.amin(X[:,0]), np.amax(X[:,0]), 0.001)
+beta = (beta.value).tolist()
+beta_0 = beta_0.value
+y = -beta_0/beta[1] - beta[0]*x/beta[1]
+plt.plot(x, y, 'r--')
+
+plt.legend(loc='upper right', numpoints=1)
+plt.show()

### 2017-11-09 Programming-Assignment-2/primal.py
PA-2 Update 2

@@ -46,23 +46,33 @@ print("Problem exited with status: {0} and value attained: {1}".format(problem.s
 label_flag = [False, False]
 for i in range(0, m):
     if y[i] == 1:
-        if label_flag[0] == False:        
-            plt.plot(X[i,0], X[i,1], 'go', label='+1')
+        point_type = 'ro' if (slackvar[i].value) > 1e-07 else 'go'
+
+        if label_flag[0] == False and point_type == 'go':
+            plt.plot(X[i,0], X[i,1], point_type, label='+1')
             label_flag[0] = True
         else:
-            plt.plot(X[i,0], X[i,1], 'go')
+            plt.plot(X[i,0], X[i,1], point_type)
     else:
-        if label_flag[1] == False:        
-            plt.plot(X[i,0], X[i,1], 'bo', label='-1')
+        point_type = 'ro' if (slackvar[i].value) > 1e-07 else 'bo'
+
+        if label_flag[1] == False and point_type == 'bo':        
+            plt.plot(X[i,0], X[i,1], point_type, label='-1')
             label_flag[1] = True
         else:
-            plt.plot(X[i,0], X[i,1], 'bo')
+            plt.plot(X[i,0], X[i,1], point_type)
 
 x = np.arange(np.amin(X[:,0]), np.amax(X[:,0]), 0.001)
-beta = (beta.value).tolist()
+beta = np.array(beta.value).reshape(-1).tolist()
 beta_0 = beta_0.value
 y = -beta_0/beta[1] - beta[0]*x/beta[1]
-plt.plot(x, y, 'r--')
+plt.plot(x, y, 'k-')
 
+y_margin1 = (1 - beta_0 - beta[0]*x)/beta[1]
+plt.plot(x, y_margin1, 'r--')
+y_margin2 = (-1 - beta_0 - beta[0]*x)/beta[1]
+plt.plot(x, y_margin2, 'r--')
+
+print("Margin width: {0}".format(2/np.sqrt(beta[0]**2 + beta[1]**2)))
 plt.legend(loc='upper right', numpoints=1)
 plt.show()

### 2017-11-09 Programming-Assignment-2/dual.py
PA-2 Update 3

@@ -1,0 +1,47 @@
+import h5py as h
+import cvxpy as C
+import numpy as np
+from matplotlib import pyplot as plt
+from argparse import ArgumentParser as AP
+
+p = AP()
+p.add_argument('--C1', required=True, type=float, help='C1 parameter')
+p.add_argument('--C2', required=True, type=float, help='C2 parameter')
+opt = p.parse_args()
+
+# Load the dataset
+f = h.File('toy.hdf5', 'r')
+X = np.array(list(f['X'])).T
+y = np.array(list(f['y'])).T
+print('Dataset loaded: X.shape = {0}'.format(X.shape))
+
+# Declare the variables to optimize over
+m, d = X.shape[0], X.shape[1]
+alpha = C.Variable(m)
+
+# Declare the objective function
+# First calculate X(tilde)X(tilde)^{T}
+constant_matrix = np.zeros((m, m))
+for i in range(0, m):
+    for j in range(0, m):
+        constant_matrix[i,j] = y[i]*y[j]*np.dot(X[i], X[j])
+
+obj_1 = C.sum_entries(alpha)
+obj_2 = C.sum_entries(C.quad_form(alpha, constant_matrix))*0.5
+objective = obj_1 - obj_2
+
+# Declare the constraints to enforce
+constraints = [C.sum_entries(alpha.T*y) == 0, alpha >= 0]
+for i in range(0, m):
+    if y[i] == 1:    
+        constraints += [alpha[i] <= opt.C1]
+    else:
+        constraints += [alpha[i] <= opt.C2]
+
+# Define the problem
+problem = C.Problem(C.Maximize(objective), constraints)
+
+# Solve the problem
+problem.solve(solver=C.CVXOPT)
+
+print("Problem exited with status: {0} and value attained: {1}".format(problem.status, problem.value))

### 2017-11-09 Programming-Assignment-2/PA_2-report.tex
PA-2 Update 4

\documentclass{article}
\usepackage[margin=0.73in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{hyperref}

\title{Report for Programming Assignment 2}
\author{Vishwak Srinivasan\\
\texttt{CS15BTECH11043}}
\date{}

\begin{document}
\maketitle

\section*{Part b}
\subsection*{Question 1}
\begin{flushleft}
Below is the table representing the objective value of the optimal solution. The program used \texttt{CVXPY}, a convex programming module in Python.
\begin{center}
\begin{tabular}{|c|c|}
\hline
Penalty Parameter combination \((C_{1}, C_{2})\) & Objective Value \\
\hline
\hline
\((1, 1)\) & \(9.60436\)\\
\hline
\((1, 10)\) & \(30.26522\)\\
\hline
\((10, 1)\) & \(11.89106\)\\
\hline
\end{tabular}
\end{center}
\end{flushleft}

\subsection*{Question 2}
The reason of different decision boundaries occuring with variation in the penalty parameters could be attributed to the relative weights assigned to the classes. For example, if we set \(C_{1} = C_{2} = 1\), we are implicitly stating that both the classes \(\{+1, -1\}\) are equally important. On the other hand, if we assign \(C_{1} = 10, C_{2} = 1\), then we are implicitly stating that \(+1\) gets a higher priority over \(-1\), in the sense that we can't afford to mis-classify many of \(+1\) datapoints, but we could rather compromise with \(-1\) datapoints. An analogous case follows for \(C_{1} = 1, C_{2} = 10\), wherein we ``weigh'' \(-1\) more than \(+1\). The variation can be seen as the decision boundary tries to keep a higher number of examples to the class having a higher penalty on one side of itself.

\end{document}

### 2017-11-09 Programming-Assignment-2/dual.py

@@ -44,4 +44,4 @@ problem = C.Problem(C.Maximize(objective), constraints)
 # Solve the problem
 problem.solve(solver=C.CVXOPT)
 
-print("Problem exited with status: {0} and value attained: {1}".format(problem.status, problem.value))
+print("Problem exited with status: {0} and value attained: {1}".format(problem.status, round(problem.value, 5)))

### 2017-11-09 Programming-Assignment-2/primal.py

@@ -40,7 +40,7 @@ problem = C.Problem(C.Minimize(objective), constraints)
 # Solve the problem
 problem.solve(solver=C.CVXOPT)
 
-print("Problem exited with status: {0} and value attained: {1}".format(problem.status, problem.value))
+print("Problem exited with status: {0} and value attained: {1}".format(problem.status, round(problem.value, 5)))
 
 # Plotting section
 label_flag = [False, False]

### 2017-11-10 Programming-Assignment-2/dual.py
PA-2 Update 5

@@ -42,6 +42,24 @@ for i in range(0, m):
 problem = C.Problem(C.Maximize(objective), constraints)
 
 # Solve the problem
-problem.solve(solver=C.CVXOPT)
+problem.solve(solver=C.ECOS, abstol=1e-10, reltol=1e-09, feastol=1e-10, max_iters=1000)
 
 print("Problem exited with status: {0} and value attained: {1}".format(problem.status, round(problem.value, 5)))
+
+# Saving the \alphas into a text file
+alpha_file = open('alpha.txt', 'w')
+for i in range(0, m):
+    alpha_file.write('{0}\n'.format(round(alpha[i].value, 8)))
+alpha_file.close()
+
+# Plotting the comparison of \alpha vs y(<\beta, x> + \beta_0)
+alpha_vals = np.around(np.genfromtxt('alpha.txt'), decimals=7)
+beta_vals = np.around(np.genfromtxt('beta.txt'), decimals=7)
+
+for i in range(0, m):
+    plt.plot(alpha_vals[i], beta_vals[i], 'b.')
+
+plt.xlim((-0.005, 1.005))
+plt.xlabel('$\\alpha_{i}$')
+plt.ylabel('$y_{i}(\\beta_{i}^{T}x_{i} + \\beta_{0})$')
+plt.show()

### 2017-11-10 Programming-Assignment-2/primal.py

@@ -38,13 +38,17 @@ for i in range(0, m):
 problem = C.Problem(C.Minimize(objective), constraints)
 
 # Solve the problem
-problem.solve(solver=C.CVXOPT)
+problem.solve(solver=C.ECOS, abstol=1e-10, reltol=1e-09, feastol=1e-10, max_iters=1000)
 
 print("Problem exited with status: {0} and value attained: {1}".format(problem.status, round(problem.value, 5)))
+
+# Saving the y(<beta, x> + beta_0) into a text file
+beta_file = open('beta.txt', 'w')
 
 # Plotting section
 label_flag = [False, False]
 for i in range(0, m):
+    beta_file.write('{0}\n'.format(round(y[i]*(np.dot(np.array(beta.value).reshape(-1), X[i]) + beta_0.value), 8)))
     if y[i] == 1:
         point_type = 'ro' if (slackvar[i].value) > 1e-07 else 'go'
 
@@ -65,8 +69,8 @@ for i in range(0, m):
 x = np.arange(np.amin(X[:,0]), np.amax(X[:,0]), 0.001)
 beta = np.array(beta.value).reshape(-1).tolist()
 beta_0 = beta_0.value
-y = -beta_0/beta[1] - beta[0]*x/beta[1]
-plt.plot(x, y, 'k-')
+y_decbound = -beta_0/beta[1] - beta[0]*x/beta[1]
+plt.plot(x, y_decbound, 'k-')
 
 y_margin1 = (1 - beta_0 - beta[0]*x)/beta[1]
 plt.plot(x, y_margin1, 'r--')
@@ -76,3 +80,14 @@ plt.plot(x, y_margin2, 'r--')
 print("Margin width: {0}".format(2/np.sqrt(beta[0]**2 + beta[1]**2)))
 plt.legend(loc='upper right', numpoints=1)
 plt.show()
+
+# Calculating prediction error
+total_error = 0
+for i in range(0, m):
+    pred = np.sign((np.dot(beta, X[i]) + beta_0))
+    print("Actual: {0}\tPredicted: {1}".format(y[i], pred))
+    if pred == 1 and y[i] == -1:
+        total_error += opt.C2
+    elif pred == -1 and y[i] == 1:
+        total_error += opt.C1
+print("Total Weighted Classification Error for C1 = {0} and C2 = {1} is {2}".format(opt.C1, opt.C2, total_error))

### 2017-11-10 Programming-Assignment-2/PA_2-report.tex
PA-2 Update 6

@@
 \usepackage{float}
+\usepackage{bookmark}
 \usepackage{hyperref}
+\newcommand{\xtilde}{\widetilde{X}}
+\newcommand{\indi}{\mathbb{I}}
+\newcommand{\setS}{\mathcal{S}}
+\newcommand{\real}{\mathbb{R}}
+\newcommand{\lagL}{\mathcal{L}}
+\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
+\newcommand{\minimizewrt}[1]{\underset{#1}{\minimize}}
 
@@
 
+\section*{Instructions for the code}
+\begin{itemize}
+\item There are two files: \texttt{primal.py} and \texttt{dual.py}. \texttt{primal.py} concerns questions b.1, b.2 and b.4. \texttt{dual.py} concerns question b.3.
+\item The programs used \texttt{CVXPY}, a convex programming module in Python.
+\item To run these programs, enter \texttt{python3 <file>.py --C1 <C1-value> --C2 <C2-value>}. \texttt{<file>} can be any one of \texttt{primal} and \texttt{dual}.
+\end{itemize}
+
+\section*{Part a}
+\subsection*{Question 1}
+\begin{flushleft}
+Strong duality holds for the below problem:
+\begin{multline}
+\minimizewrt{\beta \in \real^{p}, \hspace{0.5mm} \beta_{0} \in \real, \hspace{0.5mm} \xi \in \real^{n}} \frac{1}{2}||\beta||_{2}^{2} + C_{1}\sum_{i \in \setS_{1}} \xi_{i} + C_{2}\sum_{i \in \setS_{2}} \xi_{i} \\
+\text{subject to } \xi_{i} \geq 0, \hspace{2mm} y_{i}(x_{i}^{T}\beta + \beta_{0}) \geq 1 - \xi_{i}, \hspace{2mm} \forall i \in \{1, 2, \ldots, n\}
+\label{primal}
+\end{multline}
+
+The reasons are:
+\begin{itemize}
+\item The objective function is convex
+\item The constraints are affine in the variables to be optimized over
+\item Using the above two points and Slater's condition, we can tell \ref{primal} has strong duality.
+\end{itemize}
+\end{flushleft}
+
+\subsection*{Question 2}
+\begin{flushleft}
+Define the Lagrangian \(\mathcal{L}\) as:
+\begin{equation}
+\label{lagrangian}
+\displaystyle \lagL(\beta, \beta_{0}, \xi, \alpha, \mu) = \frac{1}{2}||\beta||_{2}^{2} + C_{1}\sum_{i \in \setS_{1}}\xi_{i} + C_{2}\sum_{i \in \setS_{2}}\xi_{i} + \sum_{i=1}^{n}\alpha_{i}(1 - \xi_{i} - y_{i}(x_{i}^{T}\beta + \beta_{0})) - \sum_{i=1}^{n}\mu_{i}(\xi_{i})
+\end{equation}
+
+By property of stationarity in \ref{lagrangian}, assuming \(\beta, \beta_{0}, \xi\) are solutions to the primal and \(\alpha, \mu\) are the solutions to the dual and using the stationarity sub-condition of the KKT conditions, we get:
+\begin{gather}
+\label{stationarity-beta}
+\displaystyle \nabla_{\beta} \lagL = \beta + \sum_{i=1}^{n}\alpha_{i}(-y_{i}x_{i}) = 0 \implies \beta = \sum_{i=1}^{n}\alpha_{i}y_{i}x_{i} \\
+\label{stationarity-beta-0}
+\displaystyle \frac{\partial \lagL}{\partial \beta_{0}} = -\sum_{i}^{n}\alpha_{i}y_{i} = 0 \implies \sum_{i}^{n}\alpha_{i}y_{i} = 0 \\
+\label{stationarity-xi}
+\displaystyle \frac{\partial \lagL}{\partial \xi_{i}} = C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} - \alpha_{i} - \mu_{i} = 0 \implies \alpha_{i} = C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} - \mu_{i}
+\end{gather}
+
+Now using the complementary slackness sub-condition of the KKT conditions, we get:
+\begin{gather}
+\label{compl-slack-alpha}
+\displaystyle \alpha_{i}(1 - \xi_{i} + y_{i}(x_{i}^{T}\beta + \beta_{0})) = 0 \hspace{4mm} \forall i \in \{1, 2, \ldots, n\} \\
+\label{compl-slack-mu}
+\displaystyle \mu_{i}\xi_{i} = 0 \hspace{4mm} \forall i \in \{1, 2, \ldots, n\}
+\end{gather}
+
+There is nothing specifically required to be written for the primal feasibility sub-condition, since the solution to the primal has to be naturally feasible. But using dual feasibility:
+\begin{equation}
+\label{dual-fease}
+\alpha_{i}, \mu_{i} \geq 0 \hspace{4mm} i \in \{1, 2, \ldots, n\}
+\end{equation}
+\end{flushleft}
+
+\subsection*{Question 3}
+\begin{flushleft}
+Note that \ref{stationarity-beta-0} can be written as \(y^{T}\alpha = 0\) by definition of the inner product. Similarly, construct a matrix:
+\begin{equation}
+\label{y-diag}
+\mathop{diag}(y) = \begin{bmatrix} y_{1} & 0 & \ldots & 0 \\ 0 & y_{2} & & \vdots \\ \vdots & & \ddots & 0 \\ 0 & \ldots & 0 & y_{n} \end{bmatrix} 
+\end{equation}
+
+Using \ref{y-diag}, we can note that:
+\begin{equation}
+\label{x-tilde}
+\xtilde = \mathop{diag}(y)X = \begin{bmatrix} y_{1}x_{1} \\ \vdots \\ y_{n}x_{n} \end{bmatrix}
+\end{equation}
+
+From \ref{stationarity-beta}, we can rewrite \(\beta\) as:
+\begin{equation}
+\label{beta-x-tilde}
+\beta = \sum_{i=1}^{n} \alpha_{i}\xtilde_{i} = \xtilde^{T}\alpha \implies ||\beta||_{2}^{2} = \beta^{T}\beta = \alpha^{T}\xtilde\xtilde^{T}\alpha
+\end{equation}
+
+Using \ref{dual-fease} and \ref{stationarity-xi}, we can tell that:
+\begin{equation}
+\label{alpha-S}
+0 \leq \alpha_{i} \leq C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} \implies 0 \leq \alpha_{\setS_{j}} \leq C_{j}\mathbf{1} \hspace{2mm} \forall j \in \{1, 2\}
+\end{equation}
+
+The above step in \ref{alpha-S} uses the fact that if \(i \in \setS_{1}\), then \(i \notin \setS_{2}\). \(\alpha_{\setS_{j}}\) denotes the vector of \(\alpha_{i}\)s such that \(i \in \setS_{j}\).
+
+\end{flushleft}
+
 \section*{Part b}
@@
 \begin{flushleft}
-Below is the table representing the objective value of the optimal solution. The program used \texttt{CVXPY}, a convex programming module in Python.
+Below is the table representing the objective value of the optimal solution.
 \begin{center}
@@
 
+\subsection*{Question 4}
+The total classification error calculated in a zero-one loss manner is below:
+\begin{center}
+\begin{tabular}{|c|c|}
+\hline
+Penalty Parameter combination \((C_{1}, C_{2})\) & Weighted Classification Error \\
+\hline
+\hline
+\((1, 1)\) & \(3\) \\
+\hline
+\((1, 10)\) & \(16\) \\
+\hline
+\((10, 1)\) & \(4\) \\
+\hline
+\end{tabular}
+\end{center}
+
 \end{document}

### 2017-11-10 Programming-Assignment-2/PA_2-report.tex
PA-2 Update 7

@@
 \newcommand{\minimize}{\mathop{\mathrm{minimize}}}
+\newcommand{\maximize}{\mathop{\mathrm{maximize}}}
 \newcommand{\minimizewrt}[1]{\underset{#1}{\minimize}}   
+\newcommand{\maximizewrt}[1]{\underset{#1}{\maximize}}
 
@@
 \begin{multline}
+\label{primal}
 \minimizewrt{\beta \in \real^{p}, \hspace{0.5mm} \beta_{0} \in \real, \hspace{0.5mm} \xi \in \real^{n}} \frac{1}{2}||\beta||_{2}^{2} + C_{1}\sum_{i \in \setS_{1}} \xi_{i} + C_{2}\sum_{i \in \setS_{2}} \xi_{i} \\
 \text{subject to } \xi_{i} \geq 0, \hspace{2mm} y_{i}(x_{i}^{T}\beta + \beta_{0}) \geq 1 - \xi_{i}, \hspace{2mm} \forall i \in \{1, 2, \ldots, n\}
-\label{primal}
 \end{multline}
@@
 \begin{itemize}
-\item The objective function is convex
-\item The constraints are affine in the variables to be optimized over
-\item Using the above two points and Slater's condition, we can tell \ref{primal} has strong duality.
+\item The objective function is convex.
+\item The constraints are affine in the variables to be optimized over.
+\item \(\xi = \mathbf{1}, \beta = \mathbf{0}, \beta_{0} = 0\) is a point of strict feasibility in constraints.
+\item Using the above three points and Slater's condition, we can tell \ref{primal} has strong duality.
 \end{itemize}
@@
 \label{stationarity-beta}
-\displaystyle \nabla_{\beta} \lagL = \beta + \sum_{i=1}^{n}\alpha_{i}(-y_{i}x_{i}) = 0 \implies \beta = \sum_{i=1}^{n}\alpha_{i}y_{i}x_{i} \\
+\displaystyle \nabla_{\beta} \lagL = \beta + \sum_{i=1}^{n}\alpha_{i}(-y_{i}x_{i}) = \mathbf{0} \implies \beta = \sum_{i=1}^{n}\alpha_{i}y_{i}x_{i} \\
 \label{stationarity-beta-0}
-\displaystyle \frac{\partial \lagL}{\partial \beta_{0}} = -\sum_{i}^{n}\alpha_{i}y_{i} = 0 \implies \sum_{i}^{n}\alpha_{i}y_{i} = 0 \\
+\displaystyle \frac{\partial \lagL}{\partial \beta_{0}} = -\sum_{i=1}^{n}\alpha_{i}y_{i} = 0 \implies \sum_{i=1}^{n}\alpha_{i}y_{i} = 0 \\
 \label{stationarity-xi}
-\displaystyle \frac{\partial \lagL}{\partial \xi_{i}} = C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} - \alpha_{i} - \mu_{i} = 0 \implies \alpha_{i} = C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} - \mu_{i}
+\displaystyle \frac{\partial \lagL}{\partial \xi_{i}} = C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} - \alpha_{i} - \mu_{i} = 0 \implies \alpha_{i} + \mu_{i} = C_{j} \hspace{1.5mm} \forall i \in \setS_{j} \hspace{1.5mm} \forall j \in \{1, 2\}
 \end{gather}
@@
 \begin{equation}
-\label{alpha-S}
-0 \leq \alpha_{i} \leq C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} \implies 0 \leq \alpha_{\setS_{j}} \leq C_{j}\mathbf{1} \hspace{2mm} \forall j \in \{1, 2\}
+\label{dual-constraints}
+0 \leq \alpha_{i} \leq C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} \implies \mathbf{0} \leq \alpha_{\setS_{j}} \leq C_{j}\mathbf{1} \hspace{2mm} \forall j \in \{1, 2\}
 \end{equation}
 
-The above step in \ref{alpha-S} uses the fact that if \(i \in \setS_{1}\), then \(i \notin \setS_{2}\). \(\alpha_{\setS_{j}}\) denotes the vector of \(\alpha_{i}\)s such that \(i \in \setS_{j}\).
+The above step in \ref{dual-constraints} uses the fact that if \(i \in \setS_{1}\), then \(i \notin \setS_{2}\).
+\(\alpha_{\setS_{j}}\) denotes the vector of \(\alpha_{i}\)s such that \(i \in \setS_{j}\).
+
+Now substituting \ref{beta-x-tilde}, \ref{stationarity-beta}, \ref{stationarity-beta-0} and \ref{stationarity-xi} in \ref{lagrangian}, we get:
+\begin{multline}
+\displaystyle \lagL = \frac{1}{2}\alpha^{T}\xtilde\xtilde^{T}\alpha + \sum_{i\in\setS_{1}} (C_{1} - \alpha_{i} - \mu_{i})\xi_{i} + \sum_{i\in\setS_{2}} (C_{2} - \alpha_{i} - \mu_{i})\xi_{i} + \\\sum_{i=1}^{n} \alpha_{i} - \sum_{i=1}^{n}\sum_{j=1}^{n}\alpha_{i}\alpha_{j} y_{i}y_{j} x_{i}^{T}x_{j} - \beta_{0}\sum_{i=1}^{n}\alpha_{i}y_{i} \hspace{4mm} \text{ (Using \ref{beta-x-tilde} and \ref{stationarity-beta}) }
+\end{multline}
+\begin{gather}
+\displaystyle \implies \lagL = \frac{1}{2}\alpha^{T}\xtilde\xtilde^{T}\alpha + \sum_{i=1}^{n} \alpha_{i} - ||\beta||_{2}^{2} \hspace{4mm} \text{ (Using \ref{stationarity-beta-0}, \ref{stationarity-beta} and \ref{stationarity-xi}) } \\
+\label{dual-objective}
+\displaystyle \implies \lagL = -\frac{1}{2}\alpha^{T}\xtilde\xtilde^{T}\alpha + \sum_{i=1}^{n} \alpha_{i} = -\frac{1}{2}\alpha^{T}\xtilde\xtilde\alpha + \alpha^{T}\mathbf{1} \hspace{4mm} \text{ (Using \ref{beta-x-tilde}) }
+\end{gather}
+
+Now since the primal problem \ref{primal} was a minimization problem, the dual will be a maximization problem.
+Using \ref{dual-objective}, \ref{dual-constraints}, we formulate the dual problem is below:
+\begin{equation}
+\label{dual}
+\maximizewrt{\alpha \in \real^{n}} -\frac{1}{2}\alpha^{T}\xtilde\xtilde\alpha + \alpha^{T}\mathbf{1} \hspace{5mm}
+\text{subject to } y^{T}\alpha = 0, \hspace{2mm} \mathbf{0} \leq \alpha_{\setS_{j}} \leq C_{j}\mathbf{1} \hspace{2mm} \forall j \in \{1, 2\}
+\end{equation}
 

### 2017-11-10 Programming-Assignment-2/dual.py

@@ -59,7 +59,7 @@ beta_vals = np.around(np.genfromtxt('beta.txt'), decimals=7)
 for i in range(0, m):
     plt.plot(alpha_vals[i], beta_vals[i], 'b.')
 
-plt.xlim((-0.005, 1.005))
+plt.xlim((-0.01, 1.01))
 plt.xlabel('$\\alpha_{i}$')
 plt.ylabel('$y_{i}(\\beta_{i}^{T}x_{i} + \\beta_{0})$')
 plt.show()

### 2017-11-10 Programming-Assignment-2/primal.py

@@ -50,7 +50,7 @@ label_flag = [False, False]
 for i in range(0, m):
     beta_file.write('{0}\n'.format(round(y[i]*(np.dot(np.array(beta.value).reshape(-1), X[i]) + beta_0.value), 8)))
     if y[i] == 1:
-        point_type = 'ro' if (slackvar[i].value) > 1e-07 else 'go'
+        point_type = 'ro' if (slackvar[i].value) > 1e-06 else 'go'
 
         if label_flag[0] == False and point_type == 'go':
             plt.plot(X[i,0], X[i,1], point_type, label='+1')
@@ -58,7 +58,7 @@ for i in range(0, m):
         else:
             plt.plot(X[i,0], X[i,1], point_type)
     else:
-        point_type = 'ro' if (slackvar[i].value) > 1e-07 else 'bo'
+        point_type = 'ro' if (slackvar[i].value) > 1e-06 else 'bo'
 
         if label_flag[1] == False and point_type == 'bo':        
             plt.plot(X[i,0], X[i,1], point_type, label='-1')

### 2017-11-10 Programming-Assignment-2/PA_2-report.tex
PA-2 Update 8

@@
 \item To run these programs, enter \texttt{python3 <file>.py --C1 <C1-value> --C2 <C2-value>}. \texttt{<file>} can be any one of \texttt{primal} and \texttt{dual}.
+\item The output of \texttt{dual.py} requires the output of \texttt{primal.py} for the same parameters \texttt{C1} and \texttt{C2}.
+So please run \texttt{primal.py} first, and then \texttt{dual.py}.
+\texttt{primal.py} writes a \texttt{beta.txt} file that is used by \texttt{dual.py}, which also uses a \texttt{alpha.txt} file.
 \end{itemize}
@@
 
-There is nothing specifically required to be written for the primal feasibility sub-condition, since the solution to the primal has to be naturally feasible. But using dual feasibility:
+Using primal feasibility:
+\begin{equation}
+\label{primal-fease}
+\xi_{i}, y_{i}(\beta^{T}x_{i} + \beta_{0}) \geq 0 \hspace{4mm} \forall i \in \{1, 2, \ldots, n\}
+\end{equation}
+
+Using dual feasibility:
 \begin{equation}
 \label{dual-fease}
-\alpha_{i}, \mu_{i} \geq 0 \hspace{4mm} i \in \{1, 2, \ldots, n\}
+\alpha_{i}, \mu_{i} \geq 0 \hspace{4mm} \forall i \in \{1, 2, \ldots, n\}
 \end{equation}
@@
 \end{equation}
+\end{flushleft}
 
-\end{flushleft}
+\subsection*{Question 4}
+\ref{primal} is a Quadratic Programming problem and a Second Order Cone Program, whereas the dual formulation \ref{dual} is a Quadratic program.
 
@@
 
+\subsection*{Question 3}
+\begin{itemize}
+\item If \(\alpha_{i} = 0\), then the points \(x_{i}\) don't contribute to the boundary of the margin, since we know from \ref{stationarity-beta} that \(\beta\) is dependent on \(\alpha_{i}\)s.
+From observation of the graph generated, this also seems to mean that these refer to the points being correctly classified.
+\item if \(\alpha_{i} \in (0, C_{k}), i \in \setS_{k}\), then the \(\mu_{i} > 0 \implies \xi_{i} = 0\) from \ref{compl-slack-mu}.
+This means that these are \(x_{i}\)s lying on the margins of the boundaries.
+\item if \(\alpha_{i} = C_{k}, i \in \setS_{k}\), then \(\mu_{i} = 0 \implies \xi_{i} \geq 0\) from \ref{compl-slack-mu}.
+This means that these refer to points being incorrectly classified or points on the margin of the boundary.
+\end{itemize}
+
 \subsection*{Question 4}

### 2017-11-10 Programming-Assignment-2/dual.py

@@ -59,7 +59,8 @@ beta_vals = np.around(np.genfromtxt('beta.txt'), decimals=7)
 for i in range(0, m):
     plt.plot(alpha_vals[i], beta_vals[i], 'b.')
 
-plt.xlim((-0.01, 1.01))
-plt.xlabel('$\\alpha_{i}$')
-plt.ylabel('$y_{i}(\\beta_{i}^{T}x_{i} + \\beta_{0})$')
+C_max = max(opt.C1, opt.C2)
+plt.xlim((-C_max*0.01, C_max*(1.01)))
+plt.xlabel('$\\alpha_{i}$', fontsize=25)
+plt.ylabel('$y_{i}(\\beta^{T}x_{i} + \\beta_{0})$', fontsize=25)
 plt.show()

### 2017-11-10 Programming-Assignment-2/primal.py

@@ -46,23 +46,31 @@ print("Problem exited with status: {0} and value attained: {1}".format(problem.s
 beta_file = open('beta.txt', 'w')
 
 # Plotting section
-label_flag = [False, False]
+plt.ylabel('$x_{1}$', fontsize=20)
+plt.xlabel('$x_{2}$', fontsize=20)
+truths = [False, False, False]
 for i in range(0, m):
     beta_file.write('{0}\n'.format(round(y[i]*(np.dot(np.array(beta.value).reshape(-1), X[i]) + beta_0.value), 8)))
     if y[i] == 1:
-        point_type = 'ro' if (slackvar[i].value) > 1e-06 else 'go'
+        point_type = 'mo' if (slackvar[i].value) > 1e-06 else 'go'
 
-        if label_flag[0] == False and point_type == 'go':
+        if point_type == 'go' and truths[0] == False:
             plt.plot(X[i,0], X[i,1], point_type, label='+1')
-            label_flag[0] = True
+            truths[0] = True
+        elif point_type == 'mo' and truths[2] == False:
+            plt.plot(X[i,0], X[i,1], point_type, label='$\\xi_{i} > 0$')
+            truths[2] = True
         else:
             plt.plot(X[i,0], X[i,1], point_type)
     else:
-        point_type = 'ro' if (slackvar[i].value) > 1e-06 else 'bo'
+        point_type = 'mo' if (slackvar[i].value) > 1e-06 else 'bo'
 
-        if label_flag[1] == False and point_type == 'bo':        
+        if point_type == 'bo' and truths[1] == False:
             plt.plot(X[i,0], X[i,1], point_type, label='-1')
-            label_flag[1] = True
+            truths[1] = True
+        elif point_type == 'mo' and truths[2] == False:
+            plt.plot(X[i,0], X[i,1], point_type, label='$\\xi_{i} > 0$')
+            truths[2] = True
         else:
             plt.plot(X[i,0], X[i,1], point_type)
 
@@ -70,22 +78,22 @@ x = np.arange(np.amin(X[:,0]), np.amax(X[:,0]), 0.001)
 beta = np.array(beta.value).reshape(-1).tolist()
 beta_0 = beta_0.value
 y_decbound = -beta_0/beta[1] - beta[0]*x/beta[1]
-plt.plot(x, y_decbound, 'k-')
[…]

### 2017-11-20 Paper-Presentation/main.tex
PA-2 Update 1
Final Submission of slides, added footnote references

@@
 \documentclass[10pt]{beamer}
+\usepackage[style=verbose,backend=biber]{biblatex}
 \usepackage{amsmath}
@@
 \newcommand{\eh}{\epsilon_{h}}
+\newcommand{\citecomma}{\({}^{{}_{,}}\)}
 %\documentclass{article}
@@
 \usetheme{CambridgeUS}
+\addbibresource{main.bib}
+\renewcommand{\citesetup}{\tiny}
 
@@
         \item<2->{Condition number: \(\kappa = \frac{L}{\gamma} = \frac{\maxeig(\h F)}{\mineig(\h F)}\), where \(L\) is smoothness parameter, \(\gamma\) is strong convexity parameter}
-        \item<3->{Convergence criteria for gradient descent relies on \(\kappa\) (\cite{benrecht-simons})}
+        \item<3->{Convergence criteria for gradient descent relies on \(\kappa\) \footfullcite{benrecht-simons}}
     \end{itemize}
@@
     \begin{itemize}    
-        \item<5->{For gradient descent, we have 1 (learning rate) and for Adam (\cite{adam-iclr}) we have 4 (learning rate, tolerance, \(\beta_{1}\) and \(\beta_{2}\))}
+        \item<5->{For gradient descent, we have 1 (learning rate) and for Adam \footfullcite{adam-iclr} we have 4 (learning rate, tolerance, \(\beta_{1}\) and \(\beta_{2}\))}
     \end{itemize}
@@
     \begin{itemize}    
-        \item<7->{\emph{Non-convex} problems have saddle-points to escape from (\cite{du-nips}, \cite{jin-icml})}
+        \item<7->{\emph{Non-convex} problems have saddle-points to escape from \footfullcite{du-nips}\citecomma\footfullcite{jin-icml}}
     \end{itemize}
@@
 \pause
-Note: If the function \(F\) satisfies \underline{strict saddle property} (\cite{ge-colt}), then the above defined optimality condition ensures ``closeness'' to a local-minimum.
+Note: If the function \(F\) satisfies \underline{strict saddle property} \footfullcite{ge-colt}, then the above defined optimality condition ensures ``closeness'' to a local-minimum.
 \pause
@@
 \begin{itemize}
-\item<1->{Method used to optimize over \emph{non-convex} functions, proven to be extremely useful at or near saddle points, by leveraging the negative curvature around that point (\cite{conn-siam}, \cite{sorensen-siam})}
+\item<1->{Method used to optimize over \emph{non-convex} functions, proven to be extremely useful at or near saddle points, by leveraging the negative curvature around that point \footfullcite{conn-siam}\citecomma\footfullcite{sorensen-siam}}
 \item<2->{Helps achieve two things:
@@
 \item<2->{Similarities to other eigenvalue problems, leading to special methods drawing inferences from existing methods e.g., truncated conjugate gradient methods, Lanczos trust-region based methods}
-\item<3->{To reach an \((\eg, \eh)\)-optimal point as defined earlier, we need \(O\left(\max\{\eh^{-1}\eg^{-2}, \eh^{-3}\}\right)\) (\cite{cartis-joc})} 
-\item<4->{To reach an \((\eg, \eh)\)-optimal point, we need \(O\left(\max\{\eg^{-3}, \eh^{-3}\}\right)\) (\cite{geovani-oms})}
+\item<3->{To reach an \((\eg, \eh)\)-optimal point as defined earlier, we need \(O\left(\max\{\eh^{-1}\eg^{-2}, \eh^{-3}\}\right)\) \footfullcite{cartis-joc}} 
+\item<4->{To reach an \((\eg, \eh)\)-optimal point, we need \(O\left(\max\{\eg^{-3}, \eh^{-3}\}\right)\) \footfullcite{geovani-oms}}
 \end{itemize}
@@
 \item<2->{Methods to solve this include Lanczos type iterations, fast approximate matrix inversion}
-\item<3->{To reach an \((\eg, \eh)\)-optimal point as defined earlier, we need \(O\left(\max\{\eg^{-2}, \eh^{-3}\}\right)\) (\cite{cartis-joc})}
-\item<4->{Assuming more tight conditions helps improve the dependence on \(\eg\) - we need \(O\left(\max\{\eg^{-3/2}, \eh^{-3}\}\right)\) (\cite{cartis-siam})}
+\item<3->{To reach an \((\eg, \eh)\)-optimal point as defined earlier, we need \(O\left(\max\{\eg^{-2}, \eh^{-3}\}\right)\) \footfullcite{cartis-joc}}
+\item<4->{Assuming more tight conditions helps improve the dependence on \(\eg\) - we need \(O\left(\max\{\eg^{-3/2}, \eh^{-3}\}\right)\) \footfullcite{cartis-siam}}
 \end{itemize}
@@
 \begin{itemize}
-\item \(-m_{t}(\sbold_{t}) \geq -m_{t}(\sbold_{t}^{C}) \geq \min\lbrace\Phi_{1}(\g F(\xbold_{t}), H_{t}), \Psi_{1}(\g F(\xbold_{t}), \Delta_{t})\rbrace \)
+\item \(-m_{t}(\sbold_{t}) \geq -m_{t}(\sbold_{t}^{C}) \geq \min\lbrace\Phi_{1}(\g F(\xbold_{t}), H_{t}), \Psi_{1}(\g F(\xbold_{t}), \Delta_{t})\rbrace \)\footnote{Refer to Equation 8a of this paper}
 \item \(-m_{t}(\sbold_{t}) \geq -m_{t}(\sbold_{t}^{E}) \geq \frac{1}{2}\nu|\mineig(H_{t})|\Delta_{t}^{2}, \text{ if } \mineig(H_{t}) < 0 \)
@@
 \begin{itemize}
-\item \(-m_{t}(\sbold_{t}) \geq -m_{t}(\sbold_{t}^{C}) \geq \max\lbrace\Phi_{2}(\sigma_{t}, \sbold_{t}, B_{H}, \g F(\xbold_{t})), \Psi_{2}(\sigma_{t}, B_{H}, \g F(\xbold_{t}))\rbrace\)
-\item \(-m_{t}(\sbold_{t}) \geq -m_{t}(\sbold_{t}^{C}) \geq \frac{\nu|\mineig(H_{t})|}{6} \max\lbrace ||\sbold_{t}^{E}||^{2}, \frac{\nu^{2}|\mineig(H_{t})|^{2}}{\sigma_{t}^{2}}\rbrace \text{ if } \mineig(H_{t}) < 0\)
+\item \(-m_{t}(\sbold_{t}) \geq -m_{t}(\sbold_{t}^{C}) \geq \max\lbrace\Phi_{2}(\sigma_{t}, \sbold_{t}, B_{H}, \g F(\xbold_{t})), \Psi_{2}(\sigma_{t}, B_{H}, \g F(\xbold_{t}))\rbrace\)\footnote{Refer to Equation 11a of this paper}
+\item \(-m_{t}(\sbold_{t}) \geq -m_{t}(\sbold_{t}^{E}) \geq \frac{\nu|\mineig(H_{t})|}{6} \max\lbrace ||\sbold_{t}^{E}||^{2}, \frac{\nu^{2}|\mineig(H_{t})|^{2}}{\sigma_{t}^{2}}\rbrace \text{ if } \mineig(H_{t}) < 0\)
 \end{itemize}
@@
 \begin{itemize}
-\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon = \min\{\epsilon_{0}, \zeta \eg\}, \zeta \in (0, 0.5)\)}
+\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon = \min\{\epsilon_{0}\footnote{Refer to Equation 12 of this paper}, \zeta \eg\}, \zeta \in (0, 0.5)\)}
 \item<2->{\(\h F(\xbold)\) satisfies Hessian regularity}
@@
 \begin{itemize}
-\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon = \min\{\epsilon_{0}, \zeta\eg\}, \zeta \in (0, 0.5)\)}
+\item<1->{\(H(\xbold)\) satisfies Inexact Hessian Regularity with \(\epsilon = \min\{\epsilon_{0}\footnote{Refer to Equation 12 of this paper}, \zeta\eg\}, \zeta \in (0, 0.5)\)}
 \item<2->{\(\h F(\xbold)\) satisfies Hessian regularity}
@@
 
-\begin{frame}[t,allowframebreaks]
-\frametitle{References}
-\bibliographystyle{plain}
-\bibliography{main.bib}
-\end{frame}
+%\begin{frame}[t,allowframebreaks]
+%\frametitle{References}
+%\bibliography{main.bib}
+%\end{frame}
 \end{document}

### 2017-11-20 Project/MLPs/MNIST.py
Start MNIST code

@@ -1,0 +1,53 @@
+import json
+import torch as t
+import numpy as np
+import torch.nn as nn
+from torchvision import transforms
+from argparse import ArgumentParser
+from torchvision.datasets import MNIST
+from torch.utils.data import DataLoader
+
+np.random.seed(29)
+t.manual_seed(29)
+
+# Get the parameters
+p = ArgumentParser()
+p.add_argument('--opt', required=True, type=str, help='Optimizer --> adam | adagrad | rmsprop | nag | cm')
+p.add_argument('--opt_params', required=True, type=str, help='File containing params in json format')
+p.add_argument('--maxiter', default=int(5e04), type=int, help='Maximum iterations')
+p.add_argument('--architecture', required=True, type=str, help='CSV file with number of nodes per layer')
+p.add_argument('--activation', default='relu', type=str, help='Activation function to be used')
+p.add_argument('--dataroot', default='./', type=str, help='Data root folder')
+p.add_argument('--cuda', default=-1, type=int, help='Enter CUDA device (> 0), or -1 if no CUDA')
+p = p.parse_args()
+
+# CUDA check
+if p.cuda != -1:
+    t.cuda.set_device(p.cuda)
+
+# Load the data
+transformations = transforms.Compose([transforms.ToTensor(), 
+                                      transforms.Normalize((0.5,), (0.5,)), 
+                                      transforms.Lambda(lambda x: x.view(-1, 784))
+                                    ])
+tr_dset = MNIST(root=p.dataroot, train=True, transform=transformations, download=True)
+te_dset = MNIST(root=p.dataroot, train=False, transform=transformations, download=True)
+tr_d_loader = DataLoader(dataset=tr_dset, batch_size=64, shuffle=True)
+te_d_loader = DataLoader(dataset=te_dset, batch_size=1000, shuffle=True)
+
+# Build MLP architecture
+arch_vals = np.genfromtxt(p.architecture, delimiter=',').reshape(-1)
+activations = {'relu': nn.ReLU(True), 
+               'sigmoid': nn.Sigmoid(True),
+               'tanh': nn.Tanh(True),
+               'leaky_relu', nn.LeakyReLU(0.2, inplace=True)
+              }
+
+model = nn.Sequential()
+for i in range(0, arch_vals.shape[0] - 1):
[…]

### 2017-11-20 Project/MLPs/MNIST.py
Complete MNIST

@@ -6,6 +6,8 @@ from torchvision import transforms
 from argparse import ArgumentParser
 from torchvision.datasets import MNIST
 from torch.utils.data import DataLoader
+from torch.autograd import Variable as V
+from utils import return_model, return_optimizer
 
 np.random.seed(29)
 t.manual_seed(29)
@@ -28,26 +30,94 @@ if p.cuda != -1:
 # Load the data
 transformations = transforms.Compose([transforms.ToTensor(), 
                                       transforms.Normalize((0.5,), (0.5,)), 
-                                      transforms.Lambda(lambda x: x.view(-1, 784))
+                                      transforms.Lambda(lambda x: x.view(784))
                                     ])
 tr_dset = MNIST(root=p.dataroot, train=True, transform=transformations, download=True)
 te_dset = MNIST(root=p.dataroot, train=False, transform=transformations, download=True)
 tr_d_loader = DataLoader(dataset=tr_dset, batch_size=64, shuffle=True)
-te_d_loader = DataLoader(dataset=te_dset, batch_size=1000, shuffle=True)
+te_d_loader = DataLoader(dataset=te_dset, batch_size=5000, shuffle=True)
 
 # Build MLP architecture
-arch_vals = np.genfromtxt(p.architecture, delimiter=',').reshape(-1)
-activations = {'relu': nn.ReLU(True), 
-               'sigmoid': nn.Sigmoid(True),
-               'tanh': nn.Tanh(True),
-               'leaky_relu', nn.LeakyReLU(0.2, inplace=True)
-              }
-
-model = nn.Sequential()
-for i in range(0, arch_vals.shape[0] - 1):
-    model.add_module('linear_{0}-{1}-{2}'.format(i, arch_vals[i], arch_vals[i+1]), nn.Linear(arch_vals[i], arch_vals[i+1]))
-    if i != arch_vals.shape[0] - 2:
-        model.add_module('{0}_{1}'.format(p.activation, i), activations[p.activation])
+arch_vals = np.genfromtxt(p.architecture, delimiter=',').reshape(-1).astype(int).tolist()
+model = return_model(arch_vals, p.activation)
+print(model)
 
 if p.cuda != -1:
     model = model.cuda()
+
+# Build Optimizer
[…]

### 2017-11-20 Project/utils.py

@@ -1,0 +1,36 @@
+import torch as t
+import torch.nn as nn
+import torch.optim as optim
+
+def return_model(arch_vals, activation):
+    activations = {'relu': nn.ReLU(True), 
+                   'sigmoid': nn.Sigmoid(),
+                   'tanh': nn.Tanh(),
+                   'leaky_relu': nn.LeakyReLU(0.2, inplace=True)
+                  }
+    model = nn.Sequential()
+    for i in range(0, len(arch_vals) - 1):
+        model.add_module('linear_{0}-{1}-{2}'.format(i, arch_vals[i], arch_vals[i+1]), nn.Linear(arch_vals[i], arch_vals[i+1]))
+        if i != len(arch_vals) - 2:
+            model.add_module('{0}_{1}'.format(activation, i), activations[activation])
+    return model
+
+def return_optimizer(optimizer, opt_params, model_params):
+    if optimizer == 'adam':
+        opt_function = optim.Adam(params=model_params, lr=opt_params['lr'])
+    if optimizer == 'adagrad':
+        opt_function = optim.Adagrad(params=model_params, lr=opt_params['lr'])
+    if optimizer == 'rmsprop':
+        opt_function = optim.RMSprop(params=model_params, lr=opt_params['lr'])
+    if optimizer == 'nag':
+        opt_function = optim.SGD(params=model_params, lr=opt_params['lr'], momentum=opt_params['momentum'], nesterov=True)
+    if optimizer == 'cm':
+        opt_function = optim.SGD(params=model_params, lr=opt_params['lr'], momentum=opt_params['momentum'])
+
+    if 'betas' in opt_function.param_groups[0].keys():
+        opt_function.param_groups[0]['betas'] = opt_params['betas']
+    if 'alpha' in opt_function.param_groups[0].keys():
+        opt_function.param_groups[0]['alpha'] = opt_params['alpha']
+    if 'weight_decay' in opt_params.keys():
+        opt_function.param_groups[0]['weight_decay'] = opt_params['weight_decay']
+    return opt_function

### 2017-11-20 Project/Trivial Functions/functions.py
Added Trivial Optimization test functions

@@ -1,0 +1,60 @@
+import numpy as np
+
+# Index of references
+# B1 -> Bohachevsky Function 1
+# B2 -> Bohachevsky Function 2
+# B3 -> Bohachevsky Function 3
+# RB -> Rosenbrock Function
+# BL -> Beale's Function
+# ST -> Styblinsky-Tang's Function
+
+def B1(X):
+    """
+        Returns Bohachevsky function 1 at a point
+    """
+    x = X[0]
+    y = X[1]
+    return x**2 + 2.0*(y**2) - 0.3*np.cos(3.0*np.pi*x) - 0.4*np.cos(4.0*np.pi*y) + 0.7
+
+def B2(X):
+    """
+        Returns Bohachevsky function 2 at a point
+    """
+    x = X[0]
+    y = X[1]
+    return x**2 + 2.0*(y**2) - 0.3*np.cos(3.0*np.pi*x)*np.cos(4.0*np.pi*y) + 0.3
+
+def B3(X):
+    """
+        Returns Bohachevsky function 3 at a point
+    """
+    x = X[0]
+    y = X[1]
+    return x**2 + 2.0*(y**2) - 0.3*np.cos(3.0*np.pi*x + 4.0*np.pi*y) + 0.3
+
+def RB(X):
+    """
+        Returns Rosenbrock's function at a point
+    """
+    x = X[0]
+    y = X[1]
+    return 25*(y - x**2)**2 + (1 - x)**2
+
+def BL(X):
+    """
+        Returns Beale's function at a point
+    """
+    x = X[0]
+    y = X[1]
+    term1 = (1.5 - x + x*y)**2
+    term2 = (2.25 - x + x*y**2)**2
+    term3 = (2.625 - x + x*y**3)**2
+    return term1 + term2 + term3
+
+def ST(X):
+    """
+        Returns Styblinsky-Tang's function at a point
+    """
+    x = X[0]
+    y = X[1]
+    return x*4 - 16*(x**2) + 5*x + y**4 - 16*(y**2) + 5*y

### 2017-11-20 Project/Trivial Functions/gradients.py

@@ -1,0 +1,77 @@
+import numpy as np
+
+# Index of references
+# B1 -> Bohachevsky Function 1
+# B2 -> Bohachevsky Function 2
+# B3 -> Bohachevsky Function 3
+# RB -> Rosenbrock Function
+# BL -> Beale's Function
+# ST -> Styblinsky-Tang's Function
+
+def B0(X):
+    """
+        Returns gradient of Bohachevsky Function 1 at a point
+    """
+    x = X[0]
+    y = X[1]
+    dx = 2*x + 3.0*np.pi*0.3*np.sin(3.0*np.pi*x)
+    dy = 4*y + 4.0*np.pi*0.4*np.sin(4.0*np.pi*y)
+    return np.array([dx, dy])
+
+def B1(X):
+    """
+        Returns gradient of Bohachevsky Function 2 at a point
+    """
+    x = X[0]
+    y = X[1]
+    dx = 2*x + 3.0*np.pi*0.3*np.sin(3.0*np.pi*x)*np.cos(4.0*np.pi*y)
+    dy = 4*y + 4.0*np.pi*0.4*np.sin(4.0*np.pi*y)*np.cos(3.0*np.pi*x)
+    return np.array([dx, dy])
+
+def B2(X):
+    """
+        Returns gradient of Bohachevsky Function 3 at a point
+    """
+    x = X[0]
+    y = X[1]
+    dx = 2*x + 3.0*np.pi*0.3*np.sin(3.0*np.pi*x + 4.0*np.pi*y)
+    dy = 4*y + 4.0*np.pi*0.4*np.sin(3.0*np.pi*x + 4.0*np.pi*y)
+    return np.array([dx, dy])
+
+def RB(X):
+    """
+        Returns gradient of Rosenbrock's function at a point
+    """
+    x = X[0]
+    y = X[1]
+    dx = -100*(y - x**2)*x + 2*(x - 1)
+    dy = 50*(y - x**2)
+    return np.array([dx, dy])
+
+def BL(X):
+    """
+        Returns gradient of Beale's function at a point
+    """
+    x = X[0]
+    y = X[1]
+	cmn1	= 2*(1.5 - x + x*y)
+	cmn2	= 2*(2.25 - x + x*y**2)
+	cmn3	= 2*(2.625 - x + x*y**3)
+	tmx1	= -1 + y
[…]

### 2017-11-20 Project/Trivial Functions/optimizers.py
Added 3 optimizers

@@ -1,0 +1,51 @@
+import numpy as np
+
+def GD(t, X, G, lr):
+    """
+        Function to perform one GD step
+        Args:
+            t   => previous timestep
+            X   => Current point
+            G   => Gradient at current point
+            lr  => Learning Rate 
+    """
+    timestep = t + 1
+    update = X - lr*G
+    return timestep, update
+
+def CM(t, X, G, V, m, lr):
+    """
+        Function to perform one Classical Momentum step
+        Args:
+            t   => previous timestep
+            X   => Current point
+            G   => Gradient at current point
+            V   => Velocity at the previous iteration
+            m   => Momentum Parameter
+            lr  => Learning Rate
+    """
+    timestep = t + 1
+    velocity = m*V - lr*G
+    update = X + velocity
+    return timestep, update, velocity
+
+def Adam(t, X, G, FM, SM, betas, lr, eps=1e-08):
+    """
+        Function to perform one Adam step
+        Args:
+            t       => previous timestep
+            X       => Current point
+            G       => Gradient at current point
+            FM      => First Moment at the previous iteration
+            SM      => Second Moment at the previous iteration
+            betas   => Beta parameter defined in Adam
+            lr      => Learning Rate
+            eps     => (optional, default=1e-08) Prevent divide by 0 error
+    """
+    timestep = t + 1
+    FM = betas[0]*FM + (1 - betas[0])*G
+    SM = betas[1]*SM + (1 - betas[1])*G*G
+    mh = FM/(1 - betas[0]**timestep)
+    vh = SM/(1 - betas[1]**timestep)
+    update = X - lr*mh/(np.sqrt(vh) + eps)
+    return timestep, update, FM, SM

### 2017-11-21 Project/MLPs/MLPAutoencoder.py
rename Classifier, start Autoencoder

@@ -1,0 +1,41 @@
+import json
+import torch as t
+import numpy as np
+import torch.nn as nn
+from torchvision import transforms
+from argparse import ArgumentParser
+from torchvision.datasets import MNIST
+from torch.utils.data import DataLoader
+from torch.autograd import Variable as V
+from utils import return_model, return_optimizer
+
+np.random.seed(29)
+t.manual_seed(29)
+
+# Get the parameters
+p = ArgumentParser()
+p.add_argument('--opt', required=True, type=str, help='Optimizer --> adam | adagrad | rmsprop | nag | cm')
+p.add_argument('--opt_params', required=True, type=str, help='File containing params in json format')
+p.add_argument('--maxiter', default=int(5e04), type=int, help='Maximum iterations')
+p.add_argument('--architecture', required=True, type=str, help='CSV file with number of nodes per layer')
+p.add_argument('--activation', default='relu', type=str, help='Activation function to be used')
+p.add_argument('--dataset', required=True, type=str, help='Dataset to use --> mnist | faces')
+p.add_argument('--dataroot', default='./', type=str, help='Data root folder')
+p.add_argument('--cuda', default=-1, type=int, help='Enter CUDA device (> 0), or -1 if no CUDA')
+p = p.parse_args()
+
+# CUDA check
+if p.cuda != -1:
+    t.cuda.set_device(p.cuda)
+
+# Load the data
+if p.dataset == 'mnist':
+    transformations = transforms.Compose([transforms.ToTensor(), 
+                                          transforms.Normalize((0.5,), (0.5,)), 
+                                          transforms.Lambda(lambda x: x.view(784))
+                                        ])
+    tr_dset = MNIST(root=p.dataroot, train=True, transform=transformations, download=True)
+    te_dset = MNIST(root=p.dataroot, train=False, transform=transformations, download=True)
+
+elif p.dataset == 'faces':
+    # Do something

### 2017-11-21 Project/MLPs/MLPClassifier.py

@@ -4,10 +4,10 @@ import numpy as np
 import torch.nn as nn
 from torchvision import transforms
 from argparse import ArgumentParser
-from torchvision.datasets import MNIST
 from torch.utils.data import DataLoader
 from torch.autograd import Variable as V
 from utils import return_model, return_optimizer
+from torchvision.datasets import MNIST, CIFAR10, SVHN
 
 np.random.seed(29)
 t.manual_seed(29)
@@ -19,6 +19,7 @@ p.add_argument('--opt_params', required=True, type=str, help='File containing pa
 p.add_argument('--maxiter', default=int(5e04), type=int, help='Maximum iterations')
 p.add_argument('--architecture', required=True, type=str, help='CSV file with number of nodes per layer')
 p.add_argument('--activation', default='relu', type=str, help='Activation function to be used')
+p.add_argument('--dataset', required=True, type=str, help='Dataset to use --> mnist | cifar10 | svhn')
 p.add_argument('--dataroot', default='./', type=str, help='Data root folder')
 p.add_argument('--cuda', default=-1, type=int, help='Enter CUDA device (> 0), or -1 if no CUDA')
 p = p.parse_args()
@@ -28,12 +29,30 @@ if p.cuda != -1:
     t.cuda.set_device(p.cuda)
 
 # Load the data
-transformations = transforms.Compose([transforms.ToTensor(), 
-                                      transforms.Normalize((0.5,), (0.5,)), 
-                                      transforms.Lambda(lambda x: x.view(784))
-                                    ])
-tr_dset = MNIST(root=p.dataroot, train=True, transform=transformations, download=True)
-te_dset = MNIST(root=p.dataroot, train=False, transform=transformations, download=True)
+if p.dataset == 'mnist':
+    transformations = transforms.Compose([transforms.ToTensor(), 
+                                          transforms.Normalize((0.5,), (0.5,)), 
+                                          transforms.Lambda(lambda x: x.view(784))
+                                        ])
+    tr_dset = MNIST(root=p.dataroot, train=True, transform=transformations, download=True)
+    te_dset = MNIST(root=p.dataroot, train=False, transform=transformations, download=True)
+
+elif p.dataset == 'cifar10':
+    transformations = transforms.Compose([transforms.ToTensor(), 
+                                          transforms.Normalize((0.5, 0.5, 0.5), (0.5, 0.5, 0.5)), 
+                                          transforms.Lambda(lambda x: x.view(3072))
[…]

### 2017-11-21 Project/utils.py

@@ -2,7 +2,7 @@ import torch as t
 import torch.nn as nn
 import torch.optim as optim
 
-def return_model(arch_vals, activation):
+def return_model(arch_vals, activation, init='random'):
     activations = {'relu': nn.ReLU(True), 
                    'sigmoid': nn.Sigmoid(),
                    'tanh': nn.Tanh(),
@@ -13,6 +13,15 @@ def return_model(arch_vals, activation):
         model.add_module('linear_{0}-{1}-{2}'.format(i, arch_vals[i], arch_vals[i+1]), nn.Linear(arch_vals[i], arch_vals[i+1]))
         if i != len(arch_vals) - 2:
             model.add_module('{0}_{1}'.format(activation, i), activations[activation])
+
+    if init == 'xavier':
+        for p in model.parameters():
+            if len(p.size()) >= 2:
+                p.data = nn.init.xavier_uniform(p)
+    elif init == 'he':
+        for p in model.parameters():
+            if len(p.size()) >= 2:
+                p.data = nn.init.kaiming_uniform(p)
     return model
 
 def return_optimizer(optimizer, opt_params, model_params):
@@ -20,6 +29,8 @@ def return_optimizer(optimizer, opt_params, model_params):
         opt_function = optim.Adam(params=model_params, lr=opt_params['lr'])
     if optimizer == 'adagrad':
         opt_function = optim.Adagrad(params=model_params, lr=opt_params['lr'])
+    if optimizer == 'adadelta':
+        opt_function = optim.Adadelta(params=model_params, lr=opt_params['lr'])
     if optimizer == 'rmsprop':
         opt_function = optim.RMSprop(params=model_params, lr=opt_params['lr'])
     if optimizer == 'nag':

### 2017-11-21 Project/MLPs/MLPAutoencoder.py
Complete MNIST autoencoder

@@ -4,6 +4,7 @@ import numpy as np
 import torch.nn as nn
 from torchvision import transforms
 from argparse import ArgumentParser
+from torchvision import utils as tv_u
 from torchvision.datasets import MNIST
 from torch.utils.data import DataLoader
 from torch.autograd import Variable as V
@@ -19,7 +20,8 @@ p.add_argument('--opt_params', required=True, type=str, help='File containing pa
 p.add_argument('--maxiter', default=int(5e04), type=int, help='Maximum iterations')
 p.add_argument('--architecture', required=True, type=str, help='CSV file with number of nodes per layer')
 p.add_argument('--activation', default='relu', type=str, help='Activation function to be used')
-p.add_argument('--dataset', required=True, type=str, help='Dataset to use --> mnist | faces')
+p.add_argument('--init', default='random', type=str, help='Initialization scheme to use --> random | he | xavier')
+p.add_argument('--dataset', required=True, type=str, help='Dataset to use --> mnist')
 p.add_argument('--dataroot', default='./', type=str, help='Data root folder')
 p.add_argument('--cuda', default=-1, type=int, help='Enter CUDA device (> 0), or -1 if no CUDA')
 p = p.parse_args()
@@ -31,11 +33,75 @@ if p.cuda != -1:
 # Load the data
 if p.dataset == 'mnist':
     transformations = transforms.Compose([transforms.ToTensor(), 
-                                          transforms.Normalize((0.5,), (0.5,)), 
                                           transforms.Lambda(lambda x: x.view(784))
                                         ])
     tr_dset = MNIST(root=p.dataroot, train=True, transform=transformations, download=True)
     te_dset = MNIST(root=p.dataroot, train=False, transform=transformations, download=True)
 
-elif p.dataset == 'faces':
-    # Do something
+else:
+    raise ValueError("No other dataset available")
+
+tr_d_loader = DataLoader(dataset=tr_dset, batch_size=64, shuffle=True)
+te_d_loader = DataLoader(dataset=te_dset, batch_size=5000, shuffle=True)
+
+# Build MLP architecture
+arch_vals = np.genfromtxt(p.architecture, delimiter=',').reshape(-1).astype(int).tolist()
[…]

### 2017-11-21 Project/MLPs/MLPClassifier.py

@@ -19,6 +19,7 @@ p.add_argument('--opt_params', required=True, type=str, help='File containing pa
 p.add_argument('--maxiter', default=int(5e04), type=int, help='Maximum iterations')
 p.add_argument('--architecture', required=True, type=str, help='CSV file with number of nodes per layer')
 p.add_argument('--activation', default='relu', type=str, help='Activation function to be used')
+p.add_argument('--init', default='random', type=str, help='Initialization scheme to use --> random | he | xavier')
 p.add_argument('--dataset', required=True, type=str, help='Dataset to use --> mnist | cifar10 | svhn')
 p.add_argument('--dataroot', default='./', type=str, help='Data root folder')
 p.add_argument('--cuda', default=-1, type=int, help='Enter CUDA device (> 0), or -1 if no CUDA')
@@ -50,15 +51,15 @@ elif p.dataset == 'svhn':
                                           transforms.Normalize((0.5, 0.5, 0.5), (0.5, 0.5, 0.5)), 
                                           transforms.Lambda(lambda x: x.view(3072))
                                         ])
-    tr_dset = SVHN(root=p.dataroot, train=True, transform=transformations, download=True)
-    te_dset = SVHN(root=p.dataroot, train=False, transform=transformations, download=True)
+    tr_dset = SVHN(root=p.dataroot, split='train', transform=transformations, download=True)
+    te_dset = SVHN(root=p.dataroot, split='test', transform=transformations, download=True)
                                     
 tr_d_loader = DataLoader(dataset=tr_dset, batch_size=64, shuffle=True)
 te_d_loader = DataLoader(dataset=te_dset, batch_size=5000, shuffle=True)
 
 # Build MLP architecture
 arch_vals = np.genfromtxt(p.architecture, delimiter=',').reshape(-1).astype(int).tolist()
-model = return_model(arch_vals, p.activation)
+model = return_model(arch_vals, p.activation, init=p.init)
 print(model)
 
 if p.cuda != -1:
@@ -74,13 +75,13 @@ if p.cuda != -1:
     loss_fn = loss_fn.cuda()
 
 optimizer_params = json.load(open(p.opt_params))
-params = [p.opt]
+params = [p.dataset, p.opt]
 for k in sorted(list(optimizer_params)):
     params.append(optimizer_params[k])
 for k in arch_vals:
-    params.append(arch_vals)
+    params.append(k)
 
[…]

### 2017-11-21 Project/utils.py

@@ -2,7 +2,7 @@ import torch as t
 import torch.nn as nn
 import torch.optim as optim
 
-def return_model(arch_vals, activation, init='random'):
+def return_model(arch_vals, activation, init, last_layer_active=False):
     activations = {'relu': nn.ReLU(True), 
                    'sigmoid': nn.Sigmoid(),
                    'tanh': nn.Tanh(),
@@ -13,15 +13,17 @@ def return_model(arch_vals, activation, init='random'):
         model.add_module('linear_{0}-{1}-{2}'.format(i, arch_vals[i], arch_vals[i+1]), nn.Linear(arch_vals[i], arch_vals[i+1]))
         if i != len(arch_vals) - 2:
             model.add_module('{0}_{1}'.format(activation, i), activations[activation])
+    if last_layer_active:
+        model.add_module('{0}_{1}'.format(activation, i+1), activations[activation])
 
     if init == 'xavier':
         for p in model.parameters():
             if len(p.size()) >= 2:
-                p.data = nn.init.xavier_uniform(p)
+                p.data = nn.init.xavier_uniform(p.data)
     elif init == 'he':
         for p in model.parameters():
             if len(p.size()) >= 2:
-                p.data = nn.init.kaiming_uniform(p)
+                p.data = nn.init.kaiming_uniform(p.data)
     return model
 
 def return_optimizer(optimizer, opt_params, model_params):

### 2017-11-21 Project/MLPs/MLPAutoencoder.py
Add Adadelta, AdaGrad, RMSprop

@@ -101,7 +101,7 @@ for i, itr in enumerate(te_d_loader):
     if p.cuda != -1:    
         x = x.cuda()
     x = V(x)
-    te_loss += loss_fn(model(x), x).data[0]*itr[0].size(0)
+    te_loss += loss_fn(model(x), x).data[0]*(itr[0].size(0))
     N += itr[0].size(0)
 te_loss = te_loss/N
 print('Test Error after {0} iterations with {1} is {2}'.format(p.maxiter, p.opt, round(te_loss, 5)))

### 2017-11-21 Project/Trivial Functions/optimizers.py

@@ -29,7 +29,7 @@ def CM(t, X, G, V, m, lr):
     update = X + velocity
     return timestep, update, velocity
 
-def Adam(t, X, G, FM, SM, betas, lr, eps=1e-08):
+def Adam(t, X, G, FM, SM, betas=(0.9, 0.99), lr, eps=1e-08):
     """
         Function to perform one Adam step
         Args:
@@ -49,3 +49,54 @@ def Adam(t, X, G, FM, SM, betas, lr, eps=1e-08):
     vh = SM/(1 - betas[1]**timestep)
     update = X - lr*mh/(np.sqrt(vh) + eps)
     return timestep, update, FM, SM
+
+def Adagrad(t, X, G, SSG, lr, eps=1e-08):
+    """
+        Function to perform one Adagrad step
+        Args:
+            t   => previous timestep
+            X   => Current point
+            G   => Gradient at current point
+            SSG => Sum of squared gradient till previous timestep
+            lr  => Learning Rate
+            eps => (optional, default=1e-08) Prevent divide by 0 error
+    """
+    timestep = t + 1
+    update = X - lr*G/(np.sqrt(SSG) + eps)
+    SSG = SSG + G*G
+    return timestep, update, SSG
+
+def Adadelta(t, X, G, EG, EDX, lr, rho=0.9, eps=1e-08):
+    """
+        Function to perform one Adadelta step
+        Args:
+            t   => previous timestep
+            X   => Current point
+            G   => Gradient at current point
+            EG  => Running average of squared gradients till previous timestep
+            EDX => Running average of squared updates till previous timestep
+            lr  => Learning Rate
+            eps => (optional, default=1e-08) Prevent divide by 0 error
+    """
+    timestep = t + 1
+    EG = rho*EG + (1 - rho)*G*G
+    delx = -np.sqrt(EDX + eps)*G/np.sqrt(EG + eps)
+    EDX = rho*EDX + (1 - rho)*delx*delx
+    update = X + delx
+    return timestep, update, EG, EDX
+
+def RMSprop(t, X, G, EG, lr, eps=1e-08):
+    """
+        Function to perform one Adadelta step
+        Args:
+            t   => previous timestep
+            X   => Current point
+            G   => Gradient at current point
+            EG  => Running average of squared gradients till previous timestep
+            lr  => Learning Rate
+            eps => (optional, default=1e-08) Prevent divide by 0 error
+    """
[…]

### 2017-11-21 Project/CNNs/CNNClassifier.py
Add CNN, partial CNN module construction

@@ -1,0 +1,141 @@
+import json
+import torch as t
+import numpy as np
+import torch.nn as nn
+from torchvision import transforms
+from argparse import ArgumentParser
+from torch.utils.data import DataLoader
+from torch.autograd import Variable as V
+from utils import return_model, return_optimizer
+from torchvision.datasets import MNIST, CIFAR10, SVHN
+
+np.random.seed(29)
+t.manual_seed(29)
+
+# Get the parameters
+p = ArgumentParser()
+p.add_argument('--opt', required=True, type=str, help='Optimizer --> adam | adagrad | rmsprop | nag | cm')
+p.add_argument('--opt_params', required=True, type=str, help='File containing params in json format')
+p.add_argument('--maxiter', default=int(5e04), type=int, help='Maximum iterations')
+p.add_argument('--architecture', required=True, type=str, help='File containing architecture params in json format')
+p.add_argument('--init', default='random', type=str, help='Initialization scheme to use --> random | he | xavier')
+p.add_argument('--dataset', required=True, type=str, help='Dataset to use --> mnist | cifar10 | svhn')
+p.add_argument('--dataroot', default='./', type=str, help='Data root folder')
+p.add_argument('--cuda', default=-1, type=int, help='Enter CUDA device (> 0), or -1 if no CUDA')
+p = p.parse_args()
+
+# CUDA check
+if p.cuda != -1:
+    t.cuda.set_device(p.cuda)
+
+# Load the data
+if p.dataset == 'mnist':
+    transformations = transforms.Compose([transforms.ToTensor(), 
+                                          transforms.Normalize((0.5,), (0.5,)), 
+                                          transforms.Lambda(lambda x: x.view(784))
+                                        ])
+    tr_dset = MNIST(root=p.dataroot, train=True, transform=transformations, download=True)
+    te_dset = MNIST(root=p.dataroot, train=False, transform=transformations, download=True)
+
+elif p.dataset == 'cifar10':
+    transformations = transforms.Compose([transforms.ToTensor(), 
+                                          transforms.Normalize((0.5, 0.5, 0.5), (0.5, 0.5, 0.5)), 
+                                          transforms.Lambda(lambda x: x.view(3072))
[…]

### 2017-11-21 Project/utils.py

@@ -2,7 +2,7 @@ import torch as t
 import torch.nn as nn
 import torch.optim as optim
 
-def return_model(arch_vals, activation, init, last_layer_active=False):
+def return_MLP_model(arch_vals, activation, init, last_layer_active=False):
     activations = {'relu': nn.ReLU(True), 
                    'sigmoid': nn.Sigmoid(),
                    'tanh': nn.Tanh(),
@@ -24,6 +24,24 @@ def return_model(arch_vals, activation, init, last_layer_active=False):
         for p in model.parameters():
             if len(p.size()) >= 2:
                 p.data = nn.init.kaiming_uniform(p.data)
+    return model
+
+def return_CNN_model(arch_params, init):
+    activations = {'relu': nn.ReLU(True), 
+                   'sigmoid': nn.Sigmoid(),
+                   'tanh': nn.Tanh(),
+                   'leaky_relu': nn.LeakyReLU(0.2, inplace=True)
+                  }
+    model = nn.Sequential()
+    for i in range(0, len(arch_params)):
+        if arch_params[i]['type'] == 'MLP':
+            fan_in, fan_out = arch_params[i]['params']['in'], arch_params[i]['params']['out']
+            model.add_module('linear_{0}_{1}-{2}'.format(i, fan_in, fan_out), nn.Linear(fan_in, fan_out))
+        elif arch_params[i]['type'] == 'CONV':
+            fan_in, fan_out = arch_params[i]['params']['inchan'], arch_params[i]['params']['outchan']
+            k_size = arch_params[i]['params']['kernel']
+            model.add_module('conv_{0}_{1}-{2}'.format(i, fan_in, fan_out), nn.Conv2d(fan_in, fan_out, kernel_size=k_size))
+        model.add_module('{0}_{1}'.format(i, arch_params[i]['activation']), activations[arch_params[i]['activation']])
     return model
 
 def return_optimizer(optimizer, opt_params, model_params):

### 2017-11-21 Project/CNNs/CNNClassifier.py
refactor and add one classifier code

@@ -1,141 +1,0 @@
-import json
-import torch as t
-import numpy as np
-import torch.nn as nn
-from torchvision import transforms
-from argparse import ArgumentParser
-from torch.utils.data import DataLoader
-from torch.autograd import Variable as V
-from utils import return_model, return_optimizer
-from torchvision.datasets import MNIST, CIFAR10, SVHN
-
-np.random.seed(29)
-t.manual_seed(29)
-
-# Get the parameters
-p = ArgumentParser()
-p.add_argument('--opt', required=True, type=str, help='Optimizer --> adam | adagrad | rmsprop | nag | cm')
-p.add_argument('--opt_params', required=True, type=str, help='File containing params in json format')
-p.add_argument('--maxiter', default=int(5e04), type=int, help='Maximum iterations')
-p.add_argument('--architecture', required=True, type=str, help='File containing architecture params in json format')
-p.add_argument('--init', default='random', type=str, help='Initialization scheme to use --> random | he | xavier')
-p.add_argument('--dataset', required=True, type=str, help='Dataset to use --> mnist | cifar10 | svhn')
-p.add_argument('--dataroot', default='./', type=str, help='Data root folder')
-p.add_argument('--cuda', default=-1, type=int, help='Enter CUDA device (> 0), or -1 if no CUDA')
-p = p.parse_args()
-
-# CUDA check
-if p.cuda != -1:
-    t.cuda.set_device(p.cuda)
-
-# Load the data
-if p.dataset == 'mnist':
-    transformations = transforms.Compose([transforms.ToTensor(), 
-                                          transforms.Normalize((0.5,), (0.5,)), 
-                                          transforms.Lambda(lambda x: x.view(784))
-                                        ])
-    tr_dset = MNIST(root=p.dataroot, train=True, transform=transformations, download=True)
-    te_dset = MNIST(root=p.dataroot, train=False, transform=transformations, download=True)
-
-elif p.dataset == 'cifar10':
-    transformations = transforms.Compose([transforms.ToTensor(), 
-                                          transforms.Normalize((0.5, 0.5, 0.5), (0.5, 0.5, 0.5)), 
-                                          transforms.Lambda(lambda x: x.view(3072))
[…]

### 2017-11-21 Project/Classifiers/Classifier.py

@@ -6,7 +6,7 @@ from torchvision import transforms
 from argparse import ArgumentParser
 from torch.utils.data import DataLoader
 from torch.autograd import Variable as V
-from utils import return_model, return_optimizer
+from utils import Model, return_optimizer
 from torchvision.datasets import MNIST, CIFAR10, SVHN
 
 np.random.seed(29)
@@ -17,8 +17,7 @@ p = ArgumentParser()
 p.add_argument('--opt', required=True, type=str, help='Optimizer --> adam | adagrad | rmsprop | nag | cm')
 p.add_argument('--opt_params', required=True, type=str, help='File containing params in json format')
 p.add_argument('--maxiter', default=int(5e04), type=int, help='Maximum iterations')
-p.add_argument('--architecture', required=True, type=str, help='CSV file with number of nodes per layer')
-p.add_argument('--activation', default='relu', type=str, help='Activation function to be used')
+p.add_argument('--architecture', required=True, type=str, help='File containing architecture params in json format')
 p.add_argument('--init', default='random', type=str, help='Initialization scheme to use --> random | he | xavier')
 p.add_argument('--dataset', required=True, type=str, help='Dataset to use --> mnist | cifar10 | svhn')
 p.add_argument('--dataroot', default='./', type=str, help='Data root folder')
@@ -57,9 +56,8 @@ elif p.dataset == 'svhn':
 tr_d_loader = DataLoader(dataset=tr_dset, batch_size=64, shuffle=True)
 te_d_loader = DataLoader(dataset=te_dset, batch_size=5000, shuffle=True)
 
-# Build MLP architecture
-arch_vals = np.genfromtxt(p.architecture, delimiter=',').reshape(-1).astype(int).tolist()
-model = return_model(arch_vals, p.activation, init=p.init)
+# Build classifier architecture
+model = Model(json.load(open(p.architecture)), p.init)
 print(model)
 
 if p.cuda != -1:

### 2017-11-21 Project/utils.py

@@ -2,47 +2,62 @@ import torch as t
 import torch.nn as nn
 import torch.optim as optim
 
-def return_MLP_model(arch_vals, activation, init, last_layer_active=False):
-    activations = {'relu': nn.ReLU(True), 
-                   'sigmoid': nn.Sigmoid(),
-                   'tanh': nn.Tanh(),
-                   'leaky_relu': nn.LeakyReLU(0.2, inplace=True)
-                  }
-    model = nn.Sequential()
-    for i in range(0, len(arch_vals) - 1):
-        model.add_module('linear_{0}-{1}-{2}'.format(i, arch_vals[i], arch_vals[i+1]), nn.Linear(arch_vals[i], arch_vals[i+1]))
-        if i != len(arch_vals) - 2:
-            model.add_module('{0}_{1}'.format(activation, i), activations[activation])
-    if last_layer_active:
-        model.add_module('{0}_{1}'.format(activation, i+1), activations[activation])
+class Model(nn.Module):
+    def __init__(self, net_params, init_scheme):
+        """
+            Class to produce models based on JSON objects
+        """
+        super(Model, self).__init__()
 
-    if init == 'xavier':
-        for p in model.parameters():
-            if len(p.size()) >= 2:
-                p.data = nn.init.xavier_uniform(p.data)
-    elif init == 'he':
-        for p in model.parameters():
-            if len(p.size()) >= 2:
-                p.data = nn.init.kaiming_uniform(p.data)
-    return model
+        activations = {'relu': nn.ReLU(True), 
+                       'sigmoid': nn.Sigmoid(),
+                       'tanh': nn.Tanh(),
+                       'leaky_relu': nn.LeakyReLU(0.2, inplace=True)
+                      }
 
-def return_CNN_model(arch_params, init):
-    activations = {'relu': nn.ReLU(True), 
-                   'sigmoid': nn.Sigmoid(),
-                   'tanh': nn.Tanh(),
-                   'leaky_relu': nn.LeakyReLU(0.2, inplace=True)
-                  }
-    model = nn.Sequential()
-    for i in range(0, len(arch_params)):
-        if arch_params[i]['type'] == 'MLP':
-            fan_in, fan_out = arch_params[i]['params']['in'], arch_params[i]['params']['out']
-            model.add_module('linear_{0}_{1}-{2}'.format(i, fan_in, fan_out), nn.Linear(fan_in, fan_out))
[…]

### 2017-11-22 Project/GANs/CIFAR10.py
Added DCGAN

@@ -1,0 +1,158 @@
+import torch as t
+import torch.nn as nn
+
+class Generator(nn.Module):
+	"""
+		Selected Generator for CIFAR10 dataset. This produces images of size 32 x 32.
+		Use this module for any GANs to be trained on CIFAR10.
+	"""
+	def __init__(self, n_z, ngpu, gen_type):
+		"""
+		Function to construct a Generator instance
+		Args:
+			n_z		: Dimensionality of the noise
+			ngpu		: Number of GPUs to be used
+			gen_type	: Type of Generator: 
+						'conv' for normal convolutional,
+						'resi' for residual blocks
+		"""
+		super(Generator, self).__init__()
+
+		assert ngpu >= 0, "Number of GPUs has to be non-negative"
+		assert n_z > 0, "Dimensionality of the noise vector has to be positive"
+		
+		if gen_type == 'conv':
+			# Architecture: Specified as follows:
+			# |   INPUT 	---->	  OUTPUT	(	   ACTIVATIONS 		  ) |
+			# |    n_z	---->	   4096		(BATCHNORM_1D, LEAKY_RELU, DROPOUT) |
+			# |  4x4x256	---->    8X8X128	(       BATCHNORM_2D, RELU 	  ) |
+			# |  8x8x128	---->    16x16x64	(       BATCHNORM_2D, RELU	  ) |
+			# |  16X16X64	---->	 32X32X3	(       BATCHNORM_2D, TANH	  ) |
+			dropout_prob	= 0.4
+			leaky_coeff	= 0.2
+			
+			# Fully Connected Section
+			layer	= 1
+			main	= nn.Sequential()
+			main.add_module('Linear_{0}-{1}-{2}'.format(layer, n_z, 4096), nn.Linear(n_z, 4096))
+			main.add_module('BatchNorm1d_{0}-{1}'.format(layer, 4096), nn.BatchNorm1d(4096))
+			main.add_module('LeakyReLU_{0}'.format(layer), nn.LeakyReLU(leaky_coeff, inplace=True))
+			main.add_module('Dropout_{0}'.format(layer), nn.Dropout(dropout_prob))
+			self.fc	= main
+			
+			# Convolutional Section
+			layer	= layer + 1
+			main	= nn.Sequential()
+			main.add_module('ConvTranspose2d_{0}-{1}-{2}'.format(layer, 256, 128),
[…]

### 2017-11-22 Project/GANs/DCGAN.py

@@ -1,0 +1,229 @@
+import torch as t
+import gan_utils as u
+import torch.nn as nn
+import torch.utils.data as d_utils
+import torchvision.utils as tv_utils
+from torch.autograd import Variable as V
+
+class DCGAN(object):
+	def __init__(self, arch, ngpu, loss='BCE'):
+		"""
+		DCGAN object. This class is a wrapper of a generalized DCGAN as explained in the paper: 
+			UNSUPERVISED REPRESENTATION LEARNING WITH DEEP CONVOLUTIONAL GENERATIVE ADVERSARIAL NETWORKS by Alec Radford et.al.
+		
+		Instance of this class initializes the Generator and the Discriminator.
+		Arguments:
+			arch			= Architecture to use:
+    							"CIFAR10" for CIFAR10 dataset
+	    						"MNIST" for MNIST dataset
+						  
+	    					  For CIFAR10/MNIST: Need to input n_z also. 
+	    					  For Generic: Need to input image_size, n_z, n_chan, hiddens
+						  
+    						  {'arch_type': <arch_type>,
+    						   'params'   : <params> as above
+    						  }
+						  
+    						  Example:
+    						  {'arch_type': "CIFAR10",
+    						   'params'   : {'n_z' : 128
+                                            }
+                              }
+
+    						  {'arch_type': "Generic",
+    						   'params'   : {'image_size'	: 32,
+        						    		 'n_z'		    : 128,
+        						    		 'n_chan'	    : 3,
+        						    		 'hiddens'	    : <see below>
+        						    		}
+    						  }
+
+                              image_size		= Height / width of the real images
+                              n_z	    		= Dimensionality of the latent space
+                              n_chan			= Number of channels of the real images
+                              hiddens			= Number of feature maps in the first layer of the generator and discriminator
+										          Format:
+										          	hiddens = {'gen': n_gen_hidden, 
+										              		   'dis': n_dis_hidden
+										              		  }
+                                        			ngpu			= Number of gpus to be allocated, if to be run on gpu
+                                        			loss			= The loss function to be used
+		"""
+		super(DCGAN, self).__init__()
+		if arch['arch_type'] == 'MNIST':
+			import MNIST as DG
+			self.Gen_net	= DG.Generator(
[…]

### 2017-11-22 Project/GANs/gan_util.py

@@ -1,0 +1,49 @@
+import torch as t
+import numpy as np
+
+def get_optimizer_with_params(optimizer_details):
+    optim_name = optimizer_details['name']
+    learn_rate = optimizer_details['learn_rate']
+    optim_params = optimizer_details['params']
+
+    if 'name' not in optimizer_details.keys() or 'learn_rate' not in optimizer_details.keys() or 'params' not in optimizer_details.keys():
+        raise ValueError("Important optimizer details not mentioned")
+
+    if optim_name == 'adadelta':
+        return t.optim.Adadelta(params=optim_params, lr=learn_rate)
+    if optim_name == 'adagrad':
+        return t.optim.Adagrad(params=optim_params, lr=learn_rate)
+    if optim_name == 'adam':
+        if 'betas' in optimizer_details.keys():
+            return t.optim.Adam(params=optim_params, lr=learn_rate, betas=optimizer_details['betas'])
+        else:
+            return t.optim.Adam(params=optim_params, lr=learn_rate)
+    if optim_name == 'adamax':
+        if 'betas' in optimizer_details.keys():
+            return t.optim.Adamax(params=optim_params, lr=learn_rate, betas=optimizer_details['betas'])
+        else:
+            return t.optim.Adamax(params=optim_params, lr=learn_rate)
+    if optim_name == 'rmsprop':
+        return t.optim.RMSprop(params=optim_params, lr=learn_rate)
+    if optim_name == 'sgd':
+        if 'momentum' in optimizer_details.keys():
+            if 'nesterov' in optimizer_details.keys():
+                return t.optim.SGD(params=optim_params, lr=learn_rate, momentum=optimizer_details['momentum'], nesterov=optimizer_details['nesterov'])
+            else:
+                return t.optim.SGD(params=optim_params, lr=learn_rate, momentum=optimizer_details['momentum'])
+        else:
+            return t.optim.SGD(params=optim_params, lr=learn_rate)
+
+def weight_init_scheme(mdl):
+    mdl_name = mdl.__class__.__name__
+    if mdl_name.find('Linear') != -1:
+        f_in = mdl.weight.size(0)
+        f_out = mdl.weight.size(1)
+        c = np.sqrt(6/(f_in + f_out))
+        mdl.weight.data.uniform_(-c, c)
+        mdl.bias.data.fill_(0)
+    if mdl_name.find('Conv') != -1:
[…]

### 2017-11-22 Project/GANs/main.py

@@ -1,0 +1,53 @@
+import sys
+import DCGAN as dc
+
+# Dataset
+dset	= sys.argv[1]
+root	= sys.argv[2]
+
+# Get dataset
+if dset == 'mnist':
+	transformations	= [transforms.Scale(image_size), transforms.ToTensor()]
+	if normalize == True:
+		transformations.append(transforms.Normalize((0.5, ), (0.5, )))
+	dataset	= dset.MNIST(root=root, download=True, transform=transforms.Compose(transformations))
+    n_chan = 1
+elif dset == 'cifar10':
+	transformations	= [transforms.Scale(image_size), transforms.ToTensor()]
+	if normalize == True:
+		transformations.append(transforms.Normalize((0.5, 0.5, 0.5), (0.5, 0.5, 0.5)))
+	dataset = dset.CIFAR10(root=root, download=True, transform=transforms.Compose(transformations))
+    n_chan = 3
+	
+# DCGAN object initialization
+image_size = 32
+n_z = 128
+hiddens	= {'gen':	64,
+		   'dis':	64
+		  }
+ngpu = 1
+loss = 'BCE'
+
+Gen_model	= dc.DCGAN(image_size=image_size, n_z=n_z, n_chan=n_chan, hiddens=hiddens, ngpu=ngpu, loss=loss)
+
+# DCGAN training scheme
+batch_size	= 100
+n_iters		= 1e05
+opt_dets	= {'gen':	{'name'         : 'adam',
+				         'learn_rate'	: 1e-04,
+				         'betas'        : (0.5, 0.99)
+				        },
+    		   'dis':	{'name'		    : 'sgd',
+	        	   		 'learn_rate'	: 1e-04,
+	        	   		 'momentum'	    : 0.9,
+	        	   		 'nesterov'	    : True
+	        	   		}
+    		  }
+
+# Optional arguments
+show_period	= 50
+display_images	= True
+misc_options	= ['init_scheme', 'save_model']
+
+# Call training
+Gen_model.train(dataset=dataset, batch_size=batch_size, n_iters=n_iters, optimizer_details=opt_dets, show_period=show_period, display_images=display_images, misc_options=misc_options)

### 2017-12-02 Project/Autoencoders/MLPAutoencoder.py
Minor changes to scripts

@@ -79,14 +79,14 @@ if p.cuda != -1:
     loss_fn = loss_fn.cuda()
 
 optimizer_params = json.load(open(p.opt_params))
-params = [p.dataset, p.opt]
+params = [p.dataset]
 for k in sorted(list(optimizer_params)):
     params.append(optimizer_params[k])
 for k in arch_vals:
     params.append(k)
 params += ['sigmoid', p.init]
 
-loss_log = open('loss_autoencoder_{0}.txt'.format(params), 'w')
+loss_log = open('./{0}/loss_autoencoder_{1}.txt'.format(p.opt, params), 'w')
 
 flag = False
 iters = 0

### 2017-12-02 Project/Autoencoders/autoencoder_script.sh

@@ -6,11 +6,12 @@ for arch in ` ls ../ARCHS/Autoencoders `;
     do    
     for opt in 'adadelta' 'adagrad' 'adam' 'cm' 'nag' 'rmsprop';
         do
+        mkdir $opt
         for opt_params in ` ls ../OPT_JSONS/$opt `;
             do
             python3 MLPAutoencoder.py --opt $opt --opt_params ../OPT_JSONS/$opt/$opt_params \
                                       --architecture ../ARCHS/Autoencoders/$arch --init xavier --dataset mnist \
-                                      --dataroot ../Datasets/MNIST --cuda $cuda;
+                                      --dataroot ../Datasets/MNIST --cuda $cuda --maxiter 10000;
             done;
         done;
     done

### 2017-12-03 Project/Trivial Functions/functions.py
minor changes in Trivial Functions

@@ -58,3 +58,5 @@ def ST(X):
     x = X[0]
     y = X[1]
     return x*4 - 16*(x**2) + 5*x + y**4 - 16*(y**2) + 5*y
+
+_func_dicts = {'B1': B1, 'B2': B2, 'B3': B3, 'RB': RB, 'BL': BL, 'ST': ST}

### 2017-12-03 Project/Trivial Functions/gradients.py

@@ -8,7 +8,7 @@ import numpy as np
 # BL -> Beale's Function
 # ST -> Styblinsky-Tang's Function
 
-def B0(X):
+def B1(X):
     """
         Returns gradient of Bohachevsky Function 1 at a point
     """
@@ -18,7 +18,7 @@ def B0(X):
     dy = 4*y + 4.0*np.pi*0.4*np.sin(4.0*np.pi*y)
     return np.array([dx, dy])
 
-def B1(X):
+def B2(X):
     """
         Returns gradient of Bohachevsky Function 2 at a point
     """
@@ -28,7 +28,7 @@ def B1(X):
     dy = 4*y + 4.0*np.pi*0.4*np.sin(4.0*np.pi*y)*np.cos(3.0*np.pi*x)
     return np.array([dx, dy])
 
-def B2(X):
+def B3(X):
     """
         Returns gradient of Bohachevsky Function 3 at a point
     """
@@ -75,3 +75,6 @@ def ST(X):
     y = X[1]
     dx = 4*x**3 - 32*x + 5
     dy = 4*y**3 - 32*y + 5
+    return np.array([dx, dy])
+
+_grad_dicts = {'B1': B1, 'B2': B2, 'B3': B3, 'RB': RB, 'BL': BL, 'ST': ST}

### 2017-12-03 Project/Trivial Functions/optimizers.py

@@ -100,3 +100,5 @@ def RMSprop(t, X, G, EG, lr, eps=1e-08):
     EG = 0.9*EG + 0.1*G*G
     update = X - lr*G/np.sqrt(EG + eps)
     return timestep, update, EG
+
+_opt_dicts = {'GD': GD, 'CM': CM, 'Adam': Adam, 'Adagrad': Adagrad, 'Adadelta': Adadelta, 'RMSprop': RMSprop}

### 2017-12-03 Project/Trivial Functions/main.py
Added a temporary main file

@@ -1,0 +1,22 @@
+import sys
+import numpy as np
+import functions as f
+import gradients as g
+import optimizers as o
+from argparse import ArgumentParser as AP
+
+p = AP()
+p.add_argument('--optim', required=True, help='Optimizers: GD | CM | Adam | Adagrad | Adadelta | RMSprop')
+p.add_argument('--lr', type=float, required=True, help='Learning rate')
+p.add_argument('--fn', required=True, help='Function name: B1 | B2 | B3 | BL | RB | ST')
+p.add_argument('--iter', default=250, type=int, help='Number of iterations to run for')
+p = p.parse_args()
+
+i = 0
+X = np.full(2, 2)
+log_file = open('log_{0}_{1}_{2}.txt'.format(p.optim, p.lr, p.fn), 'w')
+while i < p.iter:
+    log_file.write('{0}\t{1}\n'.format(i, f._func_dicts[p.fn](X)))
+    G = g._grad_dicts[p.fn](X)
+    i, X = o._opt_dicts[p.optim](i, X, G, p.lr)
+log_file.close()

### 2017-12-03 Project/Trivial Functions/optimizers.py
minor changes

@@ -13,7 +13,7 @@ def GD(t, X, G, lr):
     update = X - lr*G
     return timestep, update
 
-def CM(t, X, G, V, m, lr):
+def CM(t, X, G, V, lr, m):
     """
         Function to perform one Classical Momentum step
         Args:
@@ -29,7 +29,7 @@ def CM(t, X, G, V, m, lr):
     update = X + velocity
     return timestep, update, velocity
 
-def Adam(t, X, G, FM, SM, betas=(0.9, 0.99), lr, eps=1e-08):
+def Adam(t, X, G, lr, FM, SM, betas=(0.9, 0.99), eps=1e-08):
     """
         Function to perform one Adam step
         Args:

### 2017-12-03 Project/Trivial Functions/gradients.py
Main file which runs an optimizer on a function for multiple learning rates.

@@ -54,18 +54,18 @@ def BL(X):
     """
     x = X[0]
     y = X[1]
-	cmn1	= 2*(1.5 - x + x*y)
-	cmn2	= 2*(2.25 - x + x*y**2)
-	cmn3	= 2*(2.625 - x + x*y**3)
-	tmx1	= -1 + y
-	tmx2	= -1 + y**2
-	tmx3	= -1 + y**3
-	tmy1	= x
-	tmy2	= 2*x*y
-	tmy3	= 3*x*y**2
-	dx	= cmn1*tmx1 + cmn2*tmx2 + cmn3*tmx3
-	dy	= cmn1*tmy1 + cmn2*tmy2 + cmn3*tmy3
-	return np.array([dx, dy])
+    cmn1    = 2*(1.5 - x + x*y)
+    cmn2    = 2*(2.25 - x + x*y**2)
+    cmn3    = 2*(2.625 - x + x*y**3)
+    tmx1    = -1 + y
+    tmx2    = -1 + y**2
+    tmx3    = -1 + y**3
+    tmy1    = x
+    tmy2    = 2*x*y
+    tmy3    = 3*x*y**2
+    dx  = cmn1*tmx1 + cmn2*tmx2 + cmn3*tmx3
+    dy  = cmn1*tmy1 + cmn2*tmy2 + cmn3*tmy3
+    return np.array([dx, dy])
 
 def ST(X):
     """

### 2017-12-03 Project/Trivial Functions/main.py

@@ -4,19 +4,87 @@ import functions as f
 import gradients as g
 import optimizers as o
 from argparse import ArgumentParser as AP
+import os, errno
 
 p = AP()
 p.add_argument('--optim', required=True, help='Optimizers: GD | CM | Adam | Adagrad | Adadelta | RMSprop')
-p.add_argument('--lr', type=float, required=True, help='Learning rate')
 p.add_argument('--fn', required=True, help='Function name: B1 | B2 | B3 | BL | RB | ST')
 p.add_argument('--iter', default=250, type=int, help='Number of iterations to run for')
 p = p.parse_args()
 
 i = 0
 X = np.full(2, 2)
-log_file = open('log_{0}_{1}_{2}.txt'.format(p.optim, p.lr, p.fn), 'w')
-while i < p.iter:
-    log_file.write('{0}\t{1}\n'.format(i, f._func_dicts[p.fn](X)))
-    G = g._grad_dicts[p.fn](X)
-    i, X = o._opt_dicts[p.optim](i, X, G, p.lr)
-log_file.close()
+
+learning_rates = np.logspace(-4, -2, 12)
+
+newDir = p.fn + '/' + p.optim
+try:
+    os.makedirs(newDir)
+except OSError as e:
+    if e.errno != errno.EEXIST:
+        raise
+
+
+if p.optim == 'GD':
+    for lr in learning_rates:
+        log_file = open('{0}/log_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')
+        while i < p.iter:
+            G = g._grad_dicts[p.fn](X)
+            i, X = o._opt_dicts[p.optim](i, X, G, lr)
+        log_file.close()
+
+elif p.optim == 'CM':
+    v = 0
+    m = 0.95
+    for lr in learning_rates:
+        log_file = open('{0}/log_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')
+        while i < p.iter:
+            G = g._grad_dicts[p.fn](X)
+            i, X, v = o._opt_dicts[p.optim](i, X, G, v, lr, m)
+        log_file.close()
+
+elif p.optim == 'Adam':
+    fm = 0
+    sm = 0
+    for lr in learning_rates:
+        log_file = open('{0}/log_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')
+        while i < p.iter:
[…]

### 2017-12-03 Project/Trivial Functions/main.py
Fix minor bugs, add function initializers

@@ -1,10 +1,24 @@
 import sys
+import os, errno
 import numpy as np
 import functions as f
 import gradients as g
 import optimizers as o
 from argparse import ArgumentParser as AP
-import os, errno
+
+def give_init(function_name):
+    if function_name == 'B1':
+        return np.array([-75, 50])
+    elif function_name == 'B2':
+        return np.array([50, -50])
+    elif function_name == 'B3':
+        return np.array([-50, 50])
+    elif function_name == 'BL':
+        return np.array([1, 2])
+    elif function_name == 'RB':
+        return np.array([-2.1, 2.1])
+    elif function_name == 'ST':
+        return np.array([0, 0])
 
 p = AP()
 p.add_argument('--optim', required=True, help='Optimizers: GD | CM | Adam | Adagrad | Adadelta | RMSprop')
@@ -12,33 +26,36 @@ p.add_argument('--fn', required=True, help='Function name: B1 | B2 | B3 | BL | R
 p.add_argument('--iter', default=250, type=int, help='Number of iterations to run for')
 p = p.parse_args()
 
-i = 0
-X = np.full(2, 2)
 
-learning_rates = np.logspace(-4, -2, 12)
-
-newDir = p.fn + '/' + p.optim
+newDir = os.path.join(p.fn, p.optim)
 try:
     os.makedirs(newDir)
 except OSError as e:
     if e.errno != errno.EEXIST:
         raise
 
-
 if p.optim == 'GD':
+    learning_rates = np.around(np.logspace(-4, -2.5, 12), decimals=5)
     for lr in learning_rates:
+        i = 0
+        X = give_init(p.fn)
         log_file = open('{0}/log_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')
         while i < p.iter:
+            log_file.write('{0}\t{1}\n'.format(i, f._func_dicts[p.fn](X)))
             G = g._grad_dicts[p.fn](X)
             i, X = o._opt_dicts[p.optim](i, X, G, lr)
         log_file.close()
 
 elif p.optim == 'CM':
     v = 0
-    m = 0.95
+    m = 0.925
+    learning_rates = np.around(np.logspace(-4, -2.5, 12), decimals=5)
     for lr in learning_rates:
+        i = 0
+        X = give_init(p.fn)
[…]

### 2017-12-03 Project/Trivial Functions/main.py
Minor bug fix

@@ -47,10 +47,10 @@ if p.optim == 'GD':
         log_file.close()
 
 elif p.optim == 'CM':
-    v = 0
-    m = 0.925
     learning_rates = np.around(np.logspace(-4, -2.5, 12), decimals=5)
     for lr in learning_rates:
+        v = 0
+        m = 0.925
         i = 0
         X = give_init(p.fn)
         log_file = open('{0}/log_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')
@@ -61,10 +61,10 @@ elif p.optim == 'CM':
         log_file.close()
 
 elif p.optim == 'Adam':
-    fm = 0
-    sm = 0
     learning_rates = np.around(np.logspace(-4, -1.5, 18), decimals=5)
     for lr in learning_rates:
+        fm = 0
+        sm = 0
         i = 0
         X = give_init(p.fn)
         log_file = open('{0}/log_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')
@@ -88,10 +88,10 @@ elif p.optim == 'Adagrad':
         log_file.close()
 
 elif p.optim == 'Adadelta':
-    eg = 0
-    edx = 0
     learning_rates = np.around(np.logspace(-4, -1.5, 18), decimals=5)
     for lr in learning_rates:
+        eg = 0
+        edx = 0
         i = 0
         X = give_init(p.fn)
         log_file = open('{0}/log_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')
@@ -102,9 +102,9 @@ elif p.optim == 'Adadelta':
         log_file.close()
 
 elif p.optim == 'RMSprop':
-    eg = 0
     learning_rates = np.around(np.logspace(-4, -1.5, 18), decimals=5)
     for lr in learning_rates:
+        eg = 0
         i = 0
         X = give_init(p.fn)
         log_file = open('{0}/log_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')

### 2017-12-03 Project/Trivial Functions/main.py
Fixed adagrad bug

@@ -75,9 +75,9 @@ elif p.optim == 'Adam':
         log_file.close()
 
 elif p.optim == 'Adagrad':
-    ssg = 0
-    learning_rates = np.around(np.logspace(-4, -1.5, 18), decimals=5)
+    learning_rates = np.around(np.logspace(-4, 0, 30), decimals=5)
     for lr in learning_rates:
+        ssg = 0
         i = 0
         X = give_init(p.fn)
         log_file = open('{0}/log_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')
@@ -88,7 +88,7 @@ elif p.optim == 'Adagrad':
         log_file.close()
 
 elif p.optim == 'Adadelta':
-    learning_rates = np.around(np.logspace(-4, -1.5, 18), decimals=5)
+    learning_rates = np.around(np.logspace(-4, 0, 30), decimals=5)
     for lr in learning_rates:
         eg = 0
         edx = 0

### 2017-12-03 Project/Trivial Functions/optimizers.py

@@ -62,8 +62,8 @@ def Adagrad(t, X, G, SSG, lr, eps=1e-08):
             eps => (optional, default=1e-08) Prevent divide by 0 error
     """
     timestep = t + 1
-    update = X - lr*G/(np.sqrt(SSG) + eps)
     SSG = SSG + G*G
+    update = X - lr*G/(np.sqrt(SSG) + eps)
     return timestep, update, SSG
 
 def Adadelta(t, X, G, EG, EDX, lr, rho=0.9, eps=1e-08):

### 2017-12-03 Project/Trivial Functions/main.py
fix terrible mistake with adadelta

@@ -23,7 +23,7 @@ def give_init(function_name):
 p = AP()
 p.add_argument('--optim', required=True, help='Optimizers: GD | CM | Adam | Adagrad | Adadelta | RMSprop')
 p.add_argument('--fn', required=True, help='Function name: B1 | B2 | B3 | BL | RB | ST')
-p.add_argument('--iter', default=250, type=int, help='Number of iterations to run for')
+p.add_argument('--iter', default=1000, type=int, help='Number of iterations to run for')
 p = p.parse_args()
 
 

### 2017-12-03 Project/Trivial Functions/optimizers.py

@@ -80,7 +80,7 @@ def Adadelta(t, X, G, EG, EDX, lr, rho=0.9, eps=1e-08):
     """
     timestep = t + 1
     EG = rho*EG + (1 - rho)*G*G
-    delx = -np.sqrt(EDX + eps)*G/np.sqrt(EG + eps)
+    delx = -lr*np.sqrt(EDX + eps)*G/np.sqrt(EG + eps)
     EDX = rho*EDX + (1 - rho)*delx*delx
     update = X + delx
     return timestep, update, EG, EDX

### 2017-12-03 Project/Trivial Functions/script.py
Finding the  best parameters for the optimizers

@@ -1,0 +1,19 @@
+import subprocess
+import os
+import numpy as np
+
+functions = ['B1', 'B2', 'B3', 'BL', 'RB', 'ST']
+optims = ['GD', 'CM', 'Adam', 'Adagrad', 'Adadelta', 'RMSprop']
+
+for i in functions:
+    for j in optims:
+        subprocess.call(["python3", "main.py", "--optim", j, "--fn", i]) 
+
+for directory_1 in functions:
+    for directory_2 in optims:
+        data = []
+        file_list = os.listdir(directory_1+'/'+directory_2)
+        for filename in file_list:
+            filepath = directory_1 +'/'+ directory_2 +'/'+ filename
+            data.append(np.genfromtxt(filepath)[-1,1])
+        print(directory_1, directory_2, file_list[np.argsort(data)[-1]])

### 2017-12-04 Project/Classifiers/Classifier.py
Classifiers MLP archs

@@ -55,7 +55,6 @@ te_d_loader = DataLoader(dataset=te_dset, batch_size=5000, shuffle=True)
 
 # Build classifier architecture
 model = Model(json.load(open(p.architecture)), p.init)
-model.set_batch_size(64)
 print(model)
 
 if p.cuda != -1:
@@ -74,9 +73,9 @@ optimizer_params = json.load(open(p.opt_params))
 params = [p.dataset, p.opt]
 for k in sorted(list(optimizer_params)):
     params.append(optimizer_params[k])
-params.append(p.architecture[:-5])
+params.append(p.architecture[-11:-5])
 
-loss_log = open('loss_classifier_{0}.txt'.format(params), 'w')
+loss_log = open('./{0}/loss_classifier_{1}.txt'.format(p.opt, params), 'w')
 
 flag = False
 iters = 0
@@ -88,6 +87,7 @@ while flag != True:
             x = x.cuda()
             y = y.cuda()
         x, y = V(x), V(y)
+        model.set_batch_size(x.size(0))
         cur_loss = loss_fn(model(x), y)
         loss_log.write('{0}\t{1}\n'.format(iters, round(cur_loss.data[0], 6)))
         if iters == p.maxiter:
@@ -112,6 +112,7 @@ for i, itr in enumerate(tr_d_loader):
         x = x.cuda()
         y = y.cuda()
     x, y = V(x), V(y)
+    model.set_batch_size(x.size(0))
     forward_pass = model(x)
     maxes, pred_y = forward_pass.max(1)
     pred_y = pred_y.view(-1)
@@ -129,6 +130,7 @@ for i, itr in enumerate(te_d_loader):
         x = x.cuda()
         y = y.cuda()
     x, y = V(x), V(y)
+    model.set_batch_size(x.size(0))
     forward_pass = model(x)
     maxes, pred_y = forward_pass.max(1)
     pred_y = pred_y.view(-1)

### 2017-12-04 Project/Classifiers/classifier_script.sh

@@ -1,0 +1,36 @@
+#!/bin/bash
+
+cuda=$1
+dset=$2
+droot=$3
+s=0
+e=0
+
+if [ "$dset" == "mnist" ] ; then
+    s=1
+    e=4
+fi
+
+if [ "$dset" == "cifar10" ] ; then
+    s=5
+    e=8
+fi
+
+if [ "$dset" == "svhn" ] ; then
+    s=5
+    e=8
+fi
+
+for index in ` seq $s $e `;
+    do    
+    for opt in 'adadelta' 'adagrad' 'adam' 'cm' 'nag' 'rmsprop';
+        do
+        mkdir $opt
+        for opt_params in ` ls ../OPT_JSONS/$opt `;
+            do
+            python3 Classifier.py --opt $opt --opt_params ../OPT_JSONS/$opt/$opt_params \
+                                  --architecture ../ARCHS/Classifiers/arch_$index.json --init xavier --dataset mnist \
+                                  --dataroot $droot --cuda $cuda --maxiter 10000;
+            done;
+        done;
+    done

### 2017-12-04 Project/Classifiers/utils.py

@@ -51,8 +51,9 @@ class Model(nn.Module):
                     p.data = nn.init.kaiming_uniform(p.data)
 
     def forward(self, input):
+        pass_ = input
         if self.conv_part is not None:
-            pass_ = self.conv_part(input)
+            pass_ = self.conv_part(pass_)
         pass_ = pass_.view(self.batch_size, -1)
         if self.mlp_part is not None:
             pass_ = self.mlp_part(pass_)

### 2017-12-04 Project/Presentation/main.tex
Slides for presentation. [WIP]

\documentclass[10pt]{beamer}
\usepackage[style=verbose,backend=biber]{biblatex}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bookmark}
\usepackage{hyperref}
\newcommand{\g}{\nabla}
\newcommand{\xbold}{\mathbf{x}}
\newcommand{\citecomma}{\({}^{{}_{,}}\)}
%\documentclass{article}
%\usepackage{beamerarticle}
\usetheme{CambridgeUS}
\addbibresource{main.bib}
\renewcommand{\citesetup}{\tiny}

\title[Analysis of AG Methods]{An Empirical Analysis of the Benefits of Adaptive Gradient Methods}
\subtitle{Project done as a part of \\
CS6230 - Optimization Methods in Machine Learning}

\author{Vishwak Srinivasan \and Ayushi Patel}

\begin{document}
\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Outline}
\tableofcontents
\end{frame}

\section{Introduction}
\subsection{Gradient Descent}
\begin{frame}{Gradient Descent}
\begin{itemize}
\item<1->{Gradient Descent (GD) - based methods have been a staple approach used for optimization
          in both convex and non-convex problems for many years now.
          \begin{itemize}
          \item<2->{Huge success attributed to the simplicity of the method}
          \end{itemize}
         }
\item<3->{Start with a random initialization, and follow the direction of descent.
          \begin{itemize}
          \item<4->{Direction of descent given by the negative of the gradient of a particular point.}
          \end{itemize}
         }
\uncover<5->{\[\xbold := \xbold - \eta \g f(\xbold)\]}
\end{itemize}
\end{frame}

\subsection{Modifications to Gradient Descent}
\begin{frame}{Modifications to Gradient Descent}
\begin{itemize}
\item<1->{Polyak and Nesterov have proposed methods which \textit{provide an acceleration} to the updates. % add citations
          \begin{itemize}
          \item<2->{Collectively called \textbf{momentum based methods}}
          \end{itemize}
         }
\item<3->{Gradient descent is used as \textit{mini-batch gradient descent} in practice, which leads
          to \textit{variance in the calculated mini-batch gradients}. Methods like SVRG, SAG and % add citations
          SDCA have been proposed to help solve this problem.
          \begin{itemize}
          \item<4->{Collectively called \textbf{variance reduction methods}}
          \end{itemize}
         }
\item<5->{Choice of learning rate is cumbersome process in gradient based methods. Methods like Adam,
          Adagrad, Adadelta, RMSprop and so on \textit{adapt the learning rate} through the updates. % add citations
          \begin{itemize}
          \item<6->{Collectively called \textbf{adaptive gradient methods}}
          \end{itemize}
         }
\end{itemize}
\end{frame}

\section{Adaptive Gradient Methods}
\begin{frame}{Introduction to Adaptive Gradient Methods}
\uncover<1->{General form of any adaptive gradient method is: \[\xbold_{t+1} := \xbold - \eta_{t}\g_{\xbold_{t}} f(\xbold_{t})\] where \(\xbold_{t}\) is the value of \(\xbold\) at iteration \(t\)}
\(\newline\)

\uncover<2->{Key idea of any adaptive gradient method is to \textit{modify the learning rate} based on previous updates. Following slides will discuss some ``well-known'' adaptive gradient methods used in practice briefly.}
\end{frame}

\end{document}

### 2017-12-04 Project/Presentation/main.tex
Added Optimizers slides

@@
 \titlepage
-\end{frame}
-
-\begin{frame}{Outline}
-\tableofcontents
 \end{frame}
@@
 \section{Adaptive Gradient Methods}
+\subsection{Introduction}
 \begin{frame}{Introduction to Adaptive Gradient Methods}
@@
 
+\subsection{Adagrad}
+\begin{frame}{Adagrad} % add citations
+\uncover<1->{Define: \[G_{t} \triangleq \mathrm{diag}\left(\displaystyle \sum_{i=0}^{t} \g f(\xbold_{i}) \odot \g f(\xbold_{i})\right)\]}
+\uncover<2->{Adagrad updates parameters based on the rule below:
+\begin{equation}
+\label{adagrad}
+\xbold_{t+1} = \xbold_{t} - \frac{\eta}{\sqrt{G_{t} + \epsilon}} \odot \g f(\xbold_{t})
+\end{equation}
+where \(\epsilon\) is a small value to prevent divide by zero errors
+}
+\end{frame}
+
+\subsection{Adadelta}
+\begin{frame}{Adadelta} % add citations
+\uncover<1->{Define: \[g_{t} \triangleq \g f(\xbold_{t}) \odot \g f(\xbold_{t})\]
+\[\mathbb{E}[g]_{t} \triangleq \rho\mathbb{E}[g]_{t-1} + (1 - \rho)g_{t}\]
+\[\Delta \xbold_{t} \triangleq -\eta\displaystyle\frac{\sqrt{\mathbb{E}[\Delta \xbold]_{t-1} + \epsilon}}{\sqrt{\mathbb{E}[g]_{t} + \epsilon}}\odot\g f(\xbold_{t})\]
+\[\mathbb{E}[\Delta \xbold]_{t} \triangleq \rho\mathbb{E}[\Delta \xbold]_{t-1} + (1 - \rho)\left(\Delta \xbold_{t} \odot \Delta \xbold_{t}\right)\]}
+\uncover<2->{Adadelta updates parameters based on the rule below:
+\begin{equation}
+\label{adadelta}
+\xbold_{t+1} = \xbold_{t} + \Delta \xbold_{t}
+\end{equation}
+where \(\epsilon\) is a small value to ensure stability/relieve divide by zero errors.
+}
+\end{frame}
+
+\subsection{Adam}
+\begin{frame}{Adam} % add citations
+\uncover<1->{Define: \[m_{t} \triangleq \beta_{1}m_{t-1} + (1 - \beta_{1})\g f(\xbold_{t}), \hspace{3mm} v_{t} \triangleq \beta_{2}v_{t-1} + (1 - \beta_{2})\left(\g (\xbold_{t}) \odot \g (\xbold_{t})\right)\]
+\[\hat{m}_{t} \triangleq \displaystyle \frac{m_{t}}{1 - \beta_{1}^{t}}, \hspace{3mm} \hat{v}_{t} \triangleq \displaystyle \frac{v_{t}}{1 - \beta_{2}^{t}}\]
+}
+\uncover<2->{Adam updates parameters based on the rule below:
+\begin{equation}
+\label{adam}
+\xbold_{t+1} = \xbold_{t} - \frac{\eta}{\sqrt{\hat{v}_{t}} + \epsilon} \odot \hat{m}_{t}
+\end{equation}
+where \(\epsilon\) is a small value to prevent divide by zero errors
+}
+\end{frame}
+
+\subsection{RMSProp}
+\begin{frame}{RMSProp} % add citations
+\uncover<1->{Define: \[g_{t} \triangleq \g f(\xbold_{t}) \odot \g f(\xbold_{t})\]
+\[\mathbb{E}[g]_{t} \triangleq 0.9\mathbb{E}[g]_{t-1} + 0.1g_{t}\]
+}
+\uncover<2->{RMSProp updates parameters based on the rule below:
+\begin{equation}
+\label{rmsprop}
+\xbold_{t+1} = \xbold_{t} - \frac{\eta}{\sqrt{\mathbb{E}[g]_{t} + \epsilon}} \odot \g f(\xbold_{t})
+\end{equation}
+where \(\epsilon\) is a small value to prevent divide by zero errors
+}
+\end{frame}
+
 \end{document}

### 2017-12-04 Project/Presentation/main.tex
[WIP] added bibliography file

@@
 \documentclass[10pt]{beamer}
+%\documentclass{article}
+%\usepackage{beamerarticle}
 \usepackage[style=verbose,backend=biber]{biblatex}
@@
 \newcommand{\citecomma}{\({}^{{}_{,}}\)}
-%\documentclass{article}
-%\usepackage{beamerarticle}
 \usetheme{CambridgeUS}
 \addbibresource{main.bib}
-\renewcommand{\citesetup}{\tiny}
+\renewcommand{\footnotesize}{\tiny}
 
@@
 \begin{itemize}
-\item<1->{Polyak and Nesterov have proposed methods which \textit{provide an acceleration} to the updates. % add citations
+\item<1->{Polyak\footfullcite{polyak} and Nesterov\footfullcite{nag} have proposed methods which \textit{provide an acceleration} to the updates.
           \begin{itemize}
@@
 \item<3->{Gradient descent is used as \textit{mini-batch gradient descent} in practice, which leads
-          to \textit{variance in the calculated mini-batch gradients}. Methods like SVRG, SAG and % add citations
-          SDCA have been proposed to help solve this problem.
+          to \textit{variance in the calculated mini-batch gradients}. Methods like SVRG\footfullcite{svrg}, SAG\footfullcite{sag} and % add citations
+          SDCA\footfullcite{sdca} have been proposed to help solve this problem.
           \begin{itemize}
@@
 \subsection{Adagrad}
-\begin{frame}{Adagrad} % add citations
+\begin{frame}{Adagrad\footfullcite{adagrad}} % add citations
 \uncover<1->{Define: \[G_{t} \triangleq \mathrm{diag}\left(\displaystyle \sum_{i=0}^{t} \g f(\xbold_{i}) \odot \g f(\xbold_{i})\right)\]}
@@
 \subsection{Adadelta}
-\begin{frame}{Adadelta} % add citations
+\begin{frame}{Adadelta\footfullcite{adadelta}} % add citations
 \uncover<1->{Define: \[g_{t} \triangleq \g f(\xbold_{t}) \odot \g f(\xbold_{t})\]
@@
 \subsection{Adam}
-\begin{frame}{Adam} % add citations
+\begin{frame}{Adam\footfullcite{adam}} % add citations
 \uncover<1->{Define: \[m_{t} \triangleq \beta_{1}m_{t-1} + (1 - \beta_{1})\g f(\xbold_{t}), \hspace{3mm} v_{t} \triangleq \beta_{2}v_{t-1} + (1 - \beta_{2})\left(\g (\xbold_{t}) \odot \g (\xbold_{t})\right)\]
@@
 \subsection{RMSProp}
-\begin{frame}{RMSProp} % add citations
+\begin{frame}{RMSProp\footfullcite{rmsprop}} % add citations
 \uncover<1->{Define: \[g_{t} \triangleq \g f(\xbold_{t}) \odot \g f(\xbold_{t})\]

### 2017-12-04 Project/Presentation/main.tex
[WIP] added trivial functions in slides

@@
 
+\section{Our work}
+\begin{frame}{Our work}
+\begin{itemize}
+\item<2->{Implementation of Polyak's heavy ball method(CM), Adagrad, Adadelta, Adam, RMSprop, effectively creating a suite of first-order optimization methods.
+          \begin{itemize}
+          \item<2->{Implemented in NumPy\footfullcite{numpy}}
+          \end{itemize}}
+\item<3->{Analyzing behaviour of GD, CM, Adagrad, Adadelta, Adam, RMSprop on selected pathological test functions for optimization}
+\item<4->{Experiments on Deep Neural Networks - Multi-layer perceptrons, Autoencoders and Generative Adversarial Networks\footfullcite{gan}}
+\end{itemize}
+\end{frame}
+
+\subsection{Pathological Functions}
+\begin{frame}{Functions which were used}
+For our initial experiments, we chose 6 functions\footfullcite{test-funcs}:
+\begin{itemize}
+\item<1->{Bohachevsky's Functions 1, 2, 3}
+\item<2->{Beale's Function}
+\item<3->{Rosenbrock's Function}
+\item<4->{Styblinsky-Tang's Function}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Behaviour with Bohachevsky's Function 1}
+\begin{equation}
+\label{B1}
+f(\xbold) = x_{1}^{2} + 2x_{2}^{2} - 0.3\cos(3\pi x_{1}) - 0.4\cos(4\pi x_{2}) + 0.7
+\end{equation}
+\end{frame}
+
+\begin{frame}{Behaviour with Bohachevsky's Function 2}
+\begin{equation}
+\label{B2}
+f(\xbold) = x_{1}^{2} + 2x_{2}^{2} - 0.3\cos(3\pi x_{1})\cos(4\pi x_{2}) + 0.3
+\end{equation}
+\end{frame}
+
+\begin{frame}{Behaviour with Bohachevsky's Function 3}
+\begin{equation}
+\label{B3}
+f(\xbold) = x_{1}^{2} + 2x_{2}^{2} - 0.3\cos(3\pi x_{1} + 4\pi x_{2}) + 0.3
+\end{equation}
+\end{frame}
+
+\begin{frame}{Behaviour with Beale's Function}
+\begin{equation}
+\label{BL}
+f(\xbold) = (1.5 - x_{1} + x_{1}x_{2})^{2} + (2.25 - x_{1} + x_{1}x_{2}^{2})^{2} + (2.625 - x_{1} + x_{1}x_{2}^{3})^{2}
+\end{equation}
+\end{frame}
+
+\begin{frame}{Behaviour with Rosenbrock's Function}
+\begin{equation}
+\label{RB}
+f(\xbold;a=1, b=25) = (1 - x_{1})^{2} + 25(x_{2} - x_{1})^{2}
+\end{equation}
+\end{frame}
+
+\begin{frame}{Behaviour with Styblinsky-Tang's Function}
+\begin{equation}
+\label{ST}
+\displaystyle f(\xbold;n=2) = \sum_{i=1}^{n=2} x_{i}^{2} - 16x_{i}^{2} + 5x_{i}
+\end{equation}
+\end{frame}
 \end{document}

### 2017-12-04 Project/Trivial Functions/script.py
get best params for trivial functions

@@ -9,6 +9,7 @@ for i in functions:
     for j in optims:
         subprocess.call(["python3", "main.py", "--optim", j, "--fn", i]) 
 
+best_files = open('best_values.txt', 'w')
 for directory_1 in functions:
     for directory_2 in optims:
         data = []
@@ -16,4 +17,6 @@ for directory_1 in functions:
         for filename in file_list:
             filepath = directory_1 +'/'+ directory_2 +'/'+ filename
             data.append(np.genfromtxt(filepath)[-1,1])
-        print(directory_1, directory_2, file_list[np.argsort(data)[-1]])
+        print(directory_1, directory_2, file_list[np.argsort(data)[0]])
+        best_files.write('{0}\t{1}\t{2}\n'.format(directory_1, directory_2, file_list[np.argsort(data)[0]][3:-7]))
+best_files.close()

### 2017-12-05 Project/Trivial Functions/functions.py
plotter added

@@ -51,12 +51,4 @@ def BL(X):
     term3 = (2.625 - x + x*y**3)**2
     return term1 + term2 + term3
 
-def ST(X):
-    """
-        Returns Styblinsky-Tang's function at a point
-    """
-    x = X[0]
-    y = X[1]
-    return x*4 - 16*(x**2) + 5*x + y**4 - 16*(y**2) + 5*y
-
-_func_dicts = {'B1': B1, 'B2': B2, 'B3': B3, 'RB': RB, 'BL': BL, 'ST': ST}
+_func_dicts = {'B1': B1, 'B2': B2, 'B3': B3, 'RB': RB, 'BL': BL}

### 2017-12-05 Project/Trivial Functions/gradients.py

@@ -67,14 +67,4 @@ def BL(X):
     dy  = cmn1*tmy1 + cmn2*tmy2 + cmn3*tmy3
     return np.array([dx, dy])
 
-def ST(X):
-    """
-        Returns gradient of Styblinsky-Tang's function at a point
-    """
-    x = X[0]
-    y = X[1]
-    dx = 4*x**3 - 32*x + 5
-    dy = 4*y**3 - 32*y + 5
-    return np.array([dx, dy])
-
-_grad_dicts = {'B1': B1, 'B2': B2, 'B3': B3, 'RB': RB, 'BL': BL, 'ST': ST}
+_grad_dicts = {'B1': B1, 'B2': B2, 'B3': B3, 'RB': RB, 'BL': BL}

### 2017-12-05 Project/Trivial Functions/main.py

@@ -8,21 +8,19 @@ from argparse import ArgumentParser as AP
 
 def give_init(function_name):
     if function_name == 'B1':
-        return np.array([-75, 50])
+        return np.array([-20, 10])
     elif function_name == 'B2':
-        return np.array([50, -50])
+        return np.array([10, -10])
     elif function_name == 'B3':
-        return np.array([-50, 50])
+        return np.array([-10, 10])
     elif function_name == 'BL':
         return np.array([1, 2])
     elif function_name == 'RB':
         return np.array([-2.1, 2.1])
-    elif function_name == 'ST':
-        return np.array([0, 0])
 
 p = AP()
 p.add_argument('--optim', required=True, help='Optimizers: GD | CM | Adam | Adagrad | Adadelta | RMSprop')
-p.add_argument('--fn', required=True, help='Function name: B1 | B2 | B3 | BL | RB | ST')
+p.add_argument('--fn', required=True, help='Function name: B1 | B2 | B3 | BL | RB')
 p.add_argument('--iter', default=1000, type=int, help='Number of iterations to run for')
 p = p.parse_args()
 
@@ -40,8 +38,10 @@ if p.optim == 'GD':
         i = 0
         X = give_init(p.fn)
         log_file = open('{0}/log_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')
+        iter_file = open('{0}/iter_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')
         while i < p.iter:
             log_file.write('{0}\t{1}\n'.format(i, f._func_dicts[p.fn](X)))
+            iter_file.write('{0}\t{1}\n'.format(round(X[0], 5), round(X[1], 5)))
             G = g._grad_dicts[p.fn](X)
             i, X = o._opt_dicts[p.optim](i, X, G, lr)
         log_file.close()
@@ -54,8 +54,10 @@ elif p.optim == 'CM':
         i = 0
         X = give_init(p.fn)
         log_file = open('{0}/log_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')
+        iter_file = open('{0}/iter_{1}_{2}.txt'.format(newDir, lr, p.fn), 'w')
         while i < p.iter:
             log_file.write('{0}\t{1}\n'.format(i, f._func_dicts[p.fn](X)))
[…]

### 2017-12-05 Project/Trivial Functions/optimizers.py

@@ -66,6 +66,7 @@ def Adagrad(t, X, G, SSG, lr, eps=1e-08):
     update = X - lr*G/(np.sqrt(SSG) + eps)
     return timestep, update, SSG
 
+# FIXTHIS
 def Adadelta(t, X, G, EG, EDX, lr, rho=0.9, eps=1e-08):
     """
         Function to perform one Adadelta step

### 2017-12-05 Project/Trivial Functions/plotter.py

@@ -1,0 +1,34 @@
+import sys
+import numpy as np
+import functions as f
+import gradients as g
+import optimizers as o
+from matplotlib import pyplot as plt
+from matplotlib.colors import LogNorm
+
+# cur_func is one of B1, B2, B3, RB and BL
+cur_func = sys.argv[1]
+color_dict = {'GD': 'black', 'CM': 'blue', 'Adagrad': 'gold', 'Adam': 'green', 'RMSprop': 'red'}
+
+# Plot the contour of the function
+if cur_func == 'B1' or cur_func == 'B2' or cur_func == 'B3':
+    X, Y = np.meshgrid(np.arange(-25, 25, 0.5), np.arange(-25, 25, 0.5))
+    plt.contour(X, Y, f._func_dicts[cur_func]([X, Y]), levels=np.linspace(np.amin(f._func_dicts[cur_func]([X, Y])), np.amax(f._func_dicts[cur_func]([X, Y])), 50))
+if cur_func == 'RB':
+    X, Y = np.meshgrid(np.arange(-3, 3, 0.05), np.arange(-6, 6, 0.1))
+    plt.contour(X, Y, f._func_dicts[cur_func]([X, Y]), levels=np.linspace(np.amin(f._func_dicts[cur_func]([X, Y])), np.amax(f._func_dicts[cur_func]([X, Y])), 50))
+if cur_func == 'BL':
+    X, Y = np.meshgrid(np.arange(-4.5, 4.5, 0.05), np.arange(-4.5, 4.5, 0.05))
+    plt.contour(X, Y, f._func_dicts[cur_func]([X, Y]), levels=np.logspace(0, 5, 50), norm=LogNorm(), cmap=plt.cm.jet)
+
+# First get the iterates
+opt_params = np.genfromtxt('best_values.txt', dtype=str)
+opt_params = opt_params[opt_params[:,0] == cur_func]
+title = ''
+for i in range(0, opt_params.shape[0]):
+    iterates = np.genfromtxt('{0}/{1}/iter_{2}_{0}.txt'.format(cur_func, opt_params[i,1], opt_params[i,2]))
+    plt.plot(iterates[:,0], iterates[:,1], color=color_dict[opt_params[i,1]], label=opt_params[i,1], linewidth=2.5)
+    title = title + '{0}: {1}   '.format(opt_params[i,1], iterates[-1])
+plt.title(title)
+plt.legend(loc='upper right')
+plt.show()

### 2017-12-05 Project/Trivial Functions/script.py

@@ -1,9 +1,9 @@
-import subprocess
 import os
+import subprocess
 import numpy as np
 
-functions = ['B1', 'B2', 'B3', 'BL', 'RB', 'ST']
-optims = ['GD', 'CM', 'Adam', 'Adagrad', 'Adadelta', 'RMSprop']
+functions = ['B1', 'B2', 'B3', 'BL', 'RB']
+optims = ['GD', 'CM', 'Adam', 'Adagrad', 'RMSprop']
 
 for i in functions:
     for j in optims:
@@ -14,9 +14,10 @@ for directory_1 in functions:
     for directory_2 in optims:
         data = []
         file_list = os.listdir(directory_1+'/'+directory_2)
+        file_list = list(filter(lambda x: x.find('log') != -1, file_list))
         for filename in file_list:
             filepath = directory_1 +'/'+ directory_2 +'/'+ filename
             data.append(np.genfromtxt(filepath)[-1,1])
         print(directory_1, directory_2, file_list[np.argsort(data)[0]])
-        best_files.write('{0}\t{1}\t{2}\n'.format(directory_1, directory_2, file_list[np.argsort(data)[0]][3:-7]))
+        best_files.write('{0}\t{1}\t{2}\n'.format(directory_1, directory_2, file_list[np.argsort(data)[0]][4:-7]))
 best_files.close()

### 2017-12-05 Project/Presentation/main.tex
plotter added
Presentation complete

@@
           \end{itemize}}
-\item<3->{Analyzing behaviour of GD, CM, Adagrad, Adadelta, Adam, RMSprop on selected pathological test functions for optimization}
+\item<3->{Analyzing behaviour of GD, CM, Adagrad, Adam and RMSprop on selected pathological test functions for optimization}
 \item<4->{Experiments on Deep Neural Networks - Multi-layer perceptrons, Autoencoders and Generative Adversarial Networks\footfullcite{gan}}
@@
 \begin{frame}{Functions which were used}
-For our initial experiments, we chose 6 functions\footfullcite{test-funcs}:
+For our initial experiments, we chose 5 functions\footfullcite{test-funcs}:
 \begin{itemize}
@@
 \item<3->{Rosenbrock's Function}
-\item<4->{Styblinsky-Tang's Function}
 \end{itemize}
@@
 \end{equation}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{./images/B1.eps}
+\end{figure}
 \end{frame}
@@
 \end{equation}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{./images/B2.eps}
+\end{figure}
 \end{frame}
@@
 \end{equation}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{./images/B3.eps}
+\end{figure}
 \end{frame}
@@
 \end{equation}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{./images/BL.eps}
+\end{figure}
 \end{frame}
@@
 \end{equation}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{./images/RB.eps}
+\end{figure}
 \end{frame}
 
-\begin{frame}{Behaviour with Styblinsky-Tang's Function}
-\begin{equation}
-\label{ST}
-\displaystyle f(\xbold;n=2) = \sum_{i=1}^{n=2} x_{i}^{2} - 16x_{i}^{2} + 5x_{i}
-\end{equation}
+\subsection{Deep Learning Experiments: MLP Classifiers}
+\begin{frame}{Multi-Layer Perceptron Classifiers}
+\begin{itemize}
+\item<1->{Took Multilayer perceptron classifiers for the MNIST\footnote{\url{http://yann.lecun.com/exdb/mnist/}} and CIFAR10\footnote{\url{https://www.cs.toronto.edu/~kriz/cifar.html}} datasets}
+\item<2->{Achieved \(\approx 95\%\) on MNIST and \(\approx 60\%\) on CIFAR10
+          \begin{itemize}
+          \item<3->{Not State-of-the-art, only to analyze the behaviour of the optimization methods}
+          \end{itemize}
+         }
+\item<3->{Activation used between the layers was \texttt{ReLU}}
+\item<4->{Trained for 10000 iterations, chose optimal learning rate from \(\{0.5, 0.2, 0.1, 0.05, 0.02, 0.01, 0.005, 0.002, 0.001, 0.0005, 0.0002, 0.0001\}\)}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{MNIST with Architecture: \texttt{784-500-500-500-10} - Training Loss vs Mini-batches}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{./images/mnist-1.pdf}
+\end{figure}
+\end{frame}
+
+\begin{frame}{MNIST with Architecture: \texttt{784-1000-1000-1000-10} - Training Loss vs Mini-batches}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{./images/mnist-2.pdf}
+\end{figure}
+\end{frame}
+
+\begin{frame}{CIFAR10 with Architecture: \texttt{3072-1000-500-10} - Training Loss vs Mini-batches}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{./images/cifar10-1.pdf}
+\end{figure}
+\end{frame}
+
+\begin{frame}{CIFAR10 with Architecture: \texttt{3072-1000-1000-1000-10} - Training Loss vs Mini-batches}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{./images/cifar10-2.pdf}
+\end{figure}
+\end{frame}
+
+\subsection{Deep Learning Experiments: Autoencoders}
+\begin{frame}{Deep Autoencoders}
+\begin{itemize}
+\item<1->{Used a multilayer perceptron deep autoencoder for encoding MNIST images}
+\item<2->{Activation used between the layers was \texttt{sigmoid}}
+\item<3->{Trained for 10000 iterations, chose optimal learning rate from \(\{0.5, 0.2, 0.1, 0.05, 0.02, 0.01, 0.005, 0.002, 0.001, 0.0005, 0.0002, 0.0001\}\)}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Autoencoder architecture: \texttt{784-1000-500-1000-10} - Training Loss vs Mini-batches}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{./images/auto-1.pdf}
+\end{figure}
+\end{frame}
+\begin{frame}{Autoencoder architecture: \texttt{784-1000-500-250-30-30-250-500-1000-10}\footfullcite{hinton} - Training Loss vs Mini-batches} % add citation
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{./images/auto-2.pdf}
+\end{figure}
+\end{frame}
+
+\subsection{Deep Learning Experiments: GANs}
+\begin{frame}{Deep Convolutional Generative Adversarial Networks\footfullcite{dcgan}} %add citation
+\begin{itemize}
+\item<1->{Used a specialized architecture for generator and discriminator for MNIST digits based on Deep Convolutional GANs}
+\item<2->{Run for 50000 iterations, and noted the variations of generator loss and discriminator loss with time for Adam optimizer and Nesterov's Accelerated Gradient}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Generator Loss with time}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{./images/genr-gan.pdf}
+\end{figure}
+\end{frame}
+\begin{frame}{Discriminator Loss with time}
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{./images/disc-gan.pdf}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Quality of Images Generated}
+\begin{columns}
+\column{0.475\linewidth}
+\centering
+\textbf{Using ADAM optimizer}
+\includegraphics[width=0.75\textwidth]{./images/adam-dcgan.png}
+\column{0.475\linewidth}
+\centering
+\textbf{Using Accelerated Gradient}
+\includegraphics[width=0.75\textwidth]{./images/nag-dcgan.png}
+\end{columns}
+\end{frame}
+
+\section{Conclusion}
+\begin{frame}{Our conclusions from the experiments conducted}
+\begin{itemize}
+\item<1->{Experiments show that adaptive gradient methods in fact, only have a marginal benefit}
+\item<2->{This concurs with the claim put forward by Wilson \textit{et al}\footfullcite{marginaladaptivegradients} about the marginal benefits of Adaptive Gradient Methods}
+\item<3->{Hype or Hallelujah? \emph{Hype} it is!!}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\Huge{Thank You\footnote{Code Available at \url{https://github.com/vishwakftw/CS6230-OML}}}
 \end{frame}

### 2017-12-05 Project/GANs/DCGAN.py

@@ -1,5 +1,5 @@
 import torch as t
-import gan_utils as u
+import gan_util as u
 import torch.nn as nn
 import torch.utils.data as d_utils
 import torchvision.utils as tv_utils

### 2017-12-05 Project/GANs/main.py

@@ -2,52 +2,45 @@ import sys
 import DCGAN as dc
 
 # Dataset
-dset	= sys.argv[1]
-root	= sys.argv[2]
+dset = sys.argv[1]
+root = sys.argv[2]
 
 # Get dataset
-if dset == 'mnist':
-	transformations	= [transforms.Scale(image_size), transforms.ToTensor()]
-	if normalize == True:
-		transformations.append(transforms.Normalize((0.5, ), (0.5, )))
-	dataset	= dset.MNIST(root=root, download=True, transform=transforms.Compose(transformations))
+if dset == 'MNIST':
+    transformations = [transforms.ToTensor()]
+    transformations.append(transforms.Normalize((0.5, ), (0.5, )))
+    dataset = dset.MNIST(root=root, download=True, transform=transforms.Compose(transformations))
     n_chan = 1
-elif dset == 'cifar10':
-	transformations	= [transforms.Scale(image_size), transforms.ToTensor()]
-	if normalize == True:
-		transformations.append(transforms.Normalize((0.5, 0.5, 0.5), (0.5, 0.5, 0.5)))
-	dataset = dset.CIFAR10(root=root, download=True, transform=transforms.Compose(transformations))
+elif dset == 'CIFAR10':
+    transformations = [transforms.ToTensor()]
+    transformations.append(transforms.Normalize((0.5, 0.5, 0.5), (0.5, 0.5, 0.5)))
+    dataset = dset.CIFAR10(root=root, download=True, transform=transforms.Compose(transformations))
     n_chan = 3
-	
+    
 # DCGAN object initialization
-image_size = 32
 n_z = 128
-hiddens	= {'gen':	64,
-		   'dis':	64
-		  }
 ngpu = 1
 loss = 'BCE'
 
-Gen_model	= dc.DCGAN(image_size=image_size, n_z=n_z, n_chan=n_chan, hiddens=hiddens, ngpu=ngpu, loss=loss)
+Gen_model = dc.DCGAN(arch=dset, n_z=n_z, ngpu=ngpu, loss=loss)
 
 # DCGAN training scheme
-batch_size	= 100
-n_iters		= 1e05
-opt_dets	= {'gen':	{'name'         : 'adam',
-				         'learn_rate'	: 1e-04,
-				         'betas'        : (0.5, 0.99)
-				        },
-    		   'dis':	{'name'		    : 'sgd',
-	        	   		 'learn_rate'	: 1e-04,
[…]

### 2017-12-05 Project/Trivial Functions/plotter.py

@@ -27,7 +27,7 @@ opt_params = opt_params[opt_params[:,0] == cur_func]
 title = ''
 for i in range(0, opt_params.shape[0]):
     iterates = np.genfromtxt('{0}/{1}/iter_{2}_{0}.txt'.format(cur_func, opt_params[i,1], opt_params[i,2]))
-    plt.plot(iterates[:,0], iterates[:,1], color=color_dict[opt_params[i,1]], label=opt_params[i,1], linewidth=2.5)
+    plt.plot(iterates[:,0], iterates[:,1], color=color_dict[opt_params[i,1]], label=opt_params[i,1], linewidth=3.5)
     title = title + '{0}: {1}   '.format(opt_params[i,1], iterates[-1])
 plt.title(title)
 plt.legend(loc='upper right')

### 2017-12-05 Project/Classifiers/utils.py
final commit for CS6230-OML

@@ -1,85 +1,0 @@
-import torch as t
-import torch.nn as nn
-import torch.optim as optim
-
-class Model(nn.Module):
-    def __init__(self, net_params, init_scheme):
-        """
-            Class to produce models based on JSON objects
-        """
-        super(Model, self).__init__()
-
-        activations = {'relu': nn.ReLU(True), 
-                       'sigmoid': nn.Sigmoid(),
-                       'tanh': nn.Tanh(),
-                       'leaky_relu': nn.LeakyReLU(0.2, inplace=True)
-                      }
-        self.conv_part = None
-        self.mlp_part = None
-        for i in range(0, len(net_params)):
-            if net_params[i]['type'] == 'CONV':
-                if self.conv_part is None:
-                    self.conv_part = nn.Sequential()
-                in_ = net_params[i]['params']['in']
-                out_ = net_params[i]['params']['out']
-                k_size = net_params[i]['params']['kernel']
-                stride = net_params[i]['params']['stride']
-                pad = net_params[i]['params']['padding']
-                self.conv_part.add_module('conv_{0}-{1}-{2}'.format(i, in_, out_), 
-                                          nn.Conv2d(in_channels=in_, out_channels=out_,
-                                                    kernel_size=k_size, stride=stride,
-                                                    padding=pad))
-                if 'activation' in net_params[i].keys():    
-                    self.conv_part.add_module('{0}_{1}'.format(i, net_params[i]['activation']), activations[net_params[i]['activation']])
-            elif net_params[i]['type'] == 'MLP':
-                if self.mlp_part is None:
-                    self.mlp_part = nn.Sequential()
-                in_ = net_params[i]['params']['in']
-                out_ = net_params[i]['params']['out']
-                self.mlp_part.add_module('linear_{0}-{1}-{2}'.format(i, in_, out_),
-                                         nn.Linear(in_, out_))
-                if 'activation' in net_params[i].keys():    
-                    self.mlp_part.add_module('{0}_{1}'.format(i, net_params[i]['activation']), activations[net_params[i]['activation']])
-
-        if init_scheme == 'xavier':
-            for p in self.mlp_part.parameters():
-                if len(p.size()) >= 2:
-                    p.data = nn.init.xavier_uniform(p.data)
-        elif init_scheme == 'he':
[…]

### 2018-09-30 Assignment-1/Solns-Assgn-1.tex
Port remaining solutions to LaTeX

@@
 \usepackage{amssymb}
+\usepackage{hyperref}
+
+\newcommand{\tr}{\mathrm{tr}}
+\newcommand{\deter}{\mathrm{det}}
 
@@
 
+\section*{Question 3}
+\begin{flushleft}
+Consider an affine transformation of \(x\) given by matrices \(A \in \mathbb{R}^{k x n}\) and \(b \in \mathbb{R}^{k}\).
+We know that the domain of \(g(x)\) is \(\mathbb{R}^{n}\), which is convex.
+Note that \(Ax\) is a matrix-vector product given by:
+\begin{equation}
+Ax = \begin{bmatrix} a_{1}^{T}x \\ a_{2}^{T}x \\ \vdots \\ a_{k}^{T}x \end{bmatrix}
+\end{equation}
+Hence, \(y = Ax + b\) can be given by \(y_{i} = a_{i}^{T}x + b_{i}\quad i \in \{1, 2, \ldots, k\}\).
+Using a result in Boyd and Vanderberghe (Pg.
+74), the canonical log-sum-exp function - \(f(x)\) is convex in \(x\).
+Now the composition of a convex and a convex function is convex: therefore \(g(x) = f(y(x)) = f(Ax + b)\) is convex.
+\(\newline\)
+
+An alternative solution would be to compute the Hessian of the function. First we have to compute the gradient. Denote \(\exp(a_{i}^{T}x + b_{i}) = E_{i}\):
+\begin{equation}
+\frac{\partial g}{\partial x_{j}} = \frac{\displaystyle \sum_{i=1}^{n} \exp(a_{i}^{T}x + b_{i})a_{ij}}{\displaystyle \sum_{i=1}^{k} \exp(a_{i}^{T}x + b_{i})} = \frac{\displaystyle \sum_{i=1}^{n} E_{i}a_{ij}}{\displaystyle \sum_{i=1}^{n} E_{i}}
+\end{equation}
+
+For a critical point:
+\begin{equation}
+\label{3-grad}
+\frac{\partial g}{\partial x_{q}} = \frac{\displaystyle \sum_{i=1}^{n} E_{i}a_{iq}}{\displaystyle \sum_{i=1}^{n} E_{i}} = 0
+\end{equation}
+
+Now compute the \(p;q^{th}\) element of the Hessian (ran out of symbols).
+\begin{equation}
+\frac{\partial^{2} g}{\partial x_{p} \partial x_{q}} = \underbrace{\frac{\displaystyle \sum_{i=1}^{k}E_{i}a_{iq}\sum_{i=1}^{k}E_{i}a_{ip}}{\displaystyle \sum_{i=1}^{k} (E_{i})^{2}}}_{= 0 \text{ due to Eqn \ref{3-grad}}} + \frac{\displaystyle \sum_{i=1}^{k}E_{i}a_{iq}a_{ip}}{\displaystyle \sum_{i=1}^{k} E_{i}}
+\end{equation}
+
+Now since the other term is only left, we have to show \(z^{T}Hz = \sum_{p,q=1}^{n}z_{p}z_{q}H_{pq} \geq 0\) to show that the function is convex.
+\begin{equation}
+\frac{n^2}{\displaystyle \sum_{i=1}^{k} E_{i}}\sum_{p=1}^{n}\sum_{q=1}^{n}\sum_{i=1}^{k}z_{p}z_{q}E_{i}a_{ip}a_{iq} = \frac{n^2}{\displaystyle \sum_{i=1}^{k} E_{i}}\sum_{i=1}^{k}E_{i}\sum_{p=1}^{n}z_{p}a_{ip}\sum_{q=1}^{n}z_{q}a_{iq} = \frac{n^2}{\displaystyle \sum_{i=1}^{k} E_{i}}\sum_{i=1}^{k}E_{i}\left(\sum_{p=1}^{n}z_{p}a_{ip}\right)^{2} \geq 0
+\end{equation}
+
+The last inequality comes due to the fact that \(E_{i} > 0 \forall i\), and \(x^{2} \geq 0 \forall x\).
+\end{flushleft}
+
 \section*{Question 4}
@@
 
+\section*{Question 5}
+\subsection*{Question 5.a.}
+\begin{flushleft}
+We adopt a similar proof technique as used in Boyd and Vanderberghe (Pg. 74).
+\(\newline\)
+
+Consider \(X(t) = Z + tV\) where \(Z \succ 0\) and \(V \in S^{n}\) (symmetric matrices). Let \(t\) be the set of all values such that \(X(t)\) is PD.
+
+Now since \(X(t)\) is PD, \(X(t)^{-1}\) exists, and it given by:
+\begin{equation}
+X(t)^{-1} = [Z + tV]^{-1} = (Z[I + tZ^{-1}V])^{-1} = [I + tZ^{-1}V]^{-1}Z^{-1}
+\end{equation}
+
+A result in Matrix perturbations states that (analogous to a negative binomial expansion):
+\begin{multline}
+(I + tZ^{-1}V)^{-1} = I + tZ^{-1}V + t^{2}Z^{-1}VZ^{-1}V + \ldots\\\Rightarrow (I + tZ^{-1}V)^{-1}Z^{-1} = Z^{-1} + tZ^{-1}VZ^{-1} + t^{2}Z^{-1}VZ^{-1}VZ^{-1} + \ldots
+\end{multline}
+
+Let \(g(t)\) be given by:
+\begin{equation}
+g(t) = \tr(X(t)^{-1}) = \tr(Z^{-1}) + t\tr(Z^{-1}VZ^{-1}) + t^{2}\tr(Z^{-1}VZ^{-1}VZ^{-1}) + \ldots
+\end{equation}
+
+Now, compute the derivative of \(g(t)\) w.r.t. \(t\):
+\begin{gather}
+\frac{\partial g(t)}{\partial t} = \tr(Z^{-1}VZ^{-1}) + 2t\tr(Z^{-1}VZ^{-1}VZ^{-1}) + \ldots \\
+\Rightarrow \frac{\partial^{2} g(t)}{\partial t^{2}} = 2\tr(Z^{-1}VZ^{-1}VZ^{-1}) + \ldots \\
+\Rightarrow \frac{\partial^{2} g(t)}{\partial t^{2}} \vert_{t=0} = 2\tr(Z^{-1}VZ^{-1}VZ^{-1})
+\end{gather}
+
+Denote \(W = VZ^{-1}\). Hence:
+\begin{equation}
+\frac{\partial^{2} g(t)}{\partial t^{2}} \vert_{t=0} = 2\tr(Z^{-1}VZ^{-1}VZ^{-1}) = 2\tr(W^{T}Z^{-1}W)
+\end{equation}
+
+If \(Z\) is PD, then \(Z^{-1}\) is PD, and \(Z^{-1}V\) and \(VZ^{-1}\) are PD.
+Hence: \(\frac{\partial^{2} g(t)}{\partial t^{2}} \vert_{t=0} = 2\tr(W^{T}Z^{-1}W) \geq 0\), since the product of PD matrices is PSD.
+This means that \(g(0) = \tr(Z)\) is convex and \(Z \in S_{n}^{++}\).
+\end{flushleft}
+
+\subsection*{Question 5.b.}
+\begin{flushleft}
+Using a similar approach to 5.a, consider \(Y(t) = Z + tV\) where \(Z \succ 0\) and \(V \in S^{n}\).
+\begin{multline}
+Y(t) = Z + tV = Z^{0.5}[I + tZ^{-0.5}VZ^{-0.5}]Z^{0.5} \Rightarrow \deter(Y(t)) = \deter(Z^{0.5})\deter([I + tZ^{-0.5}VZ^{-0.5}])\deter(Z^{0.5})\\= \deter(Z)\deter([I + tZ^{-0.5}VZ^{-0.5}]).
+\end{multline}
+
+Consider the eigenvalues of \(Z^{-0.5}VZ^{-0.5}\) to be \(\{\lambda_{1}, \lambda_{2}, \ldots, \lambda_{n}\}\).
+Hence the eigenvalues of \(I + tZ^{-0.5}VZ^{-0.5}\) are \(\{1 + t\lambda_{1}, 1 + t\lambda_{2}, \ldots, 1 + t\lambda_{n}\}\).
+\(\newline\)
+
+Let \(h(t) = \deter(Z)^{\frac{1}{n}}\deter([I + tZ^{-0.5}VZ^{-0.5}])^{\frac{1}{n}}\):
+\begin{equation}
+h(t) = \deter(Z)^{\frac{1}{n}}\deter([I + tZ^{-0.5}VZ^{-0.5}])^{\frac{1}{n}} = \deter(Z)^{\frac{1}{n}}\left(\prod_{i=1}^{n}(1 + t\lambda_{i})\right)^{\frac{1}{n}}
+\end{equation}
+
+The second term involving the product is a geometric mean.
+Citing a result in Boyd and Vanderberghe about the geometric mean being concave for positive numbers, we show that \(h(t)\) is concave \(\forall t \in T\).
+The proof is complete by considering \(t = 0\).
+\end{flushleft}
+
+\section*{Question 7}
+\begin{flushleft}
+Consider the derivative of the function \(f(z) = |z|\) to be
+\begin{equation}
+f'(z) = \begin{cases} 1 & \text{ if} z \geq 0 \\ -1 & \text{ otherwise} \end{cases}
+\end{equation}
+
+Therefore, the Hessian can be written as:
+\begin{equation}
+\begin{bmatrix}
+2a & \pm 1 \\ \pm 1 & 2a
+\end{bmatrix}
+\end{equation}
+
+We need the eigenvalues to be \(> 0\), hence by finding the characteristic polynomial and extracting the roots we get \(a > \pm 0.5\).
+We take the intersection of the two sets, to finally get \(a > 0.5\).
+\end{flushleft}
+
 \section*{Question 8}
@@
 Since the domain of \(f\) is \(\mathbb{R}^{n}_{++}\), we can most certainly assure that \(\nexists \lambda \in \Lambda\), such that \(\lambda \leq 0\).
-Hence it can be said that \(f\) is \(k\)-strongly convex with \(k = \min \Lambda\).
+Hence it can be said that \(f\) is strictly convex.
 \end{flushleft}
